### 2023-08-03 manuscript/manuscript.Rmd
Initial commit

---
output: 
  bookdown::pdf_document2:
    toc: false
    keep_tex: true
fontsize: 12pt
header-includes:
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{booktabs, multirow, tabularx, float}
bibliography: references.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      fig.pos = 'H')
```

# **Title:** Data-driven, matrix-wide molecular formula assignment for ultra-high resolution ESI-MS {-}

**Authors:**  Jasen P. Finch^1\*^, Thomas Wilson^1^, Laura Lyons^1^, Helen Phillips^1^, Manfred Beckmann^1^ & John Draper^1^

^\*^ Corresponding author

**Addresses:**

^1^ Department of Life Sciences, Aberystwyth University, Aberystwyth, SY23 3DA, UK

# Abstract {-}

## Background {-}

Compound identification remains one of the primary bottlenecks in the analysis of non-targeted mass spectrometry metabolomics data.
Some annotation approaches are available; however, these are primarily for liquid chromatography-mass spectrometry (LC-MS) profiling and are unsuitable for non-targeted metabolome fingerprinting techniques.

## Results {-}

```{r assignment-summary}
tar_load(assignment_results)
tar_load(assignment_matches)
tar_load(spike_urine_standards_MF_assignment_possible)
tar_load(spiked_urine_correct_assignments)

total_assigned <- round(sum(assignment_results$n_assignments) / 
                          sum(assignment_results$n_features) * 100,1)
total_mfs <- sum(assignment_results$n_MFs)
total_kegg <- sum(assignment_matches$KEGG_compound_matches)

spiked_urine_possible <- nrow(spike_urine_standards_MF_assignment_possible)
spiked_urine_correct_percent <- spiked_urine_correct_assignments$MF %>% 
  unique() %>% 
  length() %>% 
  {. / spiked_urine_possible * 100} %>% 
  round(digits = 1)
```

We present a data-driven ensemble approach for the ‘first pass’ automated assignment of molecular formulas to mass spectra acquired from high-resolution electrospray ionisation mass spectrometry (ESI-MS). This is primarily intended for metabolomics experiment data using non-targeted metabolome fingerprinting and LC-MS profiling.
The approach incorporates the use of *m/z* correlation networks, molecular formula generation and selection as well as graphically based component plausibility selection to assign molecular formulas to *m/z* features.
The R package *assignments* was developed as a user-friendly implementation of the approach and is freely available from https://github.com/aberHRML/assignments.
The approach was tested in human urine samples spiked with known chemical standards and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*. Across the three test matrices, `r total_assigned`% of the m/z features were assigned a molecular formula giving a total of `r total_mfs` molecular formulas and matching to `r total_kegg` compounds in the KEGG compound database. Of the `r spiked_urine_possible` spiked chemical standards that it was possible to assign in the human urine samples, the approach was able to assign m/z features to `r spiked_urine_correct_percent`% of these.

## Conclusions {-}

The approach presented here and its implementation in the R package *assignments* provides investigators with an efficient and automated approach for matrix-wide ‘first pass’ assignment of molecular formulas to non-targeted ESI-MS metabolomics experiment data.

# Introduction

Metabolite identification remains one of the main bottlenecks both in time and resources in the analysis of non-targeted mass spectrometry based metabolomics approaches.
Electrospray ionisation mass spectrometry (ESI-MS) is preferred as this is a soft ionisation technique that enables easier compound identification [@domingo2018].
Irrespectively, it is possible for a single metabolite to be represented within ESI-MS spectra by multiple ionisation product adducts and isotopes [@draper2013].

Ultra-high resolution MS instrumentation utilising Orbitrap mass analyser technology is now commonplace in many laboratories, which can yield many thousands of *m/z* features in a single analytical run [@hu2005].
This has provided an unprecedented opportunity for metabolite annotation and has been applied to approaches such as 'first pass' flow infusion electrospray high-resolution MS (FIE-HRMS) fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
However, mass accuracy alone does not meet the requirements for metabolite annotation [@kind2006].

Metabolite annotation is a crucial aspect for ensuring that biological interpretation is possible from metabolomics data sets.
A number of strategies can be employed to annotate peaks in mass spectra.
These can include simple approaches such as heuristic filtering of molecular formulas generated from *m/z* or direct matching putative ionisation products to metabolite databases [@draper2009; @green2015; @kind2007];
but also ensemble approaches that integrate a number of strategies like correlation analysis, ionisation product scoring and biotransformations [@baran2013; @fernandez2014; @kuhl2012; @uppal2017].
Tandem MS can also be employed for LC-MS approaches and provides a convenient method for structural validation [@neumann2010].
However, this is unsuitable for metabolome fingerprinting-based techniques due to the presence of structural isomers and isobaric compounds.
Many of the available software packages implementing these strategies are suitable for LC-MS techniques only.

For investigators, manual consideration of putative annotations is time-consuming and often difficult to interpret due to the complex interactions between ions and molecules and their mathematical possibilities forming *m/z* features.
It is important that software package implementations can automate this process but also provide tractable results and simple input parameters for easy and wide applicability.

Correlation analysis has been commonly used to identify related metabolites within metabolomics data and is often used in LC-MS approaches to group related *m/z* features [@steuer2006;@kuhl2012]. 
These analyses can be represented as weighted undirected network graphs where the *m/z* features reflect the graph nodes and the significant correlations between them, the graph edges [@gehlenborg2010].
Applying Spearman's rank correlation allows the detection of both linear and monotopic associations between *m/z* features.
Adduct and isotopic relationships would be expected to have positive linear correlations whereas compounds associated by pathway transformations within living tissues under the control of gene and protein regulation would be able to have both positive or negative monotonic correlations [@camacho2005].

The use of correlations to associate relationships between *m/z* features provides an essential layer of data-driven, internal validation of the mathematically derived solutions.
For instance, if two features that are posited to have the same composition (eg. a C^13^ relationship) but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about associations (eg. metabolic pathway changes) of those features with other putatively assigned features within the matrix.

Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS fingerprinting and LC-MS profiling-based metabolomics data. 
This is aimed at biological experiments that include numerous replicates of sample classes for the assignment of low accurate mass (< 1000 atomic mass units (amu)) metabolite elemental compositions of biological origin.
It is an ensemble approach that combines a number of annotation strategies including correlation analysis, molecular formula selection, adduct and isotopic scoring and graphical component selection.
This is accompanied by the development of the R package *assignments* as an implementation of the approach.

# Materials and methods

## Biological sample preparation and ESI-MS analysis

```{r chemical-standards}
tar_load(spiked_urine_compound_info)

spiked_urine_compound_MFs <- spiked_urine_compound_info %>% 
  select(MF) %>% 
  distinct() %>% 
  nrow()
```

Two example sample matrices were used for assessing the molecular formula assignment approach. 
The first was human urine spiked with a diverse mixture of `r nrow(spiked_urine_compound_info)` chemical standards from a previously validated assay that included `r spiked_urine_compound_MFs` unique molecular formulas [@beckmann2020].
A full list of the names and InChI chemical identifiers of the spiked chemical standards can be found in Additional file 1.
The second was healthy leaf tissue samples collected from four ecotypes of the model grass species *Brachypodium distachyon* using the method described in @parker2008.
Further details of the sample collection and extraction are described in the Additional file 2.

The sample matrices were each analysed using FIE-HRMS fingerprinting and C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling.
These data are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949.
The FIE-HRMS fingerprinting data were spectrally processed by performing spectral binning using the R package *binneR* v`r packageVersion('binneR')` [@finch2022].
Spectral processing of the C18-UHPLC-HRMS profiling data was conducted using the R package *xcms* v`r packageVersion('xcms')` [@smith2006].
The data were then pre-treated by performing batch correction, feature occupancy filtering, imputation, relative standard deviation filtering based on QC samples and total ion count normalisation using the R package *metabolyseR* v`r packageVersion('metabolyseR')` available from GitHub (https://github.com/jasenfinch/metabolyseR).
Further details of the ESI-MS analyses and data preparation are available in Additional file 2.
All data processing and analysis were performed using the R Statistical Programming Language version `r stringr::str_c(version$major,version$minor,sep ='.')` [@Rteam2022].

## Molecular formula assignment

For each biological sample matrix, the input data for molecular formula assignment had undergone spectral processing and pre-treatment and consisted of a sample by *m/z* feature intensity matrix containing both positive and negative ionisation mode data. 
Molecular formula assignment was conducted using the R package *assignments* v`r packageVersion('assignments')` and the approach comprised four major steps which are outlined in Figure \@ref(fig:figure-1-approach).
Where assignment was possible, *m/z* features were designated with unambiguous molecular formulas, isotopes and ionisation adduct assignments. 
The following sections outline the methodology behind each major step of the approach.

\linespread{1}

```{r figure-1-approach,fig.height=6,fig.cap='An overview of the approach for assigning molecular formulas to \\textit{m/z} features from ultra-high resolution ESI-MS metabolomics experiments.'}
tar_read(approach)
```

\linespread{2}

### Correlation analysis

Spearman's rank correlation was conducted between all pairwise combinations of *m/z* features.
A Bonferroni correction was applied to the p-value of each coefficient based on the total number of features.
Correlations with a p-value below 0.05 and a minimum absolute coefficient of 0.7 were retained.

### Relationship calculation {#relationships}

```{r adducts}
tar_load(brachy_FIE_HRMS_parameters_molecular_formula_assignment)

negative_mode_adducts <-  brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::adducts() %>% 
  .$n %>% 
  stringr::str_replace_all(']1-',']^1-^') %>% 
  stringr::str_replace_all(']2-',']^2-^') %>% 
  stringr::str_replace_all('Cl37','^37^Cl') %>% 
  str_as_written_list()
positive_mode_adducts <-   brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::adducts() %>% 
  .$p %>% 
  stringr::str_replace_all(stringr::coll(']1+'),']^1+^') %>% 
  stringr::str_replace_all(stringr::coll(']2+'),']^2+^') %>% 
  stringr::str_replace_all('K41','^41^K') %>% 
  str_as_written_list()
```

Possible adduct, isotopic and transformation mass relationships were calculated for each pair of correlated *m/z*.
The pre-ionisation masses ($M$) were first calculated for all adduct, isotopic and transformation combinations for each of the *m/z*. 
An $M$ for a given ionisation product *m/z*, adduct, isotope and transformation was calculated using the following equation:

$$M = \frac{c(m/z - \Delta m_a) - \Delta m_i}{n} - \Delta m_t$$

Where $c$ is the overall ionisation product charge, $\Delta m_a$ is the adduct mass change, $\Delta m_i$ is the isotopic mass change, $n$ is the number of molecules involved in the ionisation product and $\Delta m_t$ is the transformation mass change.

The negative ionisation mode adducts included: `r negative_mode_adducts`.
The positive ionisation mode adducts included: `r positive_mode_adducts`.
All adduct mass, charge and molecule information can be found in Additional file 3 (Table S2).
The isotopes included single and double ^13^C and ^18^O for the *B.distachyon* matrices and additionally ^34^S for the spiked urine.
The isotopic mass information used can be found in Additional file 3 (Table S3).
For all matrices, `r nrow(mzAnnotation::transformation_rules)` transformations were based on the transitions of common metabolic reactions found in primary metabolism.
These transformations can be found in Additional file 3 (Table S4).

To calculate the mass error for possible relationships between the *m/z* in the correlated pair, the mass differences between all the possible $M$ values of the *m/z* were computed.
All relationships with an $M$ difference in mass error below 0.001 amu and a maximum of one transformation were retained.

### Adduct and isotopic assignment iterations

This first assignment step used only positively correlated feature relationships that corresponded to adduct and isotopic relationships without transformation. 
This assignment step consisted of molecular formula generation, molecular formula selection (Section \@ref(MF-selection)) and then multiple iterations of network graph component assembly and network graph component selection (Section \@ref(component-selection)).

### Molecular formula generation and selection {#MF-selection}

For all the relevant *m/z* relationships, molecular formulas were generated from all the possible $M$ values based on these relationships.
To reduce the computational requirements for molecular formula generation with increasing molecular mass, the possible $M$ values were limited to a maximum of 800 amu.

Generated molecular formulas were restricted to only include the elements C, H, N, O, P and S.
For a given $M$, the maximum allowed element frequencies for generated molecular formulas were restricted using the following terms: $C = M/12$, $H = 2C$, $N,O = C/2$, $P,S = C/4$.
The calculated frequencies were rounded to the nearest integer value.
Molecular formula generation was conducted for each $M$ value within a mass deviation range of $\pm$ 6 ppm.
This generation was exhaustive, using the brute force approach of the HR2 generator, without the application of the golden rules proposed by @kind2007.

To allow the most plausible molecular formulas to be selected, a percentage plausibility score was calculated based on 12 heuristic checks of the rules 2, 4, 5 and 6 from @kind2007.
These included the LEWIS and SENIOR checks, five element ratio checks and five element probability checks.
For each molecular formula, a point was awarded for each check that passed or was not applicable.
These points were then totalled and divided by 12 to give a golden rules check score ($S_{GR}$).
The proportion of the atoms in the molecule that were C, H or O was then calculated to give a CHO proportion score ($S_{CHO}$).
A percentage plausibility score ($P_{MF}$) was then calculated for the molecular formula using the equation below.

$$P_{MF} = \frac{S_{GR} + S_{CHO}}{2} \times 100  $$
The three top-ranked molecular formulas with the highest $P_{MF}$ scores, generated for a given $M$, were selected for network component assembly.

### Graphical component selection {#component-selection}

For each selected molecular formula, a component subgraph was constructed from the correlation network graph using all matching features, with features allowed across multiple components.
All components with a size of less than one and containing only isotopic features or adducts were removed.

The components were ranked from most plausible to least plausible using a component plausibility score $P_c$. 
This was based on the component size, the average component edge weight and an additional score based on the plausibility of the adduct and isotopic combinations of the features within the component (AIS).

To calculate $AIS$, the possible adducts within each ionisation mode were ranked from most probable to least probable and assigned a score ($a$) with the most probable being $n-1$, $n$ being the number of adducts within that ionisation mode and least probable adduct a score of 0. 
The adduct rank scores used for the assignment were based on the adduct order specified in Section \@ref(relationships) for all matrices.

Similarly, for the possible isotopes, these were also ranked from most (non-isotopic) to least probable and assigned a score ($i$) as for the adducts.
The isotopic rank scores used for the assignment of the *B.distachyon* matrices were as follows: Non-isotopic = 3, ^13^C = 2, ^18^O = 1, ^13^C2 = 0.
For the spiked urine matrices, the isotopic rank scores were: Non-isotopic = 4, ^13^C = 3, ^18^O = 2, ^34^S = 1, ^13^C2 = 0.

The $AIS$ for a given adduct and isotopic combination ($k$) was then calculated using:

$$AIS_k = \frac{a_k + i_k}{\max{a} + \max{i}}$$

The $AIS$ scores within each component were totalled and divided by half the total number of adduct and isotope combinations to give a component $AIS$ score ($AIS_c$).

For each component, the component plausibility score $P_c$ was calculated using:

$$P_c = d \times AIS_c \times w$$
Where $d$ is the component degree and $w$ is the average component edge weight. 
The higher $P_c$, the more plausible the component subgraph.

Features shared between multiple components were sequentially eliminated and retained for only the components with the highest $P_c$.
For any ties in $P_c$ between components, the component with the highest $P_{MF}$ was selected.
Further ties were decided by selecting the component with the lowest ppm error.
This graphical component selection routine was repeatedly performed across multiple iterations until no more assignments were obtained.

### Transformation assignment iterations

The transformation assignment iterations consisted of multiple assignment iterations where only positively or negatively correlated relationships containing a previously assigned feature and a transformation were included.
Each iteration was performed as for the previous adduct and isotopic assignment iteration and repeated until no more assignments were obtained.

## Assignment assessments

```{r KEGG-compounds}
tar_load(KEGG_compounds_valid_MFs)
```

Features matching to possible ionisation products of the spiked urine compounds were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
The calculated *m/z* were then matched to the spiked urine features using a search range of $\pm$ `r spiked_urine_search_ppm` ppm.

The molecular formulas and monoisotopic masses for compound entries in the KEGG compound database were accessed using the R package *KEGGREST* v`r packageVersion('KEGGREST')` [@kanehisa2000;@tenenbaum2021].
For the assessments of molecular formula selection success, only uncharged molecular formulas containing the elements C, H, N, O, P, or S and a monoisotopic mass below 1000 amu were retained.
This gave a total of `r nrow(KEGG_compounds_valid_MFs)` unique molecular formulas.

For the assignments in *B.distachyon* and human urine matrices, KEGG compounds specific to each organism were identified by cross-referencing KEGG compound IDs with the KEGG enzyme database for *B. distachyon*.
For these compounds, InChI chemical identifiers were obtained by cross-referencing their entries in the PubChem Substance database with the PubChem Compound database [@kim2016].
To match possible KEGG compounds specific to each organism to the molecular formulas assigned to *m/z* features, the organism-specific compound molecular formulas were first filtered to compounds only with molecular formulas that were assigned to *m/z* features.
By applying the MZedDB ionisation rules, only KEGG compounds were retained for which the adducts of the assigned features were possible.

# Results and discussion

## *assignments*: An R package for molecular formula assignment of ESI-MS metabolomics data

The R package *assignments* was developed as an implementation of this molecular formula assignment approach.
It is available to install from GitHub at https://aberhrml.github.io/assignments/.
The package is fully documented and a usage guide is provided for new users.

The input data consists of a sample by accurate *m/z* feature intensity matrix with at least four samples.
This data should be spectrally processed and appropriately pre-treated prior to assignment.
All assignment parameters are fully customisable, which includes the specification of custom adducts, isotopes and transformations.
It is possible for all the assignment criteria and results to be recovered for each assignment step, enabling the manual curation of assignments by the user.
The default parameters provided by the package have been designated to enable assignment for a wide range of biological applications.

Parallel processing is provided by the *future* R package, which allows the user to specify a number of different parallel backends [@bengtsson2021].
Benchmarking of the processing time of assignments made to the spiked urine sample matrix was performed by the *assignments* R package using varying numbers of CPU workers (Additional file 2).
With the use of high-performance computing resources with greater than 32 CPU workers resulted in assignments being obtained in under two hours for 250,000 correlations between *m/z* features (Additional file 3 (Table S5)).

## Molecular formula assignment success {#success}

```{r assignment-success}
tar_load(spiked_urine_compound_PIPs)
tar_load(spiked_urine_feature_matches)
tar_load(spiked_urine_unassigned)
tar_load(spiked_urine_assigned)
tar_load(spiked_urine_incorrect_assignments)

brachy_shared_MFs <- tar_read(brachy_FIE_HRMS_results_molecular_formula_assignment) %>% 
  assignments::assignments() %>% 
  select(MF) %>% 
  inner_join(tar_read(brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment) %>% 
               assignments::assignments() %>% 
               select(MF),
             by = 'MF') %>% 
  distinct() %>% 
  nrow()

brachy_shared_kegg <- intersect(
  tar_read(brachy_FIE_HRMS_KEGG_matches)$MF %>% 
    unique(),
  tar_read(brachy_RP_UHPLC_HRMS_KEGG_matches)$MF %>% 
    unique()
) %>% length()

positive_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
 assignments::adducts() %>% 
  .$p %>% 
  length()

negative_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
 assignments::adducts() %>% 
  .$n %>% 
  length()

spiked_urine_MF_PIPs <- spiked_urine_compound_PIPs %>% 
  select(MF,Adduct,Isotope) %>% 
  distinct()

spiked_urine_feature_matches_MF_n <- spiked_urine_feature_matches$MF %>% 
  unique() %>% length()

percent_unassigned <- {nrow(spiked_urine_unassigned) / 
    nrow(spiked_urine_feature_matches) * 100} %>% 
  signif(digits = 3)

spiked_urine_unassigned_MF_n <- spiked_urine_unassigned %>% 
  anti_join(
    spiked_urine_assigned %>% 
      select(MF) %>% 
      distinct(), 
    by = 'MF') %>% 
  .$MF %>% 
  unique() %>% 
  length()

spiked_urine_correct_assignments_MF_n <- spiked_urine_correct_assignments %>%
  .$MF %>% 
  unique() %>% 
  length()

spike_urine_standards_MF_assignment_impossible_n <- spiked_urine_feature_matches_MF_n -
  nrow(spike_urine_standards_MF_assignment_possible)

percent_standards_MF_assigned <- {spiked_urine_correct_assignments_MF_n / 
  nrow(spike_urine_standards_MF_assignment_possible) * 100} %>% 
  signif(digits = 3)
```

An overview of the molecular formula assignments made in both the urine and *B. distachyon* FIE-HRMS matrices is provided in Table \@ref(tab:assignment-results), which shows that similar proportions of *m/z* features were assigned molecular formulas.
However, there was more than double the relative molecular formula assignment coverage for the C18-UHPLC-HRMS data of *B. distachyon* compared to the FIE-HRMS matrices.
The chromatographic separation of the C18-UHPLC-HRMS increases the likelihood that a number of correlated ionisation products will be present for a given compound due to a reduction in the effects of ion suppression [@gowda2014].
There is also a decreased chance of interference in the correlations between related features due to the presence of structural isomers and isobaric compounds.
A total of `r brachy_shared_MFs` molecular formulas were assigned in both the FIE-HRMS and C18-UHPLC-HRMS data of *B. distachyon*.

\linespread{1}

```{r assignment-results}
tar_read(assignment_results) %>%
  left_join(tar_read(assignment_matches), 
            by = c("matrix", "technique")) %>% 
  mutate(matrix = paste(matrix,technique)#,
  ) %>%
  mutate(n_assignments = glue::glue(
    '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
  select(-technique) %>% 
  dplyr::rename(`\\textit{m/z} features` = n_features,
                correlations = n_correlations,
                relationships = n_relationships,
                `assigned features` = n_assignments,
                `molecular formulas (MF)` = n_MFs,
                `KEGG compound matches` = KEGG_compound_matches,
                `KEGG MF matches` = KEGG_MF_matches,
                `matched chemical standards` = matched_standards) %>% 
  tidyr::gather(Variable,Value,-matrix) %>% 
  mutate(Value = replace(Value,
                         Value == 'NA',
                       'N/A')) %>% 
  tidyr::spread(matrix,Value) %>% 
  slice(c(1,3,8,2,7,5,4,6)) %>%
  set_colnames(stringr::str_split_fixed(colnames(.),
                                        ' ',2)[,2]) %>% 
  knitr::kable(format = 'latex',
               booktabs = TRUE,
               linesep = '',
               escape = FALSE,
               align = c(
                 rep('l',1),
                 rep('r',3)
               ),
               caption = 'A summary of the molecular formula assignment results for the example sample matrices. The number of matched chemical standards is given as the number of unique molecular formulas.') %>% 
  kable_styling(latex_options = "HOLD_position") %>% 
  add_header_above(c(
    " ", 
    "\\\\textit{B. distachyon} leaf" = 2, 
    "Human urine" = 1),
    escape = FALSE)
```

\linespread{2}

Although there were proportionately more molecular formulas assigned to *m/z* features in the C18-UHPLC-HRMS compared to the FIE-HRMS *B. distachyon* matrix, there was a disproportionate number of these molecular formulas that matched to compounds in the KEGG database specific to *B. distachyon* and able to form the relevant ionisation products (Table \@ref(tab:assignment-results)).
However, only `r brachy_shared_kegg` of the molecular formulas matched to compounds in the KEGG database were found to be common between the two techniques.
The low number of assigned molecular formulas partly reflects the incompleteness of the KEGG database, especially with respect to the secondary metabolites that are prolific in C18-UHPLC-HRMS [@breitling2013].
The assignments for all the matrices along with the matched KEGG compound IDs are provided in Additional file 4.

To assess the assignments in the example urine matrix with respect to the spiked chemical standards, the predicted ionisation products were first identified by directly matching these to the observed *m/z*.
Based on the `r positive_mode_adducts + negative_mode_adducts` adducts and four isotopic rules specified, there were `r nrow(spiked_urine_MF_PIPs)` predicted ionisation product *m/z* for these chemical standards.
A total of `r nrow(spiked_urine_feature_matches)` of these ionisation products could be directly matched to the observed *m/z* features which included `r spiked_urine_feature_matches_MF_n` of the `r spiked_urine_compound_MFs` unique molecular formulas of the spiked chemical standards. 
These directly matched *m/z* features are provided in Additional file 5.
For `r nrow(spiked_urine_unassigned)` (`r percent_unassigned`%) of these matched *m/z* features, no molecular formula was assigned by the approach, which included `r spiked_urine_unassigned_MF_n` of the molecular formulas of the spiked chemical standards.
However, `r nrow(spiked_urine_assigned)` of the matched features were assigned a molecular formula with `r nrow(spiked_urine_correct_assignments)` of these corresponding to one of the spiked compound ionisation products. This included `r spiked_urine_correct_assignments_MF_n` of the spiked chemical standards molecular formulas (Table \@ref(tab:assignment-results)).
The *m/z* features assigned to molecular formulas corresponding to the spiked chemical standards are provided in Additional file 6.

The remaining `r nrow(spiked_urine_incorrect_assignments)` matched *m/z* features were assigned alternative molecular formula ionisation products.
The *m/z* of the predicted ionisation features that it was not possible to assign molecular formulas were as a result of either not being correlated with other *m/z* features that corresponded to a correct adduct or isotope mass difference or whose subgraph components were not selected, with more plausible solutions instead selected based on the correlation network.

Accounting for the `r spike_urine_standards_MF_assignment_impossible_n` chemical standards that were not assigned due to a lack of relevant correlated mass relationships, the algorithm was able to assign molecular formulas to features corresponding to `r percent_standards_MF_assigned`% of the `r nrow(spike_urine_standards_MF_assignment_possible)` possible chemical standards.
This highlights the difficulty in annotating these complex biological mixtures containing hundreds of compounds, which can create many thousands of possible ionisation products.
Interference in correlations due to the presence of isobaric compounds and MS instrument noise can weaken correlations between features, particularly in those with low abundance.

## The molecular formula assignment approach
```{r sample-size}
sample_size_n_features <- 4000
sample_size <- pwr::pwr.r.test(
  r = 0.7,
  power = 0.8,
  sig.level = 0.05/sample_size_n_features)$n %>% 
  round()
```

```{r mf-selection}
tar_load(KEGG_compound_MF_ranks)

kegg_top_rank_percent <- KEGG_compound_MF_ranks %>% 
  mutate(top_10 = Rank <= 10) %>% 
  group_by(top_10) %>% 
  summarise(N = n()) %>% 
  mutate(percent = N / sum(N) * 100) %>% 
  {.$percent[.$top_10 == TRUE]} %>% 
  signif(digits = 3)
```

The use of correlation analysis to identify related features requires a suitable number of observations to ensure that statistically significant associations can be detected. 
This is especially important in non-targeted metabolomics data sets that can contain thousands of *m/z* features and p-value adjustment is required to account for multiple testing. 
For example, to detect a significant lower limit of Pearson's correlation coefficient of 0.7 for a matrix containing `r sample_size_n_features` *m/z* features, an approximate minimum sample size of `r sample_size` observations would be required (p < 0.05, Bonferroni corrected) [@cohen2013]. 
Metabolomics fingerprinting and profiling experiments typically require high replication of 10-15 observations per class in order to discriminate between the sample classes [@enot2008]. 
Therefore, a sample set comparing just four classes brings this to within a practical range for applying this approach in many cases, such as the example *B. distachyon* ecotype comparison sample set provided here.
The assignments obtained from the *B. distachyon* example matrix also show the application of the assignment approach to sample sets that contain multiple underlying experimental classes.

The approaches of the *MFAssignR* R package and *Formularity* software provide alternative means of assigning molecular formulas to both fingerprinting and LC-MS data, with *Formularity* utilising a compound identification database to aid molecular formula selection [@schum2020;@tolic2017].
While these approaches provide the advantage of being able to assign molecular formulas using a single replicate with substantially low computational demands, they do not incorporate the use of correlations between *m/z* features.
This can lead to incorrect assignments and the potential miss assignment of isotopic formulae to uncorrelated features.
Both the *CAMERA* and *MAIT* R packages utilise correlations for adduct and isotope annotation but only for LC-MS data [@fernandez2014;@kuhl2012].
The MAIT package also includes the addition of biotransformations to improve peak annotation performance.

Assignment approaches using molecular formula selection strategies have been modifications of the seven golden rules proposed by @kind2007 to identify criteria for chemically correct formulae.
These include the assignment approach of the R package *MFAssignR* and formula filtering for PubChem database matching [@lommen2014; @schum2020].
These approaches apply the heuristic rules discretely where molecular formulas are removed if they do not meet the criteria.
The scoring approach used here for selection builds on this use of discrete rules by enabling generated molecular formulas to be ranked in terms of their plausibility. It also allows the inclusion of non-carbon containing molecules such as phosphate and sulphate that are common in biological systems.
The effectiveness of this ranking is reflected in the fact that, of the `r nrow(KEGG_compound_MF_ranks)` valid KEGG compound molecular formulas, `r kegg_top_rank_percent`% of these formulas could be ranked within the top 10 for plausibility when molecular formulas were generated for the compound accurate masses inside a range of `r KEGG_alternative_MFs_ppm` ppm.
This is also shown in Figure \@ref(fig:figure-2-kegg-mf-selection) where the plausibility ranking is still effective even for masses above 500 amu where it is possible for many thousands of formulas to be generated.

\linespread{1}

```{r figure-2-kegg-mf-selection,fig.height=4,fig.cap='Molecular formula plausibility ($P_{MF}$) rankings of KEGG compound molecular formulas. The $P_{MF}$ rankings were based on the monoisotopic masses of 6745 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 6 ppm.'}
tar_read(KEGG_compound_MF_ranks_mass_plot) +
  tar_read(KEGG_compound_MF_ranks_n_plot)
```

\linespread{2}

As shown in Figure \@ref(fig:figure-3-feature-solutions), there can be multiple graphical molecular formula component solutions can exist for a given *m/z* feature.
The component plausibility score was the primary criterion for which components were selected. 
This score combines information about how connected the features are within a given component, the average strength of these connections and the expected commonality of the adduct and isotopic combinations present.
By using this information about the connections of *m/z* within the spectrum, it is possible for an *m/z* to be assigned a less plausible molecular formula if it is more highly connected with strong connections containing more common adduct and isotopic combinations.
Combining information between the ionisation modes further increases available information and more informative components [@kuhl2012].

\linespread{1}

```{r figure-3-feature-solutions,fig.height=5,fig.width=8,fig.cap = 'Graphical component solutions for the spiked urine negative mode feature 80.97488 \\textit{m/z}. The components were generated during the first adduct and isotopic assignment iteration. The selected graphical component is bordered in red. The values of attributes are shown below each component. The nodes for the negative mode feature 80.97488 \\textit{m/z} are highlighted in blue.',dev='png',dpi=300}
tar_read(spiked_urine_feature_solutions_plot)
```

\linespread{2}

Adduct formation will vary depending on the sample origin and the frequency of isotopes will be dependent on the elemental compositions of the constituent compounds [@kruve2017].
The input rankings of the adducts and isotopes ranks used will alter the preference of selected components so gains can possibly be made in assignment accuracy with the optimisation of these rankings on a per sample type basis.

Although examples used here include FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using non-targeted ultra-high resolution mass accuracy ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
Similarly, data from other chromatographic techniques for LC-HRMS profiling could be used such as hydrophilic interaction chromatography (HILIC), where suitable adducts and adduct rankings could instead be specified [@spagou2010].

This approach was designed for use with any sample tissues and fluids of biological origin and analysed using non-targeted ultra-high resolution mass accuracy ESI-MS data.
The additional input of knowledge for transformations specific to the biological sample can increase assignment success.
However, the approach could be applied without the use of transformations where these are not relevant.
The assigned molecular formulas could be putatively matched to compound databases such as KEGG using ionisation rules such as those of the MZedDB database and then apply functional enrichment analyses using a tool such as the R package *FELLA* [@draper2009;@picart2018].
They could also be used to provide some information about precursor ions in LC-MS/MS fragmentation analyses.

# Conclusions

The approach presented here, and its implementation in the *assignments* R package, provides investigators with an automated means to assign molecular formulas to non-targeted ESI-MS metabolomics with the use of high-performance computing resources.

Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted ionisation products could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS metabolome fingerprinting.
It was also possible to assign molecular formulas of common primary and secondary metabolites present in the KEGG compound database to *m/z* features from leaf tissue of *B. distachyon* using both FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques.

This approach can contribute to considerably reducing the workload of 'first pass' putative metabolite identification in non-targeted metabolomics analyses.
Assignments can be used as inputs for enrichment analyses, providing a fast route to the biological interpretation of inherently complex data sets.

# Declarations

## Ethics approval and consent to participate

Not applicable

## Consent for publication

Not applicable

## Availability of data and materials

The metabolomics data files reported in this paper are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949 (https://www.ebi.ac.uk/metabolights/MTBLS6949). 
The R code use to analyse the data and generate the manuscript is available on GitHub at https://github.com/jasenfinch/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS.

## Competing interests

The authors declare that they have no competing interests.

## Funding

JF received funding from the charity Woodland Heritage. JF and LL were funded by The Biotechnology and Biological Science Research Council (BBSRC), Natural Environmental Research Council (NERC), Defra and Scottish Government for funding as part of the BAC-STOP and FUTURE OAK projects (grant refs:BB/T010886/1 , BB/T01069X/1).  The financial support of HP was provided by Aberystwyth University and the UK Medical Research Council (MRC) funded TW and MB (MRC grant ref: MR/S010483/1).

## Authors' contributions

JF: conceptualisation, methodology, sample preparation, software development, data analysis and writing-original draft; TW: conceptualisation and manuscript revision; LL: sample preparation and data collection; HP: data collection; MB: conceptualisation, supervision and manuscript revision; JD: supervision and manuscript revision.

## Acknowlegements

Not applicable

```{r word-count}
# jfmisc::writingChecks('manuscript/manuscript.Rmd',
#                       checks = 'word count')
```

# Additional files

Additional file 1.csv - Table S1 -
Chemical identifiers of the compounds spiked into the urine extract.

Additional file 2.pdf - Supplementary materials and methods -
Further details outlining the materials and methods used in this article.

Additional file 3.pdf - Supplementary tables -
Supplementary tables S2, S3, S4 and S5 showing adduct, isotopic, transformation and R package benchmarking information respectively.

Additional file 4.csv - Table S6 -
The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.

Additional file 5.csv Table S7 -
The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.

Additional file 6.csv Table S8 -
The *m/z* features assigned to spiked chemical compounds in urine.

# References

### 2023-08-03 manuscript/supplementary_methods.Rmd
Initial commit

---
output: pdf_document
bibliography: references.bib
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

# Supplementary materials and methods

## Sample collection and extraction protocols

### *Brachypodium distachyon* tissue collection and extraction

Plants of *Brachypodium distachyon* ecotypes ABR1, ABR5, ABR6 and Bd21 were grown for 21 days in environmentally controlled growth rooms (Polysec, R. J. Hicks Refrigeration, Aberystwyth, UK) at 23 ^$\circ$^C under a 16 hour light period. 
A total of 15 plants of each ecotype were harvested by detaching the 2nd fully developed leaf from the base of the plant and dissecting the middle 4 cm segment. 
Each sample was then placed in a 2 mL Eppendorf tube containing a 4mm steel ball bearing and snap frozen in liquid nitrogen.
Samples were stored at -80 ^$\circ$^C until extraction. 
All harvesting was conducted in the growth room to avoid environmental fluctuations.

The frozen samples were milled using a Retsch 301 Mixer Mill at 30 Hz for 30 seconds and then placed on crushed ice. 
700 $\mu$L of pre-chilled extraction solvent (CHCl~3~:MeOH:H~2~O; 1:2.5:1; v:v:v) was immediately added and suspended by vortexing (Scientific Industries Vortex Genie-2). 
Samples were placed on an orbital shaker (FATSN002, Favorgen Biotech Corp) for 20 minutes at 1,400 rpm and a temperature of 4 ^$\circ$^C. 
After shaking, the samples were centrifuged (EBA 12R, Hettich) at 13,000 rpm for 6 minutes at 0 ^$\circ$^C. 
The supernatant of each sample was then transferred to a new 2 mL Eppendorf tube and the pellet was discarded. 
The samples were stored at -80 ^$\circ$^C until MS analysis.

### Human urine extraction and addition of chemical standards

Mixed gender, unfiltered pooled human urine (BioIVT, Catalog# HUMANURINEPNN, Lot# 0393918) from healthy donors was extracted  by adding 500 $\mu$L of pre-chilled MeOH (Extraction Grade, Fisher Scientific) and vortexing. 
Samples were placed on an orbital shaker (FATSM002, Favorgen Biotech Corp) for 20 minutes at 1,400 rpm and 4 ^$\circ$^C in the dark. 
The samples were centrifuged (EBA 12 R, Hettich) at 13,000 rpm for 5 minutes at 4 ^$\circ$^C and 20 $\mu$L of the supernatant was transferred to a clean glass HPLC vial, containing a 200 $\mu$L flat bottom micro insert (Chromacol) and diluted with 80 $\mu$L of H~2~O:MeOH (3:7) directly in the vial.

The `r tar_read(spiked_urine_compound_info) %>% nrow()` chemical standards were dissolved in a solvent mixture of methanol, water and formic acid (70%:30%:0.1%), each 4.5 $\mu$g/ml concentration.
700 $\mu$L of urine and 300 $\mu$L of the standards mixture were combined to produce the spiked urine sample.

## MS analysis

FIE-HRMS mass spectra were acquired on an Exactive Orbitrap (ThermoFinnigan, San Jose CA) mass spectrometer coupled to an Accela (ThermoFinnigan, San Jose CA) ultra-performance liquid chromatography (LC) system for the *B. distachyon* samples and the Human urine on a Q-Exactive Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system. 
A sample volume of 20 $\mu$L was injected and delivered to the electrospray ionisation (ESI) source via a 'plug' flow solvent of pre-mixed HPLC grade MeOH (Fisher Scientific) and ultra-pure H~2~O (18.2 $\Omega$) at a ratio of 7:3. 
The initial flow rate was 60 $\mu$L min^-1^ for 0.4 minutes and increased to 200 $\mu$L min^-1^ over 0.8 minutes. 
The flow rate was maintained at 200 $\mu$L min^-1^ for 0.3 minutes then increased to 600 $\mu$L min^-1^ over 1.5 minutes. 
Then the flow rate was then returned to 60 $\mu$L min^-1^ for 0.5 minutes. 
The total gradient time was 3.5 minutes. The capillary temperature and voltage were set at 270 ^$\circ$^C and 37.50 kV respectively.

For the *B. distachyon* samples, mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes. 
Positive ions were acquired between 55.000 and 1000.000 *m/z*, and  63.000 and 1000.000 *m/z* for negative ions.
Mass spectra for the spiked human urine were acquired with two scan events between 55-280 *m/z* and 270-1200 *m/z* for each of the positive and negative acquisition modes.
For all scan events, the scan rate was 1 Hz with a mass resolution of 100,000. 
The automatic gain control (AGC) target was 5 x10^5^ and the total ion injection time 250 ms.

Each biological matrix was analysed in separate analytical runs.
For the spiked urine, a total of 50 replicate injections were performed.
Three blank injections were preformed prior to the analytical replicate injections, using the solvent mixture used in the sample preparation of the particular matrix. 
For the *B. distachyon* samples, the samples were organised into randomised blocks, with each ecotype class represented by two samples in each block.
Similarly to the spiked urine samples, three blank injections were performed first, followed by three quality control samples (QC) injections.
The QC samples were made up of equal parts of each of the 60 samples in the run.
A QC sample injection then followed each of the random blocks.

C18-UHPLC-HRMS analyses were performed on an Orbitrap Fusion Tribrid mass spectrometer (Thermo Scientific) that was coupled to a Ultimate 3000 liquid chromatography tower (Dionex, Thermo Scientific). 
Mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes between a range of 100-2000 m/z. For all scan events, the scan rate was 1 Hz with a mass resolution of 100,000. The automatic gain control (AGC) target was 5 x10^5.

A Hypersil Gold reverse phase C18 column (2.1 mm x 150 mm; particle size 1.9 $\mu$m) was used for chromatography, which was maintained at a temperature of 60 $^\circ$C. 
The mobile phases consisted of ultra-pure water - formic acid (100:0.1) (A), LC-MS grade methanol -formic acid (100:0.1) (B) and HPLC grade isopropyl alcohol - LCMS grade methanol (1:1) (C). 
The initial condition was A:B (99:1.0) for 0.5 minutes and the percentage of B increased linearly over 6 minutes, to 60.0%. 
The percentage of B was increased further for another 4 minutes to 100% and held for 2 minutes. 
C was then introduced over 0.1 minutes to 100% and held for 1.4 minutes.
Prior to re-equilibration, 100% C was switched to 100% B in 0.1 minutes and then returned to the starting conditions of A:B (99:1).
This was held for 3.4 minutes giving a total method time was 17 minutes. 
The flow rate was kept at 400 $\mu$L min^-1^ except for between 12.1 and 13.6 minutes where it was reduced to 300 $\mu$L min^-1^.

## Spectral processing

After the MS data were acquired, the raw mass spectra data files (.RAW, ThermoFinnigan) were converted to the universal mass spectrometry open file format, mzML.
The conversion and vendor centroiding were performed using *msconvert* from the Proteowizard tool pack.

The FIE-HRMS fingerprinting data were spectrally processed by performing spectral binning to 0.01 amu using the R package *binneR* v`r packageVersion('binneR')`.
The infusion scans were detected using a total ion count threshold of above 50% of the maximum total ion count across the infusion profile.

For C18-UHPLC-HRMS, spectral processing was performed using the R package *xcms* v`r packageVersion('xcms')`.
Peak detection used the *centWave* algorithm, with the tolerated deviation between consecutive scans set at 8.2 parts per million (ppm), a minimum and maximum peak width of 16 and 43 seconds respectively and a signal to noise threshold of 7. 
The *Obiwarp* algorithm was used for retention time correction with a bin size of 0.7, a gap opening penalty of 2.9 and a gap enlargement penalty of 2.7. 
Peak grouping was performed using peak density, with a bin size of 0.00775 amu, a chromatographic bandwidth of 5 seconds and a minimum group fraction of two thirds. 
The *fillPeaks* method was used to integrate the areas of missing peaks using the default parameters.

## Data pre-treatment

The pre-treatment routines were performed separately for positive and negative ionisation mode features, with the features subsequently combined into a single intensity matrix.

For FIE-HRMS of the spiked urine, *m/z* were first filtered based on an overall occupancy threshold of two thirds.
All missing values were then imputed using random forest imputation. Only *m/z* features with a relative standard deviation (RSD) value below 50% were retained and finally a total ion count (TIC) normalisation was applied to each sample.

For the FIE-HRMS of the *B. distachyon* samples, a median centring batch correction was first performed where the median of each feature for each randomised block of samples were adjusted to the overall median.
The QC samples were used to calculate the RSD of each feature across the run batch.
These samples were first prepared by retaining only features with an occupancy above two thirds, and imputing missing values using random forest imputation.
Features with an RSD below 50% were retained and the QC samples were subsequently removed.
A maximum occupancy feature filter was then applied where features were retained if at least a single ecotype class had an occupancy of above two thirds of the samples.
Class-wise imputation using random forest was performed on features with a class occupancy greater than two thirds. 
Finally, a TIC normalisation was applied to each of the samples.

For the C18-UHPLC-HRMS of the *B. distachyon* samples, these data were pre-treatment as for FIE-HRMS fingerprinting data with the exception that a RSD filtering threshold of 20% was instead applied.

## Performance benchmarking of the *assignments* R package

```{r benchmarking-info}
targets::tar_load(cpu_information)
targets::tar_load(memory_information)
n_reps <- xfun::numbers_to_words(n_reps)
n_correlations <-  n_correlations %>% 
  prettyNum(big.mark=",",
            scientific=FALSE) %>% 
  str_as_written_list()
n_cores <- n_cores %>% 
  str_as_written_list()
```

Processing time was benchmarked for all the combinations of maximum number of correlations (`r n_correlations`) and CPU workers (`r n_cores`).
For each test, the molecular formula assignment routine was performed on the pre-treated spiked urine data matrix with all other parameters as the defaults using the R package *assignments* v`r packageVersion('assignments')`.
Tests were conducted on a Dell high performance computing blade with an `r cpu_information$Value[cpu_information$Type == 'Model name']` model processor with `r cpu_information$Value[cpu_information$Type == 'CPU(s)']` CPUs, `r memory_information$Size[memory_information$Type == 'MemTotal'] %>% as.character()`B of RAM and `r osVersion` operating system.

The R package *rbenchmark* v`r packageVersion('rbenchmark')` was used to measure processing time with `r n_reps` replications of each correlation and CPU workers combination.

### 2023-08-03 manuscript/supplementary_tables.Rmd
Initial commit

---
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \newcommand{\beginsupplement}{\setcounter{table}{1}  \renewcommand{\thetable}{S\arabic{table}}}
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE)
```

\beginsupplement

```{r table-s2}
tar_load(brachy_FIE_HRMS_parameters_molecular_formula_assignment)

adducts <-  brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::adducts() %>% 
  purrr::flatten_chr()

mzAnnotation::adduct_rules() %>% 
  dplyr::filter(Name %in% adducts) %>% 
  dplyr::select(
    Adduct = Name,
    Charge,
    `No. molecules` = xM,
    `Mass difference` = Add
  ) %>% 
  mutate(`Mass difference` = round(`Mass difference`,5)) %>% 
  knitr::kable(caption = 'Information for positive and negative mode adducts used for calculating ionisation products.',
               format = 'latex',
               booktabs = TRUE)
```

```{r table-s3}
mzAnnotation::isotope_rules() %>% 
  dplyr::filter(Isotope %in% c('13C','13C2','18O','34S')) %>% 
  dplyr::select(Isotope,
                `Mass difference` = `Mass Difference`) %>%
  dplyr::mutate(Isotope = c('\\textsuperscript{13}C',
                            '\\textsuperscript{13}C2',
                            '\\textsuperscript{18}O',
                            '\\textsuperscript{34}S')) %>% 
  knitr::kable(caption = 'Isotopic mass difference information.',
               format = 'latex',
               escape = FALSE,
               booktabs = TRUE)
```

```{r table-s4}
transformations <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
  assignments::transformations()

mzAnnotation::transformation_rules() %>% 
  dplyr::filter(`MF Change` %in% transformations) %>% 
  dplyr::select(
    Transformation = Name,
    `MF change` = `MF Change`,
    `Mass difference` = `Difference`
  ) %>% 
  knitr::kable(caption = 'Transformation mass difference information.',
               format = 'latex',
               booktabs = TRUE)
```

```{r table-s5}
tar_read(processing_time_results) %>% 
  dplyr::select(workers,
                `Number of correlations` = `# correlations`,`time (seconds)`) %>% 
  dplyr::mutate(`time (minutes)` = `time (seconds)` / 60,
                dplyr::across(tidyselect::where(is.numeric),~signif(.x,digits = 3)),
                `Number of correlations` = scales::comma(
                  `Number of correlations`,
                )) %>% 
  dplyr::select(-`time (seconds)`) %>% 
  tidyr::spread(workers,`time (minutes)`) %>% 
  knitr::kable(
    format = 'latex',
    booktabs = TRUE,
    caption = 'Processing time (minutes) results for the R package \\textit{assignments}. Feature \\textit{m/z} were used from spiked urine matrix and the numbers of correlations limited by using the correlations with the highest absolute coefficients up to the designated number. Molecular formula assignment was then performed across the range of numbers of central processing unit (CPU) workers.',
    align = rep('r',4)
  ) %>% 
  kableExtra::kable_styling() %>% 
  add_header_above(c(" ", "CPU workers" = 3))
```

### 2023-08-03 R/functions/KEGGmatches.R

@@ -1,0 +1,13 @@
+
+KEGGmatches <- function(mf,adduct,IRs){
+  rule_expr <- mzAnnotation::adduct_rules() %>% 
+    dplyr::filter(Name == adduct) %>% 
+    .$Rule %>% 
+    rlang::parse_expr()
+  
+  IRs %>%
+    cheminf::filterMF(mf) %>% 
+    cheminf::filterIP(rule = !!rule_expr) %>% 
+    cheminf::entries() %>% 
+    mutate(Adduct = adduct)
+}

### 2023-08-03 R/functions/approach.R

@@ -1,0 +1,56 @@
+
+assignmentApproach <- function(){
+  nodes <- tibble::tibble(
+    title = c(
+      '',
+      'Correlation analysis',
+      'Relationship calculation',
+      'Adduct and isotopic assignment iterations',
+      'Transformation assignment iterations',
+      ''),
+    body = c(
+      '*m/z* feature intensity matrix',
+      'Create a network graph of linked features.',
+      'Calculation of isotopic, adduct and transformation relationships between correlated features.',
+      'Component subgraphs are extracted based on molecular formulas generated from ionisation products of related features. Feature assignments are selected by sequential elimination from components based on component plausibility scores.',
+      'Identify features that are linked to assigned nodes by correlations with a transformation relationship. Multiple iterations are performed to allow possible assignments to propagate through the correlation network as more features are assigned.',
+      'Assigned *m/z* features'
+    ),
+    x = 1,
+    y = c(5,4.3,3.5,2.5,1.25,0.25)
+  ) %>% 
+    dplyr::mutate(label = body %>% 
+             stringr::str_wrap(65) %>% 
+             stringr::str_replace_all('\\n','<br>') %>% 
+             {glue::glue('**{title}**<br>{.}')} %>% 
+               stringr::str_replace_all(stringr::coll('****<br>'),'')
+             )
+  
+  edges <- tibble::tibble(
+    x = c(1),
+    xend = c(1),
+    y = c(4.85,4.07,3.22,2.07,0.76),
+    yend = c(4.53,3.79,2.93,1.75,0.39)
+  )
+  
+  ggplot2::ggplot() +
+    ggtext::geom_richtext(
+      data = nodes,
+      ggplot2::aes(x = x,
+                   y = y,
+                   label = label),
+      size = 3) +
+    ggplot2::geom_segment(
+      data = edges,
+      ggplot2::aes(
+        x = x,
+        xend = xend,
+        y = y,
+        yend = yend
+      ),
+      arrow = ggplot2::arrow(length = ggplot2::unit(2,'mm')),
+      size = 0.5
+    ) +
+    # ggplot2::theme_bw()
+    ggraph::theme_graph()
+}

### 2023-08-03 R/functions/benchmarking.R

@@ -1,0 +1,62 @@
+memUsage <- function(feature_data,max_cors,n_cores,interval){
+  
+  if (n_cores > 1) {
+    future::plan('multicore',workers = n_cores)  
+  } else {
+    future::plan('sequential')
+  }
+  
+  parameters <- assignments::assignmentParameters('FIE-HRMS')
+  parameters@correlations_parameters$maxCor <- max_cors
+  
+  mem_usage <- profvis::profvis({
+    assignment <- assignments::assignMFs(
+      feature_data,
+      parameters) 
+  },
+  interval = interval) %>%
+    {
+      .$x$message$prof 
+    } %>%
+    tibble::as_tibble() %>%
+    {
+      tibble::tibble(
+        min = min(.$memalloc),
+        max = max(.$memalloc)) %>%
+        dplyr::mutate(
+          `RAM usage (MB)` = max - min,
+          workers = n_cores,
+          `# correlations` = max_cors) %>%
+        dplyr::select(-min,-max)
+    }
+  
+  return(mem_usage)
+}
+
+procBenchmark <- function(feature_data,max_cors,n_cores,reps){
+  if (n_cores > 1) {
+    future::plan('multicore',workers = n_cores)  
+  } else {
+    future::plan('sequential')
+  }
+  
+  parameters <- assignments::assignmentParameters('FIE-HRMS')
+  parameters@correlations_parameters$maxCor <- max_cors
+  
+  proc_time <- rbenchmark::benchmark(
+    {
+      assignment <- assignments::assignMFs(
+        feature_data,
+        parameters) 
+    },
+    replications = reps,
+    columns = c('replications',
+                'elapsed')
+  ) %>%
+    tibble::as_tibble() %>% 
+    dplyr::rename(`time (seconds)` = elapsed) %>% 
+    dplyr::mutate(workers = n_cores,
+           `# correlations` = max_cors)
+  
+  return(proc_time)
+}

### 2023-08-03 R/functions/plotFeatureSolutions.R

@@ -1,0 +1,37 @@
+plotFeatureSolutions <- function(assignment,components,feature,iteration){
+  components %>%
+    rowwise() %>% 
+    group_split() %>% 
+    map(~{
+      component_stats <- assignments:::component(
+        assignment,
+        .x$component,
+        iteration,
+        type = 'all'
+      ) %>% 
+        assignments::nodes() %>% 
+        dplyr::filter(Feature == feature) %>% 
+        mutate(across(where(is.numeric),~signif(.x,digits = 3)))
+      
+      assignments::plotComponent(
+        assignment,
+        .x$component,
+        iteration,
+        type = 'all',
+        highlight = feature,
+        border = .x$border,
+        label_size = 2.5,
+        axis_offset = 0.2
+      ) +
+        labs(  caption = glue::glue('
+          P<sub>c</sub> = {component_stats$`Component Plausibility`};
+          Degree = {component_stats$Degree};
+          AIS<sub>c</sub> = {component_stats$AIS};
+          
+          P<sub>MF</sub> = {component_stats$`MF Plausibility (%)`}%;
+          Δ ppm = {component_stats$`PPM error`}')
+        )
+    }) %>% 
+    patchwork::wrap_plots() +
+    patchwork::plot_layout(guides = 'collect')
+}

### 2023-08-03 R/functions/str_as_written_list.R

@@ -1,0 +1,10 @@
+str_as_written_list <- function(string_vector,sep = ', '){
+  seperator <- sep
+  
+  string_vector[1:(length(string_vector) - 2)] <- stringr::str_c(string_vector[1:(length(string_vector) - 2)],seperator)
+  string_vector[length(string_vector)] <- stringr::str_c(' and ',string_vector[length(string_vector)])
+  
+  string_vector <- stringr::str_c(string_vector,collapse = '')
+  
+  return(string_vector)
+}

### 2023-08-03 R/functions/tar_KEGG_matches.R

@@ -1,0 +1,40 @@
+
+tar_KEGG_matches <- function(name,db){
+  envir <- tar_option_get("envir")
+  tidy_eval <- tar_option_get("tidy_eval")
+  
+  name <- tar_deparse_language(rlang::enexpr(name))
+  assignments_name <- paste0(name,'_results_molecular_formula_assignment') %>% 
+    rlang::sym()
+  IRs_name <- rlang::enexpr(db)
+  
+  results_name <- paste0(name,'_KEGG_matches')
+  
+  results_expr <- rlang::expr(!!assignments_name %>% 
+                                assignments::assignments() %>% 
+                                dplyr::select(MF,Adduct) %>% 
+                                dplyr::distinct() %>% 
+                                dplyr::rowwise() %>% 
+                                dplyr::group_split() %>% 
+                                purrr::map_dfr(
+                                  ~KEGGmatches(.x$MF,.x$Adduct,!!IRs_name)
+                                ) %>% 
+                                dplyr::distinct() %>% 
+                                dplyr::rowwise() %>% 
+                                dplyr::mutate(MF = cheminf::smilesToMF(SMILES)) %>% 
+                                dplyr::ungroup()
+  )
+  
+  command_results <- tar_tidy_eval(
+    results_expr,
+    envir = envir,
+    tidy_eval = tidy_eval
+  )
+  
+  target_results <- tar_target_raw(
+    results_name,
+    command_results
+  )
+  
+  return(list(target_results))
+}

### 2023-08-03 R/functions/validMF.R

@@ -1,0 +1,4 @@
+
+validMF <- function(MF){
+  grepl("^(([CHONPS]*)(([0-9])*))+$",MF) 
+}

### 2023-08-03 R/targets/KEGG_compound_targets.R

@@ -1,0 +1,120 @@
+
+KEGG_compound_IRs <- tarchetypes::tar_map(
+  values = tibble(
+    organism = c('human','brachy'),
+    KEGG_id = c('hsa','bdi')
+  ),
+  ## Identify KEGG compounds found organism
+  tar_target(
+    KEGG_IDs,
+    {
+      enzymes <- KEGGREST::keggLink(KEGG_id,'enzyme') %>%
+        names()
+      KEGGREST::keggLink('compound','enzyme') %>%
+        {tibble::tibble(Enzyme = names(.),
+                        Compound = .)} %>%
+        dplyr::filter(Enzyme %in% enzymes) %>%
+        dplyr::select(Compound) %>%
+        dplyr::distinct() %>%
+        tibble::deframe() %>%
+        stringr::str_remove_all('cpd:')
+    }
+  ),
+  ## Filter InChIs to only include specified organism
+  tar_target(
+    specific_KEGG_InChIs,{
+      future::plan('multicore',workers = 16)
+      
+      specific_KEGG_InChIs <- KEGG_InChIs %>% 
+        dplyr::filter(ID %in% KEGG_IDs) %>% 
+        tidyr::drop_na() %>% 
+        dplyr::rowwise() %>% 
+        group_split() %>% 
+        furrr::future_map_dfr(~{
+          .x %>%
+            dplyr::mutate(
+              SMILES = cheminf::convert(InChI,
+                                        'INCHI',
+                                        'SMILES'))
+        })
+      
+      future::plan('multicore',workers = 64)
+      
+      specific_KEGG_InChIs
+    }
+  ),
+  
+  ## Calculate the chemical descriptors for orgainism specific KEGG compounds
+  tar_target(
+    KEGG_IRs,
+    specific_KEGG_InChIs %>%
+      dplyr::mutate(NAME = "") %>% 
+      cheminf::metaboliteDB(.)
+  ),
+  
+  
+  names = 'organism'
+)
+
+KEGG_compound_targets <- list(
+  ## Retrieve KEGG compounds
+  tar_target(
+    KEGG_compounds,
+    KEGGREST::keggList('compound') %>% 
+      names() %>% 
+      stringr::str_remove('cpd:') %>% 
+      tibble::tibble(ENTRY = .) %>% 
[…]

### 2023-08-03 R/targets/MF_selection_targets.R

@@ -1,0 +1,73 @@
+
+KEGG_alternative_MFs_ppm <- 6
+
+MF_selection_targets <- list(
+  
+  ## Calculate golden rules for KEGG MFs
+  KEGG_compounds_valid_MFs_golden_rules = tar_target(
+    KEGG_compounds_valid_MFs_golden_rules,
+    KEGG_compounds_valid_MFs$MF %>% 
+      mzAnnotation::goldenRules()
+  ),
+  
+  ## Calculate KEGG MF plausibilities
+  KEGG_compound_plausibilities = tar_target(
+    KEGG_compound_plausibilities,
+    KEGG_compounds_valid_MFs_golden_rules %>% 
+      mzAnnotation::goldenRulesScore()
+  ),
+  
+  ## Generate alternative KEGG MF plausibilities
+  KEGG_alternative_MFs = tar_target(
+    KEGG_alternative_MFs,
+    KEGG_compounds_valid_MFs %>% 
+      dplyr::rowwise() %>% 
+      dplyr::group_split() %>% 
+      furrr::future_map(~ mzAnnotation::ipMF(.x$M,
+                                             adduct = 'M',
+                                             ppm = KEGG_alternative_MFs_ppm) %>% 
+                          dplyr::mutate(Rank = rank(100 - `Plausibility (%)`,
+                                                    ties.method = 'min'),
+                                        `KEGG MF` = .x$MF),
+                        .options = furrr::furrr_options(seed = TRUE))
+  ),
+  
+  ## Retrieve the MF plausibility ranks of the KEGG compounds
+  KEGG_compound_MF_ranks = tar_target(
+    KEGG_compound_MF_ranks,
+    KEGG_alternative_MFs %>% 
+      furrr::future_map(~{
+        .x %>% 
+          dplyr::mutate(N = nrow(.)) %>% 
+          dplyr::filter(MF == `KEGG MF`) %>% 
+          dplyr::select(MF,`Measured M`,`Plausibility (%)`,Rank,N)
+      }) %>% 
+      dplyr::bind_rows()
+  ),
+  
+  ## Plot the KEGG compound MF ranks against the exact mass
+  KEGG_compound_MF_ranks_mass_plot = tar_target(
+    KEGG_compound_MF_ranks_mass_plot,
+    KEGG_compound_MF_ranks %>% 
+      ggplot2::ggplot(ggplot2::aes(x = Rank,
+                                   y = `Measured M`)) +
[…]

### 2023-08-03 R/targets/assignment_results_targets.R

@@ -1,0 +1,48 @@
+
+assignment_results_targets <- list(
+  tar_target(
+    assignment_results,
+    list(
+      `spiked-urine_FIE-HRMS` = spiked_urine_results_molecular_formula_assignment,
+      `brachy_FIE-HRMS` = brachy_FIE_HRMS_results_molecular_formula_assignment,
+      `brachy_C18-UHPLC-HRMS` = brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment
+    ) %>%   map(~tibble(
+      n_features = ncol(.x@data),
+      n_correlations = nrow(.x@correlations),
+      n_relationships = nrow(.x@relationships),
+      n_assignments = .x %>% 
+        assignments::assignments() %>% 
+        nrow(),
+      n_MFs = .x %>% 
+        assignments::summariseAssignments() %>% 
+        nrow()
+    )) %>% 
+      bind_rows(.id = 'matrix') %>% 
+      tidyr::separate(matrix,
+                      c('matrix','technique'),sep = '_')
+  ),
+  
+  tar_target(
+    assignment_matches,
+    list(
+      `spiked-urine_FIE-HRMS` = spiked_urine_KEGG_matches,
+      `brachy_FIE-HRMS` = brachy_FIE_HRMS_KEGG_matches,
+      `brachy_C18-UHPLC-HRMS` = brachy_RP_UHPLC_HRMS_KEGG_matches
+    ) %>% 
+      map(~tibble(
+        KEGG_MF_matches = .x$MF %>% 
+          unique() %>% 
+          length(),
+        KEGG_compound_matches = nrow(.x)
+      )) %>% 
+      bind_rows(.id = 'matrix') %>% 
+      mutate(
+        matched_standards = c(spiked_urine_correct_assignments$MF %>% 
+                                unique() %>% 
+                                length(),
+                              rep(NA,2))
+      ) %>% 
+      tidyr::separate(matrix,
+                      c('matrix','technique'),sep = '_')
+  )
+)

### 2023-08-03 R/targets/benchmarking_targets.R

@@ -1,0 +1,67 @@
+n_correlations <- seq(3,5,1) %>% 
+  {10^. * 2.5}
+n_cores <- seq(32,96,32)
+
+benchmark_intervals <- tidyr::expand_grid(
+  n_correlations = n_correlations,
+  n_cores = n_cores
+)
+
+interval <- 0.5
+n_reps <- 1
+
+benchmarking_branches <- tarchetypes::tar_map(
+  values = benchmark_intervals,
+  targets::tar_target(
+    processing_time,
+    procBenchmark(
+      spiked_urine_results_pre_treatment,
+      n_correlations,
+      n_cores,
+      reps = n_reps
+    )
+  )
+)
+
+benchmarking_targets <- list(
+  benchmarking_branches,
+  
+  tarchetypes::tar_combine(
+    processing_time_results,
+    benchmarking_branches,
+    command = dplyr::bind_rows(!!!.x,.id = 'combination')
+  ),
+  
+  targets::tar_target(
+    processing_time_plot,
+    processing_time_results %>% 
+      mutate(`Time (minutes)` = `time (seconds)` / 60,
+             workers = factor(workers),
+             `# correlations` = as.integer(`# correlations`)
+      ) %>% 
+      ggplot(aes(
+        x = `# correlations`,
+        y = `Time (minutes)`,
+        fill = workers
+      )) +
+      geom_line(linetype = 2) +
+      geom_point(shape = 21,
+                 size = 3) +
+      scale_x_log10(labels = scales::comma,breaks = n_correlations) +
+      ggthemes::scale_fill_ptol() +
+      jfmisc::theme_neat() +
+      labs(x = 'No. correlations',
+           fill = 'CPU workers')
+  ),
+  
+  targets::tar_target(
+    cpu_information,
+    minder::cpuInfo()
+  ),
+  
+  targets::tar_target(
+    memory_information,
+    minder::memoryInfo() 
+  )
+)
+

### 2023-08-03 R/targets/brachy_targets.R

@@ -1,0 +1,49 @@
+
+techniques <- c(
+  'FIE-HRMS',
+  'RP-UHPLC-HRMS'
+) 
+
+brachy_targets <- purrr::map(
+  techniques,
+  ~{
+    target_name <- .x %>% 
+      stringr::str_replace_all('-','_') %>% 
+      paste0('brachy_',.) %>% 
+      rlang::sym()
+    
+    if (stringr::str_detect(.x,'LC')){
+      pre_treatment_parameters <- metabolyseR::analysisParameters('pre-treatment')
+      metabolyseR::changeParameter(pre_treatment_parameters,'RSDthresh') <- 20
+    } else {
+      pre_treatment_parameters <- NULL
+    }
+    
+    list( 
+      hrmtargets::tar_input_file_path(!!target_name,
+                                      mzML_files = metaboData::filePaths(.x,
+                                                                         'BdistachyonEcotypes',
+                                                                         ask = FALSE),
+                                      sample_info = metaboData::runinfo(.x,
+                                                                        'BdistachyonEcotypes',
+                                                                        ask = FALSE)),
+      hrmtargets::tar_spectral_processing(!!target_name,
+                                          type = .x %>%
+                                            stringr::str_remove('UHP'),
+                                          export_path = NULL),
+      hrmtargets::tar_pre_treatment(!!target_name,
+                                    parameters = pre_treatment_parameters,
+                                    export_path = NULL),
+      hrmtargets::tar_mf_assignment(
+        !!target_name,
+        parameters = assignments::assignmentParameters(.x %>%
+                                                      stringr::str_remove('UHP')),
+        export_path = NULL
+      ),
+      
+      ## Match assignments to B. distachyon specific KEGG compounds
+      tar_KEGG_matches(!!target_name,
+                       KEGG_IRs_brachy)
+    )
+  }
+)

### 2023-08-03 R/targets/manuscript_targets.R

@@ -1,0 +1,160 @@
+
+manuscript_targets <- list(
+  ## Approach plot
+  tar_target(
+    approach,
+    assignmentApproach()
+  ),
+  
+  ## Supplementary materials and methods
+  tarchetypes::tar_render(
+    supplementary_methods,
+    'manuscript/supplementary_methods.Rmd',
+    output_file = 'Additional file 2.pdf',
+    output_dir = 'exports/manuscript/supplementary_materials'
+  ),
+  
+  ## Supplementary tables
+  tarchetypes::tar_render(
+    supplementary_tables,
+    'manuscript/supplementary_tables.Rmd',
+    output_file = 'Additional file 3.pdf',
+    output_dir = 'exports/manuscript/supplementary_materials'
+  ),
+  
+  ## Supplementary table S1 of spiked compounds
+  tarchetypes::tar_file(
+    table_s1,
+    spiked_urine_compound_info %>% 
+      dplyr::mutate(MF = SMILES %>% 
+                      cheminf::chemicalDescriptors() %>% 
+                      .$MF) %>% 
+      dplyr::select(Name,InChI,`Pubchem CID` = CID,MF) %>% 
+      jfmisc::exportCSV(
+        file = 'exports/manuscript/supplementary_materials/Additional file 1.csv')
+  ),
+  
+  ## Supplementary table S5 of all assignments
+  tarchetypes::tar_file(
+    table_s6,
+    bind_rows(
+      list(
+        `Human_urine_FIE-HRMS` = spiked_urine_results_molecular_formula_assignment %>% 
+          assignments::assignments(),
+        `Brachypodium distachyon_leaf_FIE-HRMS` = brachy_FIE_HRMS_results_molecular_formula_assignment %>% 
+          assignments::assignments(),
+        `Brachypodium distachyon_leaf_C18-UHPLC-HRMS` = brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment %>% 
+          assignments::assignments()
+      ),
+      .id = 'Type'
+    )  %>%
+      tidyr::separate(
+        Type,
+        c('Organism',
+          'Matrix',
+          'Technique'),
+        sep = '_'
+      ) %>% 
+      left_join(
+        bind_rows(
+          list(
+            `Human_urine_FIE-HRMS` = spiked_urine_KEGG_matches,
+            `Brachypodium distachyon_leaf_FIE-HRMS` = brachy_FIE_HRMS_KEGG_matches,
[…]

### 2023-08-03 R/targets/spiked_urine_targets.R

@@ -1,0 +1,545 @@
+
+spiked_urine_search_ppm <- 6
+
+spiked_urine_assignment_parameters <- {
+  ap <- assignments::assignmentParameters('FIE')
+  assignments::adducts(ap)$n <- c("[M-H]1-",
+                                  "[M+Cl]1-",
+                                  "[M+K-2H]1-",
+                                  "[M+Cl37]1-",
+                                  "[M-2H]2-",
+                                  "[2M-H]1-")
+  assignments::ppm(ap) <- spiked_urine_search_ppm
+  assignments::adducts(ap)$p <- assignments::adducts(ap)$p[!stringr::str_detect(assignments::adducts(ap)$p,
+                                                                                stringr::coll("[M+NH4]1+"))]
+  assignments::isotopes(ap) <- c('13C','18O','34S','13C2')
+  ap
+}
+
+spiked_urine_example_feature <- 'n80.97488' 
+
+spiked_urine_targets <- list(
+  ## spiked_urine compound info file
+  tarchetypes::tar_file(
+    spiked_urine_compound_info_file,
+    {
+      'spiked_chemical_standards_info.csv' %>% 
+        piggyback::pb_download(dest = 'data',
+                               tag = 'spiked-urine')
+      
+      'data/spiked_chemical_standards_info.csv' 
+    }
+  ),
+  
+  ## load spiked_urine compound info
+  tar_target(
+    spiked_urine_compound_info,
+    readr::read_csv(spiked_urine_compound_info_file) %>% 
+      dplyr::rowwise() %>% 
+      dplyr::group_split() %>% 
+      purrr::map_dfr(~ .x %>% 
+                       dplyr::mutate(SMILES = cheminf::convert(
+                         InChI,
+                         'INCHI',
+                         'SMILES'),
+                         MF = cheminf::smilesToMF(SMILES)))
+  ),
+  
+  ## calculate spiked compound descriptors
+  tar_target(
+    spiked_urine_compound_db,
+    cheminf::metaboliteDB(
+      spiked_urine_compound_info %>% 
+        dplyr::rename(ID = CID,
+                      NAME = Name)
+    )
+  ),
+  
+  ## calculate possible spiked compound adducts
+  tar_target(
+    spiked_urine_compound_adducts,
+    spiked_urine_compound_db %>%
+      cheminf::entries() %>%
+      .$ID %>%
+      purrr::map(cheminf::calcAdducts,
[…]

### 2023-08-03 R/utils.R

@@ -1,0 +1,18 @@
+## Resolve conflicts
+# conflict_prefer(quiet = TRUE)
+
+# options(future.debug = TRUE)
+
+## Set targets options
+targets::tar_option_set(
+  error = 'continue',
+  memory = 'transient',
+  garbage_collection = TRUE
+)
+
+## Parallel backend
+metaboMisc::suitableParallelPlan(
+  workers = jfmisc::suitableParallelWorkers(
+    RAM_per_worker = 32
+  )
+)

### 2023-08-03 _targets.R

@@ -1,0 +1,15 @@
+## Generated by projecttemplates (https://jasenfinch.github.io/projecttemplates/) v0.5.0
+
+"R/targets/" %>%
+  list.files(full.names = TRUE) %>%
+  purrr::walk(source)  
+
+list(
+  KEGG_compound_targets,
+  MF_selection_targets,
+  spiked_urine_targets,
+  brachy_targets,
+  assignment_results_targets,
+  benchmarking_targets,
+  manuscript_targets
+)

### 2023-08-03 compile.sh

@@ -1,0 +1,3 @@
+logfile=$(date +"misc/logs/%Y-%m-%d_%H-%M-%S_compile.log")
+
+/usr/lib/R/bin/Rscript 'misc/run.R'  >> $logfile 2>&1

### 2023-08-03 misc/docker/build.sh

@@ -1,0 +1,1 @@
+docker build . --file misc/docker/Dockerfile -t data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_esi-ms

### 2023-08-03 misc/docker/run.sh

@@ -1,0 +1,4 @@
+docker run \
+  --entrypoint /usr/bin/bash \
+  -v $(pwd):/home/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS \
+  data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_esi-ms

### 2023-08-03 misc/run.R

@@ -1,0 +1,7 @@
+## Generated by projecttemplates (https://jasenfinch.github.io/projecttemplates/) v0.5.0
+## Execute this script to run the project analysis
+targets::tar_make()
+
+pipeline_graph <- targets::tar_visnetwork(label = c("time", "size"))
+visNetwork::visSave(pipeline_graph,
+                    file = "exports/pipeline_graph.html")

### 2023-08-07 manuscript/additional_files.Rmd
Move additional file information to a separate file

--- 2023-08-03 manuscript/manuscript.Rmd
@@
----
-output: 
-  bookdown::pdf_document2:
-    toc: false
-    keep_tex: true
-fontsize: 12pt
-header-includes:
-  - \linespread{2}
-  - \usepackage{lineno}
-  - \linenumbers
-  - \usepackage{booktabs, multirow, tabularx, float}
-bibliography: references.bib
-csl: vancouver.csl
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,
-                      warning = FALSE,
-                      message = FALSE,
-                      fig.align = 'center',
-                      fig.pos = 'H')
-```
-
-# **Title:** Data-driven, matrix-wide molecular formula assignment for ultra-high resolution ESI-MS {-}
-
-**Authors:**  Jasen P. Finch^1\*^, Thomas Wilson^1^, Laura Lyons^1^, Helen Phillips^1^, Manfred Beckmann^1^ & John Draper^1^
-
-^\*^ Corresponding author
-
-**Addresses:**
-
-^1^ Department of Life Sciences, Aberystwyth University, Aberystwyth, SY23 3DA, UK
-
-# Abstract {-}
-
-## Background {-}
-
-Compound identification remains one of the primary bottlenecks in the analysis of non-targeted mass spectrometry metabolomics data.
-Some annotation approaches are available; however, these are primarily for liquid chromatography-mass spectrometry (LC-MS) profiling and are unsuitable for non-targeted metabolome fingerprinting techniques.
-
-## Results {-}
-
-```{r assignment-summary}
-tar_load(assignment_results)
-tar_load(assignment_matches)
-tar_load(spike_urine_standards_MF_assignment_possible)
-tar_load(spiked_urine_correct_assignments)
-
-total_assigned <- round(sum(assignment_results$n_assignments) / 
-                          sum(assignment_results$n_features) * 100,1)
-total_mfs <- sum(assignment_results$n_MFs)
-total_kegg <- sum(assignment_matches$KEGG_compound_matches)
-
-spiked_urine_possible <- nrow(spike_urine_standards_MF_assignment_possible)
-spiked_urine_correct_percent <- spiked_urine_correct_assignments$MF %>% 
-  unique() %>% 
-  length() %>% 
-  {. / spiked_urine_possible * 100} %>% 
-  round(digits = 1)
-```
-
-We present a data-driven ensemble approach for the ‘first pass’ automated assignment of molecular formulas to mass spectra acquired from high-resolution electrospray ionisation mass spectrometry (ESI-MS).
-This is primarily intended for metabolomics experiment data using non-targeted metabolome fingerprinting and LC-MS profiling.
-The approach incorporates the use of *m/z* correlation networks, molecular formula generation and selection as well as graphically based component plausibility selection to assign molecular formulas to *m/z* features.
-The R package *assignments* was developed as a user-friendly implementation of the approach and is freely available from https://github.com/aberHRML/assignments.
-The approach was tested in human urine samples spiked with known chemical standards and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
-Across the three test matrices, `r total_assigned`% of the m/z features were assigned a molecular formula giving a total of `r total_mfs` molecular formulas and matching to `r total_kegg` compounds in the KEGG compound database.
-Of the `r spiked_urine_possible` spiked chemical standards that it was possible to assign in the human urine samples, the approach was able to assign m/z features to `r spiked_urine_correct_percent`% of these.
-
-## Conclusions {-}
-
-The approach presented here and its implementation in the R package *assignments* provides investigators with an efficient and automated approach for matrix-wide ‘first pass’ assignment of molecular formulas to non-targeted ESI-MS metabolomics experiment data.
-
-# Introduction
-
-Metabolite identification remains one of the main bottlenecks both in time and resources in the analysis of non-targeted mass spectrometry based metabolomics approaches.
-Electrospray ionisation mass spectrometry (ESI-MS) is preferred as this is a soft ionisation technique that enables easier compound identification [@domingo2018].
-Irrespectively, it is possible for a single metabolite to be represented within ESI-MS spectra by multiple ionisation product adducts and isotopes [@draper2013].
-
-Ultra-high resolution MS instrumentation utilising Orbitrap mass analyser technology is now commonplace in many laboratories, which can yield many thousands of *m/z* features in a single analytical run [@hu2005].
-This has provided an unprecedented opportunity for metabolite annotation and has been applied to approaches such as 'first pass' flow infusion electrospray high-resolution MS (FIE-HRMS) fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
-However, mass accuracy alone does not meet the requirements for metabolite annotation [@kind2006].
-
-Metabolite annotation is a crucial aspect for ensuring that biological interpretation is possible from metabolomics data sets.
-A number of strategies can be employed to annotate peaks in mass spectra.
-These can include simple approaches such as heuristic filtering of molecular formulas generated from *m/z* or direct matching putative ionisation products to metabolite databases [@draper2009; @green2015; @kind2007];
-but also ensemble approaches that integrate a number of strategies like correlation analysis, ionisation product scoring and biotransformations [@baran2013; @fernandez2014; @kuhl2012; @uppal2017].
-Tandem MS can also be employed for LC-MS approaches and provides a convenient method for structural validation [@neumann2010].
-However, this is unsuitable for metabolome fingerprinting-based techniques due to the presence of structural isomers and isobaric compounds.
-Many of the available software packages implementing these strategies are suitable for LC-MS techniques only.
-
-For investigators, manual consideration of putative annotations is time-consuming and often difficult to interpret due to the complex interactions between ions and molecules and their mathematical possibilities forming *m/z* features.
-It is important that software package implementations can automate this process but also provide tractable results and simple input parameters for easy and wide applicability.
-
-Correlation analysis has been commonly used to identify related metabolites within metabolomics data and is often used in LC-MS approaches to group related *m/z* features [@steuer2006;@kuhl2012]. 
-These analyses can be represented as weighted undirected network graphs where the *m/z* features reflect the graph nodes and the significant correlations between them, the graph edges [@gehlenborg2010].
-Applying Spearman's rank correlation allows the detection of both linear and monotopic associations between *m/z* features.
-Adduct and isotopic relationships would be expected to have positive linear correlations whereas compounds associated by pathway transformations within living tissues under the control of gene and protein regulation would be able to have both positive or negative monotonic correlations [@camacho2005].
-
-The use of correlations to associate relationships between *m/z* features provides an essential layer of data-driven, internal validation of the mathematically derived solutions.
-For instance, if two features that are posited to have the same composition (eg.
-a C^13^ relationship) but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about associations (eg.
-metabolic pathway changes) of those features with other putatively assigned features within the matrix.
-
-Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS fingerprinting and LC-MS profiling-based metabolomics data.
-This is aimed at biological experiments that include numerous replicates of sample classes for the assignment of low accurate mass (< 1000 atomic mass units (amu)) metabolite elemental compositions of biological origin.
-It is an ensemble approach that combines a number of annotation strategies including correlation analysis, molecular formula selection, adduct and isotopic scoring and graphical component selection.
-This is accompanied by the development of the R package *assignments* as an implementation of the approach.
-
-# Materials and methods
-
-## Biological sample preparation and ESI-MS analysis
-
-```{r chemical-standards}
-tar_load(spiked_urine_compound_info)
-
-spiked_urine_compound_MFs <- spiked_urine_compound_info %>% 
-  select(MF) %>% 
-  distinct() %>% 
-  nrow()
-```
-
-Two example sample matrices were used for assessing the molecular formula assignment approach. 
-The first was human urine spiked with a diverse mixture of `r nrow(spiked_urine_compound_info)` chemical standards from a previously validated assay that included `r spiked_urine_compound_MFs` unique molecular formulas [@beckmann2020].
-A full list of the names and InChI chemical identifiers of the spiked chemical standards can be found in Additional file 1.
-The second was healthy leaf tissue samples collected from four ecotypes of the model grass species *Brachypodium distachyon* using the method described in @parker2008.
-Further details of the sample collection and extraction are described in the Additional file 2.
-
-The sample matrices were each analysed using FIE-HRMS fingerprinting and C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling.
-These data are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949.
-The FIE-HRMS fingerprinting data were spectrally processed by performing spectral binning using the R package *binneR* v`r packageVersion('binneR')` [@finch2022].
-Spectral processing of the C18-UHPLC-HRMS profiling data was conducted using the R package *xcms* v`r packageVersion('xcms')` [@smith2006].
-The data were then pre-treated by performing batch correction, feature occupancy filtering, imputation, relative standard deviation filtering based on QC samples and total ion count normalisation using the R package *metabolyseR* v`r packageVersion('metabolyseR')` available from GitHub (https://github.com/jasenfinch/metabolyseR).
-Further details of the ESI-MS analyses and data preparation are available in Additional file 2.
-All data processing and analysis were performed using the R Statistical Programming Language version `r stringr::str_c(version$major,version$minor,sep ='.')` [@Rteam2022].
-
-## Molecular formula assignment
-
-For each biological sample matrix, the input data for molecular formula assignment had undergone spectral processing and pre-treatment and consisted of a sample by *m/z* feature intensity matrix containing both positive and negative ionisation mode data.
-Molecular formula assignment was conducted using the R package *assignments* v`r packageVersion('assignments')` and the approach comprised four major steps which are outlined in Figure \@ref(fig:figure-1-approach).
-Where assignment was possible, *m/z* features were designated with unambiguous molecular formulas, isotopes and ionisation adduct assignments. 
-The following sections outline the methodology behind each major step of the approach.
-
-\linespread{1}
-
-```{r figure-1-approach,fig.height=6,fig.cap='An overview of the approach for assigning molecular formulas to \\textit{m/z} features from ultra-high resolution ESI-MS metabolomics experiments.'}
-tar_read(approach)
-```
-
-\linespread{2}
-
-### Correlation analysis
-
-Spearman's rank correlation was conducted between all pairwise combinations of *m/z* features.
-A Bonferroni correction was applied to the p-value of each coefficient based on the total number of features.
-Correlations with a p-value below 0.05 and a minimum absolute coefficient of 0.7 were retained.
-
-### Relationship calculation {#relationships}
-
-```{r adducts}
-tar_load(brachy_FIE_HRMS_parameters_molecular_formula_assignment)
-
-negative_mode_adducts <-  brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
-  assignments::adducts() %>% 
-  .$n %>% 
-  stringr::str_replace_all(']1-',']^1-^') %>% 
-  stringr::str_replace_all(']2-',']^2-^') %>% 
-  stringr::str_replace_all('Cl37','^37^Cl') %>% 
-  str_as_written_list()
-positive_mode_adducts <-   brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
-  assignments::adducts() %>% 
-  .$p %>% 
-  stringr::str_replace_all(stringr::coll(']1+'),']^1+^') %>% 
-  stringr::str_replace_all(stringr::coll(']2+'),']^2+^') %>% 
-  stringr::str_replace_all('K41','^41^K') %>% 
-  str_as_written_list()
-```
-
-Possible adduct, isotopic and transformation mass relationships were calculated for each pair of correlated *m/z*.
-The pre-ionisation masses ($M$) were first calculated for all adduct, isotopic and transformation combinations for each of the *m/z*. 
-An $M$ for a given ionisation product *m/z*, adduct, isotope and transformation was calculated using the following equation:
-
-$$M = \frac{c(m/z - \Delta m_a) - \Delta m_i}{n} - \Delta m_t$$
-
-Where $c$ is the overall ionisation product charge, $\Delta m_a$ is the adduct mass change, $\Delta m_i$ is the isotopic mass change, $n$ is the number of molecules involved in the ionisation product and $\Delta m_t$ is the transformation mass change.
-
-The negative ionisation mode adducts included: `r negative_mode_adducts`.
-The positive ionisation mode adducts included: `r positive_mode_adducts`.
-All adduct mass, charge and molecule information can be found in Additional file 3 (Table S2).
-The isotopes included single and double ^13^C and ^18^O for the *B.distachyon* matrices and additionally ^34^S for the spiked urine.
-The isotopic mass information used can be found in Additional file 3 (Table S3).
-For all matrices, `r nrow(mzAnnotation::transformation_rules)` transformations were based on the transitions of common metabolic reactions found in primary metabolism.
-These transformations can be found in Additional file 3 (Table S4).
-
-To calculate the mass error for possible relationships between the *m/z* in the correlated pair, the mass differences between all the possible $M$ values of the *m/z* were computed.
-All relationships with an $M$ difference in mass error below 0.001 amu and a maximum of one transformation were retained.
-
-### Adduct and isotopic assignment iterations
-
-This first assignment step used only positively correlated feature relationships that corresponded to adduct and isotopic relationships without transformation. 
-This assignment step consisted of molecular formula generation, molecular formula selection (Section \@ref(MF-selection)) and then multiple iterations of network graph component assembly and network graph component selection (Section \@ref(component-selection)).
-
-### Molecular formula generation and selection {#MF-selection}
-
-For all the relevant *m/z* relationships, molecular formulas were generated from all the possible $M$ values based on these relationships.
-To reduce the computational requirements for molecular formula generation with increasing molecular mass, the possible $M$ values were limited to a maximum of 800 amu.
-
-Generated molecular formulas were restricted to only include the elements C, H, N, O, P and S.
-For a given $M$, the maximum allowed element frequencies for generated molecular formulas were restricted using the following terms: $C = M/12$, $H = 2C$, $N,O = C/2$, $P,S = C/4$.
-The calculated frequencies were rounded to the nearest integer value.
-Molecular formula generation was conducted for each $M$ value within a mass deviation range of $\pm$ 6 ppm.
-This generation was exhaustive, using the brute force approach of the HR2 generator, without the application of the golden rules proposed by @kind2007.
-
-To allow the most plausible molecular formulas to be selected, a percentage plausibility score was calculated based on 12 heuristic checks of the rules 2, 4, 5 and 6 from @kind2007.
-These included the LEWIS and SENIOR checks, five element ratio checks and five element probability checks.
-For each molecular formula, a point was awarded for each check that passed or was not applicable.
-These points were then totalled and divided by 12 to give a golden rules check score ($S_{GR}$).
-The proportion of the atoms in the molecule that were C, H or O was then calculated to give a CHO proportion score ($S_{CHO}$).
-A percentage plausibility score ($P_{MF}$) was then calculated for the molecular formula using the equation below.
-
-$$P_{MF} = \frac{S_{GR} + S_{CHO}}{2} \times 100  $$
-The three top-ranked molecular formulas with the highest $P_{MF}$ scores, generated for a given $M$, were selected for network component assembly.
-
-### Graphical component selection {#component-selection}
-
-For each selected molecular formula, a component subgraph was constructed from the correlation network graph using all matching features, with features allowed across multiple components.
-All components with a size of less than one and containing only isotopic features or adducts were removed.
-
-The components were ranked from most plausible to least plausible using a component plausibility score $P_c$. 
-This was based on the component size, the average component edge weight and an additional score based on the plausibility of the adduct and isotopic combinations of the features within the component (AIS).
-
-To calculate $AIS$, the possible adducts within each ionisation mode were ranked from most probable to least probable and assigned a score ($a$) with the most probable being $n-1$, $n$ being the number of adducts within that ionisation mode and least probable adduct a score of 0.
-The adduct rank scores used for the assignment were based on the adduct order specified in Section \@ref(relationships) for all matrices.
-
-Similarly, for the possible isotopes, these were also ranked from most (non-isotopic) to least probable and assigned a score ($i$) as for the adducts.
-The isotopic rank scores used for the assignment of the *B.distachyon* matrices were as follows: Non-isotopic = 3, ^13^C = 2, ^18^O = 1, ^13^C2 = 0.
-For the spiked urine matrices, the isotopic rank scores were: Non-isotopic = 4, ^13^C = 3, ^18^O = 2, ^34^S = 1, ^13^C2 = 0.
-
-The $AIS$ for a given adduct and isotopic combination ($k$) was then calculated using:
-
-$$AIS_k = \frac{a_k + i_k}{\max{a} + \max{i}}$$
-
-The $AIS$ scores within each component were totalled and divided by half the total number of adduct and isotope combinations to give a component $AIS$ score ($AIS_c$).
-
-For each component, the component plausibility score $P_c$ was calculated using:
-
-$$P_c = d \times AIS_c \times w$$
-Where $d$ is the component degree and $w$ is the average component edge weight. 
-The higher $P_c$, the more plausible the component subgraph.
-
-Features shared between multiple components were sequentially eliminated and retained for only the components with the highest $P_c$.
-For any ties in $P_c$ between components, the component with the highest $P_{MF}$ was selected.
-Further ties were decided by selecting the component with the lowest ppm error.
-This graphical component selection routine was repeatedly performed across multiple iterations until no more assignments were obtained.
-
-### Transformation assignment iterations
-
-The transformation assignment iterations consisted of multiple assignment iterations where only positively or negatively correlated relationships containing a previously assigned feature and a transformation were included.
-Each iteration was performed as for the previous adduct and isotopic assignment iteration and repeated until no more assignments were obtained.
-
-## Assignment assessments
-
-```{r KEGG-compounds}
-tar_load(KEGG_compounds_valid_MFs)
-```
-
-Features matching to possible ionisation products of the spiked urine compounds were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
-The calculated *m/z* were then matched to the spiked urine features using a search range of $\pm$ `r spiked_urine_search_ppm` ppm.
-
-The molecular formulas and monoisotopic masses for compound entries in the KEGG compound database were accessed using the R package *KEGGREST* v`r packageVersion('KEGGREST')` [@kanehisa2000;@tenenbaum2021].
-For the assessments of molecular formula selection success, only uncharged molecular formulas containing the elements C, H, N, O, P, or S and a monoisotopic mass below 1000 amu were retained.
-This gave a total of `r nrow(KEGG_compounds_valid_MFs)` unique molecular formulas.
-
-For the assignments in *B.distachyon* and human urine matrices, KEGG compounds specific to each organism were identified by cross-referencing KEGG compound IDs with the KEGG enzyme database for *B.
-distachyon*.
-For these compounds, InChI chemical identifiers were obtained by cross-referencing their entries in the PubChem Substance database with the PubChem Compound database [@kim2016].
-To match possible KEGG compounds specific to each organism to the molecular formulas assigned to *m/z* features, the organism-specific compound molecular formulas were first filtered to compounds only with molecular formulas that were assigned to *m/z* features.
-By applying the MZedDB ionisation rules, only KEGG compounds were retained for which the adducts of the assigned features were possible.
-
-# Results and discussion
-
-## *assignments*: An R package for molecular formula assignment of ESI-MS metabolomics data
-
-The R package *assignments* was developed as an implementation of this molecular formula assignment approach.
-It is available to install from GitHub at https://aberhrml.github.io/assignments/.
-The package is fully documented and a usage guide is provided for new users.
-
-The input data consists of a sample by accurate *m/z* feature intensity matrix with at least four samples.
-This data should be spectrally processed and appropriately pre-treated prior to assignment.
-All assignment parameters are fully customisable, which includes the specification of custom adducts, isotopes and transformations.
-It is possible for all the assignment criteria and results to be recovered for each assignment step, enabling the manual curation of assignments by the user.
-The default parameters provided by the package have been designated to enable assignment for a wide range of biological applications.
-
-Parallel processing is provided by the *future* R package, which allows the user to specify a number of different parallel backends [@bengtsson2021].
-Benchmarking of the processing time of assignments made to the spiked urine sample matrix was performed by the *assignments* R package using varying numbers of CPU workers (Additional file 2).
-With the use of high-performance computing resources with greater than 32 CPU workers resulted in assignments being obtained in under two hours for 250,000 correlations between *m/z* features (Additional file 3 (Table S5)).
-
-## Molecular formula assignment success {#success}
-
-```{r assignment-success}
-tar_load(spiked_urine_compound_PIPs)
-tar_load(spiked_urine_feature_matches)
-tar_load(spiked_urine_unassigned)
-tar_load(spiked_urine_assigned)
-tar_load(spiked_urine_incorrect_assignments)
-
-brachy_shared_MFs <- tar_read(brachy_FIE_HRMS_results_molecular_formula_assignment) %>% 
-  assignments::assignments() %>% 
-  select(MF) %>% 
-  inner_join(tar_read(brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment) %>% 
-               assignments::assignments() %>% 
-               select(MF),
-             by = 'MF') %>% 
-  distinct() %>% 
-  nrow()
-
-brachy_shared_kegg <- intersect(
-  tar_read(brachy_FIE_HRMS_KEGG_matches)$MF %>% 
-    unique(),
-  tar_read(brachy_RP_UHPLC_HRMS_KEGG_matches)$MF %>% 
-    unique()
-) %>% length()
-
-positive_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
- assignments::adducts() %>% 
-  .$p %>% 
-  length()
-
-negative_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
- assignments::adducts() %>% 
-  .$n %>% 
-  length()
-
-spiked_urine_MF_PIPs <- spiked_urine_compound_PIPs %>% 
-  select(MF,Adduct,Isotope) %>% 
-  distinct()
-
-spiked_urine_feature_matches_MF_n <- spiked_urine_feature_matches$MF %>% 
-  unique() %>% length()
-
-percent_unassigned <- {nrow(spiked_urine_unassigned) / 
-    nrow(spiked_urine_feature_matches) * 100} %>% 
-  signif(digits = 3)
-
-spiked_urine_unassigned_MF_n <- spiked_urine_unassigned %>% 
-  anti_join(
-    spiked_urine_assigned %>% 
-      select(MF) %>% 
-      distinct(), 
-    by = 'MF') %>% 
-  .$MF %>% 
-  unique() %>% 
-  length()
-
-spiked_urine_correct_assignments_MF_n <- spiked_urine_correct_assignments %>%
-  .$MF %>% 
-  unique() %>% 
-  length()
-
-spike_urine_standards_MF_assignment_impossible_n <- spiked_urine_feature_matches_MF_n -
-  nrow(spike_urine_standards_MF_assignment_possible)
-
-percent_standards_MF_assigned <- {spiked_urine_correct_assignments_MF_n / 
-  nrow(spike_urine_standards_MF_assignment_possible) * 100} %>% 
-  signif(digits = 3)
-```
-
-An overview of the molecular formula assignments made in both the urine and *B.
-distachyon* FIE-HRMS matrices is provided in Table \@ref(tab:assignment-results), which shows that similar proportions of *m/z* features were assigned molecular formulas.
-However, there was more than double the relative molecular formula assignment coverage for the C18-UHPLC-HRMS data of *B. distachyon* compared to the FIE-HRMS matrices.
-The chromatographic separation of the C18-UHPLC-HRMS increases the likelihood that a number of correlated ionisation products will be present for a given compound due to a reduction in the effects of ion suppression [@gowda2014].
-There is also a decreased chance of interference in the correlations between related features due to the presence of structural isomers and isobaric compounds.
-A total of `r brachy_shared_MFs` molecular formulas were assigned in both the FIE-HRMS and C18-UHPLC-HRMS data of *B. distachyon*.
-
-\linespread{1}
-
-```{r assignment-results}
-tar_read(assignment_results) %>%
-  left_join(tar_read(assignment_matches), 
-            by = c("matrix", "technique")) %>% 
-  mutate(matrix = paste(matrix,technique)#,
-  ) %>%
-  mutate(n_assignments = glue::glue(
-    '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
-    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
-  select(-technique) %>% 
-  dplyr::rename(`\\textit{m/z} features` = n_features,
-                correlations = n_correlations,
-                relationships = n_relationships,
-                `assigned features` = n_assignments,
-                `molecular formulas (MF)` = n_MFs,
-                `KEGG compound matches` = KEGG_compound_matches,
-                `KEGG MF matches` = KEGG_MF_matches,
-                `matched chemical standards` = matched_standards) %>% 
-  tidyr::gather(Variable,Value,-matrix) %>% 
-  mutate(Value = replace(Value,
-                         Value == 'NA',
-                       'N/A')) %>% 
-  tidyr::spread(matrix,Value) %>% 
-  slice(c(1,3,8,2,7,5,4,6)) %>%
-  set_colnames(stringr::str_split_fixed(colnames(.),
-                                        ' ',2)[,2]) %>% 
-  knitr::kable(format = 'latex',
-               booktabs = TRUE,
-               linesep = '',
-               escape = FALSE,
-               align = c(
-                 rep('l',1),
-                 rep('r',3)
-               ),
-               caption = 'A summary of the molecular formula assignment results for the example sample matrices.
-The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
-  kable_styling(latex_options = "HOLD_position") %>% 
-  add_header_above(c(
-    " ", 
-    "\\\\textit{B. distachyon} leaf" = 2, 
-    "Human urine" = 1),
-    escape = FALSE)
-```
-
-\linespread{2}
-
-Although there were proportionately more molecular formulas assigned to *m/z* features in the C18-UHPLC-HRMS compared to the FIE-HRMS *B.
-distachyon* matrix, there was a disproportionate number of these molecular formulas that matched to compounds in the KEGG database specific to *B.
-distachyon* and able to form the relevant ionisation products (Table \@ref(tab:assignment-results)).
-However, only `r brachy_shared_kegg` of the molecular formulas matched to compounds in the KEGG database were found to be common between the two techniques.
-The low number of assigned molecular formulas partly reflects the incompleteness of the KEGG database, especially with respect to the secondary metabolites that are prolific in C18-UHPLC-HRMS [@breitling2013].
-The assignments for all the matrices along with the matched KEGG compound IDs are provided in Additional file 4.
-
-To assess the assignments in the example urine matrix with respect to the spiked chemical standards, the predicted ionisation products were first identified by directly matching these to the observed *m/z*.
-Based on the `r positive_mode_adducts + negative_mode_adducts` adducts and four isotopic rules specified, there were `r nrow(spiked_urine_MF_PIPs)` predicted ionisation product *m/z* for these chemical standards.
-A total of `r nrow(spiked_urine_feature_matches)` of these ionisation products could be directly matched to the observed *m/z* features which included `r spiked_urine_feature_matches_MF_n` of the `r spiked_urine_compound_MFs` unique molecular formulas of the spiked chemical standards.
-These directly matched *m/z* features are provided in Additional file 5.
-For `r nrow(spiked_urine_unassigned)` (`r percent_unassigned`%) of these matched *m/z* features, no molecular formula was assigned by the approach, which included `r spiked_urine_unassigned_MF_n` of the molecular formulas of the spiked chemical standards.
-However, `r nrow(spiked_urine_assigned)` of the matched features were assigned a molecular formula with `r nrow(spiked_urine_correct_assignments)` of these corresponding to one of the spiked compound ionisation products.
-This included `r spiked_urine_correct_assignments_MF_n` of the spiked chemical standards molecular formulas (Table \@ref(tab:assignment-results)).
-The *m/z* features assigned to molecular formulas corresponding to the spiked chemical standards are provided in Additional file 6.
-
-The remaining `r nrow(spiked_urine_incorrect_assignments)` matched *m/z* features were assigned alternative molecular formula ionisation products.
-The *m/z* of the predicted ionisation features that it was not possible to assign molecular formulas were as a result of either not being correlated with other *m/z* features that corresponded to a correct adduct or isotope mass difference or whose subgraph components were not selected, with more plausible solutions instead selected based on the correlation network.
-
-Accounting for the `r spike_urine_standards_MF_assignment_impossible_n` chemical standards that were not assigned due to a lack of relevant correlated mass relationships, the algorithm was able to assign molecular formulas to features corresponding to `r percent_standards_MF_assigned`% of the `r nrow(spike_urine_standards_MF_assignment_possible)` possible chemical standards.
-This highlights the difficulty in annotating these complex biological mixtures containing hundreds of compounds, which can create many thousands of possible ionisation products.
-Interference in correlations due to the presence of isobaric compounds and MS instrument noise can weaken correlations between features, particularly in those with low abundance.
-
-## The molecular formula assignment approach
-```{r sample-size}
-sample_size_n_features <- 4000
-sample_size <- pwr::pwr.r.test(
-  r = 0.7,
-  power = 0.8,
-  sig.level = 0.05/sample_size_n_features)$n %>% 
-  round()
-```
-
-```{r mf-selection}
-tar_load(KEGG_compound_MF_ranks)
-
-kegg_top_rank_percent <- KEGG_compound_MF_ranks %>% 
-  mutate(top_10 = Rank <= 10) %>% 
-  group_by(top_10) %>% 
-  summarise(N = n()) %>% 
-  mutate(percent = N / sum(N) * 100) %>% 
-  {.$percent[.$top_10 == TRUE]} %>% 
-  signif(digits = 3)
-```
-
-The use of correlation analysis to identify related features requires a suitable number of observations to ensure that statistically significant associations can be detected. 
-This is especially important in non-targeted metabolomics data sets that can contain thousands of *m/z* features and p-value adjustment is required to account for multiple testing. 
-For example, to detect a significant lower limit of Pearson's correlation coefficient of 0.7 for a matrix containing `r sample_size_n_features` *m/z* features, an approximate minimum sample size of `r sample_size` observations would be required (p < 0.05, Bonferroni corrected) [@cohen2013].
-Metabolomics fingerprinting and profiling experiments typically require high replication of 10-15 observations per class in order to discriminate between the sample classes [@enot2008]. 
-Therefore, a sample set comparing just four classes brings this to within a practical range for applying this approach in many cases, such as the example *B.
-distachyon* ecotype comparison sample set provided here.
-The assignments obtained from the *B. distachyon* example matrix also show the application of the assignment approach to sample sets that contain multiple underlying experimental classes.
-
-The approaches of the *MFAssignR* R package and *Formularity* software provide alternative means of assigning molecular formulas to both fingerprinting and LC-MS data, with *Formularity* utilising a compound identification database to aid molecular formula selection [@schum2020;@tolic2017].
-While these approaches provide the advantage of being able to assign molecular formulas using a single replicate with substantially low computational demands, they do not incorporate the use of correlations between *m/z* features.
-This can lead to incorrect assignments and the potential miss assignment of isotopic formulae to uncorrelated features.
-Both the *CAMERA* and *MAIT* R packages utilise correlations for adduct and isotope annotation but only for LC-MS data [@fernandez2014;@kuhl2012].
-The MAIT package also includes the addition of biotransformations to improve peak annotation performance.
-
-Assignment approaches using molecular formula selection strategies have been modifications of the seven golden rules proposed by @kind2007 to identify criteria for chemically correct formulae.
-These include the assignment approach of the R package *MFAssignR* and formula filtering for PubChem database matching [@lommen2014; @schum2020].
-These approaches apply the heuristic rules discretely where molecular formulas are removed if they do not meet the criteria.
-The scoring approach used here for selection builds on this use of discrete rules by enabling generated molecular formulas to be ranked in terms of their plausibility.
-It also allows the inclusion of non-carbon containing molecules such as phosphate and sulphate that are common in biological systems.
-The effectiveness of this ranking is reflected in the fact that, of the `r nrow(KEGG_compound_MF_ranks)` valid KEGG compound molecular formulas, `r kegg_top_rank_percent`% of these formulas could be ranked within the top 10 for plausibility when molecular formulas were generated for the compound accurate masses inside a range of `r KEGG_alternative_MFs_ppm` ppm.
-This is also shown in Figure \@ref(fig:figure-2-kegg-mf-selection) where the plausibility ranking is still effective even for masses above 500 amu where it is possible for many thousands of formulas to be generated.
-
-\linespread{1}
-
-```{r figure-2-kegg-mf-selection,fig.height=4,fig.cap='Molecular formula plausibility ($P_{MF}$) rankings of KEGG compound molecular formulas.
-The $P_{MF}$ rankings were based on the monoisotopic masses of 6745 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 6 ppm.'}
-tar_read(KEGG_compound_MF_ranks_mass_plot) +
-  tar_read(KEGG_compound_MF_ranks_n_plot)
-```
-
-\linespread{2}
-
-As shown in Figure \@ref(fig:figure-3-feature-solutions), there can be multiple graphical molecular formula component solutions can exist for a given *m/z* feature.
-The component plausibility score was the primary criterion for which components were selected. 
-This score combines information about how connected the features are within a given component, the average strength of these connections and the expected commonality of the adduct and isotopic combinations present.
-By using this information about the connections of *m/z* within the spectrum, it is possible for an *m/z* to be assigned a less plausible molecular formula if it is more highly connected with strong connections containing more common adduct and isotopic combinations.
-Combining information between the ionisation modes further increases available information and more informative components [@kuhl2012].
-
-\linespread{1}
-
-```{r figure-3-feature-solutions,fig.height=5,fig.width=8,fig.cap = 'Graphical component solutions for the spiked urine negative mode feature 80.97488 \\textit{m/z}.
-The components were generated during the first adduct and isotopic assignment iteration.
-The selected graphical component is bordered in red.
-The values of attributes are shown below each component.
-The nodes for the negative mode feature 80.97488 \\textit{m/z} are highlighted in blue.',dev='png',dpi=300}
-tar_read(spiked_urine_feature_solutions_plot)
-```
-
-\linespread{2}
-
-Adduct formation will vary depending on the sample origin and the frequency of isotopes will be dependent on the elemental compositions of the constituent compounds [@kruve2017].
-The input rankings of the adducts and isotopes ranks used will alter the preference of selected components so gains can possibly be made in assignment accuracy with the optimisation of these rankings on a per sample type basis.
-
-Although examples used here include FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using non-targeted ultra-high resolution mass accuracy ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
-Similarly, data from other chromatographic techniques for LC-HRMS profiling could be used such as hydrophilic interaction chromatography (HILIC), where suitable adducts and adduct rankings could instead be specified [@spagou2010].
-
-This approach was designed for use with any sample tissues and fluids of biological origin and analysed using non-targeted ultra-high resolution mass accuracy ESI-MS data.
-The additional input of knowledge for transformations specific to the biological sample can increase assignment success.
-However, the approach could be applied without the use of transformations where these are not relevant.
-The assigned molecular formulas could be putatively matched to compound databases such as KEGG using ionisation rules such as those of the MZedDB database and then apply functional enrichment analyses using a tool such as the R package *FELLA* [@draper2009;@picart2018].
-They could also be used to provide some information about precursor ions in LC-MS/MS fragmentation analyses.
-
-# Conclusions
-
-The approach presented here, and its implementation in the *assignments* R package, provides investigators with an automated means to assign molecular formulas to non-targeted ESI-MS metabolomics with the use of high-performance computing resources.
-
-Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted ionisation products could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS metabolome fingerprinting.
-It was also possible to assign molecular formulas of common primary and secondary metabolites present in the KEGG compound database to *m/z* features from leaf tissue of *B.
-distachyon* using both FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques.
-
-This approach can contribute to considerably reducing the workload of 'first pass' putative metabolite identification in non-targeted metabolomics analyses.
-Assignments can be used as inputs for enrichment analyses, providing a fast route to the biological interpretation of inherently complex data sets.
-
-# Declarations
-
-## Ethics approval and consent to participate
-
-Not applicable
-
-## Consent for publication
-
-Not applicable
-
-## Availability of data and materials
-
-The metabolomics data files reported in this paper are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949 (https://www.ebi.ac.uk/metabolights/MTBLS6949). 
-The R code use to analyse the data and generate the manuscript is available on GitHub at https://github.com/jasenfinch/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS.
-
-## Competing interests
-
-The authors declare that they have no competing interests.
-
-## Funding
-
-JF received funding from the charity Woodland Heritage.
-JF and LL were funded by The Biotechnology and Biological Science Research Council (BBSRC), Natural Environmental Research Council (NERC), Defra and Scottish Government for funding as part of the BAC-STOP and FUTURE OAK projects (grant refs:BB/T010886/1 , BB/T01069X/1).
-The financial support of HP was provided by Aberystwyth University and the UK Medical Research Council (MRC) funded TW and MB (MRC grant ref: MR/S010483/1).
-
-## Authors' contributions
-
-JF: conceptualisation, methodology, sample preparation, software development, data analysis and writing-original draft; TW: conceptualisation and manuscript revision; LL: sample preparation and data collection; HP: data collection; MB: conceptualisation, supervision and manuscript revision; JD: supervision and manuscript revision.
-
-## Acknowlegements
-
-Not applicable
-
-```{r word-count}
-# jfmisc::writingChecks('manuscript/manuscript.Rmd',
-#                       checks = 'word count')
-```
-
 # Additional files
@@
 The *m/z* features assigned to spiked chemical compounds in urine.
-
-# References

### 2023-08-07 manuscript/manuscript.Rmd
Add email address to manuscript
Move additional file information to a separate file

@@
 
-^\*^ Corresponding author
+^\*^ Correspondence to jsf9@aber.ac.uk
 
@@
 
-# Additional files
-
-Additional file 1.csv - Table S1 -
-Chemical identifiers of the compounds spiked into the urine extract.
-
-Additional file 2.pdf - Supplementary materials and methods -
-Further details outlining the materials and methods used in this article.
-
-Additional file 3.pdf - Supplementary tables -
-Supplementary tables S2, S3, S4 and S5 showing adduct, isotopic, transformation and R package benchmarking information respectively.
-
-Additional file 4.csv - Table S6 -
-The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.
-
-Additional file 5.csv Table S7 -
-The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.
-
-Additional file 6.csv Table S8 -
-The *m/z* features assigned to spiked chemical compounds in urine.
-
 # References

### 2023-08-07 R/targets/manuscript_targets.R

@@ -6,6 +6,14 @@ manuscript_targets <- list(
     assignmentApproach()
   ),
   
+  ## Additional file info
+  tarchetypes::tar_render(
+    additional_files,
+    'manuscript/additional_files.Rmd',
+    output_file = 'Additional_files.pdf',
+    output_dir = 'exports/manuscript/supplementary_materials'
+  ),
+  
   ## Supplementary materials and methods
   tarchetypes::tar_render(
     supplementary_methods,

### 2023-08-14 manuscript/manuscript.Rmd
Add the source of human urine to the main manuscript text

@@
 Two example sample matrices were used for assessing the molecular formula assignment approach. 
-The first was human urine spiked with a diverse mixture of `r nrow(spiked_urine_compound_info)` chemical standards from a previously validated assay that included `r spiked_urine_compound_MFs` unique molecular formulas [@beckmann2020].
+The first was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with a diverse mixture of `r nrow(spiked_urine_compound_info)` chemical standards from a previously validated assay that included `r spiked_urine_compound_MFs` unique molecular formulas [@beckmann2020].
 A full list of the names and InChI chemical identifiers of the spiked chemical standards can be found in Additional file 1.
@@
 
-Not applicable
+The human urine was purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) with no ethical approval required for use.
 

### 2023-11-04 R/targets/standards_targets.R
add static branching to process the spiked urine and standards matrices. Temporarily deactivate the spiked-urine evaluation targets

@@ -1,545 +1,595 @@
 
-spiked_urine_search_ppm <- 6
+search_ppm <- 4
 
-spiked_urine_assignment_parameters <- {
-  ap <- assignments::assignmentParameters('FIE')
-  assignments::adducts(ap)$n <- c("[M-H]1-",
-                                  "[M+Cl]1-",
-                                  "[M+K-2H]1-",
-                                  "[M+Cl37]1-",
-                                  "[M-2H]2-",
-                                  "[2M-H]1-")
-  assignments::ppm(ap) <- spiked_urine_search_ppm
-  assignments::adducts(ap)$p <- assignments::adducts(ap)$p[!stringr::str_detect(assignments::adducts(ap)$p,
-                                                                                stringr::coll("[M+NH4]1+"))]
-  assignments::isotopes(ap) <- c('13C','18O','34S','13C2')
-  ap
+assignment_parameters <- {
+  p <- assignments::assignmentParameters("FIE-HRMS") 
+  assignments::adducts(p)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
+  assignments::adducts(p)$p <- c("[M+Na]1+","[M+H]1+","[M+K]1+","[M+K41]1+","[M+2H]2+","[2M+H]1+")
+  assignments::isotopes(p) <- c("13C","18O","34S")
+  assignments::ppm(p) <- 4
+  p
 }
 
 spiked_urine_example_feature <- 'n80.97488' 
 
-spiked_urine_targets <- list(
-  ## spiked_urine compound info file
-  tarchetypes::tar_file(
-    spiked_urine_compound_info_file,
-    {
-      'spiked_chemical_standards_info.csv' %>% 
-        piggyback::pb_download(dest = 'data',
-                               tag = 'spiked-urine')
-      
-      'data/spiked_chemical_standards_info.csv' 
-    }
+
+matrix_info <- tibble::tibble(
+  sample = c(
+    'standards',
+    'spiked_urine'
   ),
-  
-  ## load spiked_urine compound info
-  tar_target(
-    spiked_urine_compound_info,
-    readr::read_csv(spiked_urine_compound_info_file) %>% 
-      dplyr::rowwise() %>% 
-      dplyr::group_split() %>% 
-      purrr::map_dfr(~ .x %>% 
-                       dplyr::mutate(SMILES = cheminf::convert(
-                         InChI,
[…]

### 2023-11-04 _targets.R

@@ -7,7 +7,7 @@
 list(
   KEGG_compound_targets,
   MF_selection_targets,
-  spiked_urine_targets,
+  standards_targets,
   brachy_targets,
   assignment_results_targets,
   benchmarking_targets,

### 2023-11-05 R/targets/brachy_targets.R
update the brachy assignment parameters

@@ -4,8 +4,33 @@ techniques <- c(
   'RP-UHPLC-HRMS'
 ) 
 
-brachy_targets <- purrr::map(
+assignment_parameters <- list(
+  `FIE-HRMS` = assignments::assignmentParameters("FIE-HRMS"),
+  `RP-UHPLC-HRMS` = assignments::assignmentParameters("RP-LC-HRMS")
+)
+
+assignments::adducts(assignment_parameters$`FIE-HRMS`)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
+assignments::adducts(assignment_parameters$`RP-UHPLC-HRMS`)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
+
+assignment_parameters <- assignment_parameters %>% 
+  purrr::map(
+    ~{
+      assignments::isotopes(.x) <- c('13C','18O')
+      .x
+    }
+  )
+
+assignment_parameters <- assignment_parameters %>% 
+  purrr::map(
+    ~{
+      assignments::ppm(.x) <- 4
+      .x
+    }
+  )
+
+brachy_targets <- purrr::map2(
   techniques,
+  assignment_parameters,
   ~{
     target_name <- .x %>% 
       stringr::str_replace_all('-','_') %>% 
@@ -36,8 +61,8 @@ brachy_targets <- purrr::map(
                                     export_path = NULL),
       hrmtargets::tar_mf_assignment(
         !!target_name,
-        parameters = assignments::assignmentParameters(.x %>%
-                                                      stringr::str_remove('UHP')),
+        parameters = .y,
+        separate = TRUE,
         export_path = NULL
       ),
       

### 2023-11-05 R/targets/manuscript_targets.R
add targets for evaluating standards samples

@@ -33,7 +33,7 @@ manuscript_targets <- list(
   ## Supplementary table S1 of spiked compounds
   tarchetypes::tar_file(
     table_s1,
-    spiked_urine_compound_info %>% 
+    standards_compound_info %>% 
       dplyr::mutate(MF = SMILES %>% 
                       cheminf::chemicalDescriptors() %>% 
                       .$MF) %>% 

### 2023-11-05 R/targets/standards_targets.R

@@ -34,6 +34,28 @@ matrix_info <- tibble::tibble(
   )
 )
 
+assignment_outcomes <- list(
+  standards = c(
+    'No correlations',
+    'No relevant relationships',
+    'No relevant relationships',
+    'Eliminated, unassigned',
+    'Alternative adduct, alternative MF',
+    'Matching adduct, MF outside top 3',
+    'Matching adduct, matching MF'
+  ),
+  spiked_urine = c(
+    'No correlations',
+    'No relevant relationships',
+    'No relevant relationships',
+    'Eliminated, unassigned',
+    'Alternative adduct, alternative MF',
+    'Matching adduct, MF outside top 3',
+    'Matching adduct, alternative MF',
+    'Matching adduct, matching MF'
+  ) 
+)
+
 standards_processing_targets <- matrix_info %>% 
   dplyr::rowwise() %>% 
   dplyr::group_split() %>% 
@@ -88,508 +110,529 @@ standards_processing_targets <- matrix_info %>%
                 )
               )
             )
-              )
           )
+        )
       )   
     }
   )
 
-standards_targets <- list(
-  standards_processing_targets
+standards_compound_targets <- list(
+  ## chemical standards compound info file
+  tarchetypes::tar_file(
+    standards_compound_info_file,
+    {
+      'spiked_chemical_standards_info.csv' %>%
+        piggyback::pb_download(dest = 'data',
+                               tag = 'spiked-urine')
+
+      'data/spiked_chemical_standards_info.csv'
+    }
+  ),
+
+  ## load chemical standards compound info
+  tar_target(
+    standards_compound_info,
+    readr::read_csv(standards_compound_info_file) %>%
+      dplyr::rowwise() %>%
+      dplyr::mutate(SMILES = cheminf::convert(
+        InChI,
+        'INCHI',
+        'SMILES'),
+        MF = cheminf::smilesToMF(SMILES))
+  ),
+
+  ## calculate chemical standards compound descriptors
+  tar_target(
+    standards_compound_db,
+    cheminf::metaboliteDB(
+      standards_compound_info %>%
+        dplyr::rename(ID = CID,
+                      NAME = Name)
+    )
+  ),
+
+  ## calculate possible adducts for the standards
+  tar_target(
+    standards_compound_adducts,
+    standards_compound_db %>%
+      cheminf::entries() %>%
+      .$ID %>%
+      furrr::future_map(cheminf::calcAdducts,
[…]

### 2023-11-05 R/targets/standards_targets.R
add assignment outcomes summary

@@ -627,7 +627,23 @@ standards_figures_targets <- list(
   
   ## Match assignments to human specific KEGG compounds
   tar_KEGG_matches(spiked_urine,
-                   KEGG_IRs_human)
+                   KEGG_IRs_human),
+  
+  tar_target(
+    assignment_outcomes_summary,
+    list(
+      standards = standards_assignment_outcomes_summary,
+      spiked_urine = spiked_urine_assignment_outcomes_summary
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'sample'
+      ) %>% 
+      dplyr::select(-`%`) %>% 
+      tidyr::spread(
+        sample,`# IPs`,
+        fill = 0
+      )
+  )
 )
 
 standards_targets <- list(

### 2023-11-05 R/targets/brachy_targets.R
return brachy FIE-HRMS back to 6 ppm

@@ -20,13 +20,7 @@ assignment_parameters <- assignment_parameters %>%
     }
   )
 
-assignment_parameters <- assignment_parameters %>% 
-  purrr::map(
-    ~{
-      assignments::ppm(.x) <- 4
-      .x
-    }
-  )
+assignments::ppm(assignment_parameters$`RP-UHPLC-HRMS`) <- 4
 
 brachy_targets <- purrr::map2(
   techniques,

### 2023-11-05 R/targets/standards_targets.R
add compound outcomes table

@@ -401,6 +401,36 @@ standards_evaluation_targets <- matrix_info %>%
               )
           )
         ),
+        
+        tar_target_raw(
+          paste0(.x$sample,'_assignments_T'),
+          rlang::expr(
+            !!rlang::sym(paste0(.x$sample,'_assignments')) %>%
+              dplyr::bind_rows() %>% 
+              dplyr::filter(
+                stringr::str_detect(
+                  Iteration,
+                  'T'
+                )
+              )
+          )
+        ),
+        
+        tar_target_raw(
+          paste0(.x$sample,'_compound_assignments_T'),
+          rlang::expr(
+            !!rlang::sym(paste0(.x$sample,'_feature_matches')) %>% 
+              dplyr::inner_join(
+                !!rlang::sym(paste0(.x$sample,'_assignments_T')) %>% 
+                  dplyr::select(
+                    MF,
+                    Isotope,
+                    Adduct
+                  ),
+                by = dplyr::join_by(MF, Isotope, Adduct)
+              ) 
+          )
+        ),
 
         tar_target_raw(
           paste0(.x$sample,'_assignment_outcomes'),
@@ -643,6 +673,133 @@ standards_figures_targets <- list(
         sample,`# IPs`,
         fill = 0
       )
+  ),
+  
+  tar_target(
+    assignment_outcomes_non_iso,
+    list(
+      standards = standards_assignment_outcomes_non_iso,
+      spiked_urine = spiked_urine_assignment_outcomes_non_iso
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'sample'
+      ) 
+  ),
+  
+  tar_target(
+    correct_assignments,
+    list(
+      standards = standards_correct_assignments,
+      spiked_urine = spiked_urine_correct_assignments
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'sample'
+      )
+  ),
+  
+  tar_target(
+    compound_assignments_T,
+    list(
+      standards = standards_compound_assignments_T,
+      spiked_urine = spiked_urine_compound_assignments_T
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'sample'
+      )
+  ),
+  
+  tar_target(
+    standards_compound_outcomes,
+    assignment_outcomes_non_iso %>% 
+      dplyr::select(sample,MF) %>% 
+      dplyr::distinct() %>% 
+      dplyr::count(sample) %>% 
+      dplyr::rename(
+        `Matched to m/z features` = n
+      ) %>% 
[…]

### 2023-11-05 manuscript/manuscript.Rmd
add the standards mix to the results table

@@
                                         ' ',2)[,2]) %>% 
-  knitr::kable(format = 'latex',
+  knitr::kable(
+    format = 'latex',
     booktabs = TRUE,
@@
     "\\\\textit{B. distachyon} leaf" = 2,
-    "Human urine" = 1),
+    "Human urine" = 1,
+    "Standards mix" = 1),
     escape = FALSE)

### 2023-11-05 R/targets/assignment_results_targets.R

@@ -3,6 +3,7 @@ assignment_results_targets <- list(
   tar_target(
     assignment_results,
     list(
+      `standards-mix_FIE-HRMS` = standards_results_molecular_formula_assignment,
       `spiked-urine_FIE-HRMS` = spiked_urine_results_molecular_formula_assignment,
       `brachy_FIE-HRMS` = brachy_FIE_HRMS_results_molecular_formula_assignment,
       `brachy_C18-UHPLC-HRMS` = brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment
@@ -36,13 +37,33 @@ assignment_results_targets <- list(
         KEGG_compound_matches = nrow(.x)
       )) %>% 
       bind_rows(.id = 'matrix') %>% 
-      mutate(
-        matched_standards = c(spiked_urine_correct_assignments$MF %>% 
-                                unique() %>% 
-                                length(),
-                              rep(NA,2))
-      ) %>% 
       tidyr::separate(matrix,
-                      c('matrix','technique'),sep = '_')
+                      c('matrix','technique'),sep = '_') %>% 
+      dplyr::full_join(
+        correct_assignments %>% 
+          dplyr::select(
+            sample,
+            MF
+          ) %>% 
+          dplyr::distinct() %>% 
+          dplyr::count(
+            sample
+          ) %>% 
+          dplyr::rename(
+            matched_standards = n
+          ) %>% 
+          dplyr::mutate(
+            sample = sample %>% 
+              stringr::str_replace(
+                '_',
+                '-'
+              ),
+            technique = 'FIE-HRMS'
+          ),
+        by = dplyr::join_by(
+          matrix == sample,
+          technique
+        )
+      )  
   )
 )

### 2023-11-05 manuscript/manuscript.Rmd
added the adduct and isotopic step and transformation step assignment frequencies to table 1

@@
             by = c("matrix", "technique")) %>% 
-  mutate(matrix = paste(matrix,technique)#,
+  mutate(
+    matrix = paste(matrix,technique)
   ) %>%
@@
                 relationships = n_relationships,
+                `adduct and isotopic step assignments` = n_ai_assignments,
+                `transformation step assignments` = n_t_assignments,
                 `assigned features` = n_assignments,
@@
   tidyr::spread(matrix,Value) %>% 
-  slice(c(1,3,8,2,7,5,4,6)) %>%
+  slice(c(1,3,10,8,9,2,7,5,4,6)) %>%
   set_colnames(stringr::str_split_fixed(colnames(.),

### 2023-11-05 R/targets/assignment_results_targets.R

@@ -11,6 +11,24 @@ assignment_results_targets <- list(
       n_features = ncol(.x@data),
       n_correlations = nrow(.x@correlations),
       n_relationships = nrow(.x@relationships),
+      n_ai_assignments = .x %>% 
+        assignments::assignments() %>% 
+        dplyr::filter(
+          stringr::str_detect(
+            Iteration,
+            'A&I'
+          )
+        ) %>% 
+        nrow(),
+      n_t_assignments = .x %>% 
+        assignments::assignments() %>% 
+        dplyr::filter(
+          stringr::str_detect(
+            Iteration,
+            'T'
+          )
+        ) %>% 
+        nrow(),
       n_assignments = .x %>% 
         assignments::assignments() %>% 
         nrow(),

### 2023-11-06 manuscript/additional_files.Rmd
Move table 1 to supplementary table S6

@@
 Additional file 3.pdf - Supplementary tables -
-Supplementary tables S2, S3, S4 and S5 showing adduct, isotopic, transformation and R package benchmarking information respectively.
+Supplementary tables S2, S3, S4, S5, S6 showing adduct, isotopic, transformation, R package benchmarking information and a summary of the molecular formula assignment results for the example sample matrices respectively.
 
-Additional file 4.csv - Table S6 -
+Additional file 4.csv - Table S7 -
 The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.
 
-Additional file 5.csv Table S7 -
+Additional file 5.csv Table S8 -
 The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.
 
-Additional file 6.csv Table S8 -
+Additional file 6.csv Table S9 -
 The *m/z* features assigned to spiked chemical compounds in urine.

### 2023-11-06 manuscript/manuscript.Rmd
Move table 1 to supplementary table S6

--- 2023-08-14 manuscript/manuscript.Rmd
@@
 ```{r assignment-results}
-tar_read(assignment_results) %>%
-  left_join(tar_read(assignment_matches), 
-            by = c("matrix", "technique")) %>% 
-  mutate(matrix = paste(matrix,technique)#,
-  ) %>%
-  mutate(n_assignments = glue::glue(
-    '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
-    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
-  select(-technique) %>% 
-  dplyr::rename(`\\textit{m/z} features` = n_features,
-                correlations = n_correlations,
-                relationships = n_relationships,
-                `assigned features` = n_assignments,
-                `molecular formulas (MF)` = n_MFs,
-                `KEGG compound matches` = KEGG_compound_matches,
-                `KEGG MF matches` = KEGG_MF_matches,
-                `matched chemical standards` = matched_standards) %>% 
-  tidyr::gather(Variable,Value,-matrix) %>% 
-  mutate(Value = replace(Value,
-                         Value == 'NA',
-                       'N/A')) %>% 
-  tidyr::spread(matrix,Value) %>% 
-  slice(c(1,3,8,2,7,5,4,6)) %>%
-  set_colnames(stringr::str_split_fixed(colnames(.),
-                                        ' ',2)[,2]) %>% 
-  knitr::kable(format = 'latex',
-               booktabs = TRUE,
-               linesep = '',
-               escape = FALSE,
-               align = c(
-                 rep('l',1),
-                 rep('r',3)
-               ),
-               caption = 'A summary of the molecular formula assignment results for the example sample matrices.
-The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
-  kable_styling(latex_options = "HOLD_position") %>% 
-  add_header_above(c(
-    " ", 
-    "\\\\textit{B. distachyon} leaf" = 2, 
-    "Human urine" = 1),
-    escape = FALSE)
 ```

### 2023-11-06 manuscript/supplementary_tables.Rmd
Move table 1 to supplementary table S6

@@
 ```{r setup,include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,
+knitr::opts_chunk$set(
+  echo = FALSE,
+  warning = FALSE,
   message = FALSE)
@@
 ```
+
+```{r table-s6}
+tar_read(assignment_results) %>%
+  dplyr::full_join(
+    tar_read(assignment_matches) %>% 
+      dplyr::mutate(
+        matrix = matrix %>% 
+          replace(
+            matrix == 'standards',
+            'standards-mix'
+          )
+      ), 
+    by = c("matrix", "technique")) %>% 
+  mutate(
+    matrix = paste(matrix,technique)
+  ) %>%
+  mutate(
+    n_assignments = glue::glue(
+      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
+    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
+  select(-technique) %>% 
+  dplyr::rename(
+    `\\textit{m/z} features` = n_features,
+    correlations = n_correlations,
+    relationships = n_relationships,
+    `adduct and isotopic step assignments` = n_ai_assignments,
+    `transformation step assignments` = n_t_assignments,
+    `assigned features` = n_assignments,
+    `molecular formulas (MF)` = n_MFs,
+    `KEGG compound matches` = KEGG_compound_matches,
+    `KEGG MF matches` = KEGG_MF_matches,
+    `matched chemical standards` = matched_standards) %>% 
+  tidyr::gather(Variable,Value,-matrix) %>% 
+  mutate(
+    Value = replace(Value,
+                    Value == 'NA',
+                    'N/A')) %>% 
+  tidyr::spread(matrix,Value) %>% 
+  slice(c(1,4,9,2,10,3,8,5,6,7)) %>%
+  set_colnames(
+    stringr::str_split_fixed(colnames(.),
+                             ' ',2)[,2]) %>% 
+  knitr::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    linesep = '',
+    escape = FALSE,
+    align = c(
+      rep('l',1),
+      rep('r',3)
+    ),
+    caption = 'A summary of the molecular formula assignment results for the example sample matrices.
+The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
+  kable_styling() %>%
+  add_header_above(c(
+    " ",
+    "\\\\textit{B. distachyon} leaf" = 2,
+    "Human urine" = 1,
+    "Standards mix" = 1),
+    escape = FALSE)
+```

### 2023-11-06 R/targets/manuscript_targets.R

@@ -44,7 +44,7 @@ manuscript_targets <- list(
   
   ## Supplementary table S5 of all assignments
   tarchetypes::tar_file(
-    table_s6,
+    table_s7,
     bind_rows(
       list(
         `Human_urine_FIE-HRMS` = spiked_urine_results_molecular_formula_assignment %>% 
@@ -105,7 +105,7 @@ manuscript_targets <- list(
   
   ## Supplementary table S6 directly matched chemical standard ionisation products
   tarchetypes::tar_file(
-    table_s7,
+    table_s8,
     spiked_urine_feature_matches %>% 
       select(MF,
              Isotope,
@@ -120,7 +120,7 @@ manuscript_targets <- list(
   
   ## Supplementary table S7 assigned chemical standards
   tarchetypes::tar_file(
-    table_s8,
+    table_s9,
     spiked_urine_correct_assignments %>% 
       select(MF,
              Isotope,
@@ -148,9 +148,9 @@ manuscript_targets <- list(
           c(supplementary_methods[1],
             supplementary_tables[1],
             table_s1,
-            table_s6,
             table_s7,
-            table_s8),
+            table_s8,
+            table_s9),
           flags = '-r9Xj')
       
       zipfile

### 2023-11-06 manuscript/manuscript.Rmd
update target names in the manuscript where necessary. Deactivated variables/targets that no longer exist

@@
 ```{r assignment-summary}
-tar_load(assignment_results)
-tar_load(assignment_matches)
-tar_load(spike_urine_standards_MF_assignment_possible)
-tar_load(spiked_urine_correct_assignments)
-
-total_assigned <- round(sum(assignment_results$n_assignments) / 
-                          sum(assignment_results$n_features) * 100,1)
-total_mfs <- sum(assignment_results$n_MFs)
-total_kegg <- sum(assignment_matches$KEGG_compound_matches)
-
-spiked_urine_possible <- nrow(spike_urine_standards_MF_assignment_possible)
-spiked_urine_correct_percent <- spiked_urine_correct_assignments$MF %>% 
-  unique() %>% 
-  length() %>% 
-  {. / spiked_urine_possible * 100} %>% 
-  round(digits = 1)
+# tar_load(assignment_results)
+# tar_load(assignment_matches)
+# tar_load(spike_urine_standards_MF_assignment_possible)
+# tar_load(spiked_urine_correct_assignments)
+# 
+# total_assigned <- round(sum(assignment_results$n_assignments) / 
+#                           sum(assignment_results$n_features) * 100,1)
+# total_mfs <- sum(assignment_results$n_MFs)
+# total_kegg <- sum(assignment_matches$KEGG_compound_matches)
+# 
+# spiked_urine_possible <- nrow(spike_urine_standards_MF_assignment_possible)
+# spiked_urine_correct_percent <- spiked_urine_correct_assignments$MF %>% 
+#   unique() %>% 
+#   length() %>% 
+#   {. / spiked_urine_possible * 100} %>% 
+#   round(digits = 1)
 ```
@@
 The approach was tested in human urine samples spiked with known chemical standards and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
-Across the three test matrices, `r total_assigned`% of the m/z features were assigned a molecular formula giving a total of `r total_mfs` molecular formulas and matching to `r total_kegg` compounds in the KEGG compound database.
-Of the `r spiked_urine_possible` spiked chemical standards that it was possible to assign in the human urine samples, the approach was able to assign m/z features to `r spiked_urine_correct_percent`% of these.
+Across the three test matrices, `r #total_assigned`% of the m/z features were assigned a molecular formula giving a total of `r #total_mfs` molecular formulas and matching to `r #total_kegg` compounds in the KEGG compound database.
+Of the `r #spiked_urine_possible` spiked chemical standards that it was possible to assign in the human urine samples, the approach was able to assign m/z features to `r #spiked_urine_correct_percent`% of these.
 
@@
 ```{r chemical-standards}
-tar_load(spiked_urine_compound_info)
+tar_load(standards_compound_info)
 
-spiked_urine_compound_MFs <- spiked_urine_compound_info %>% 
+standards_compound_MFs <- standards_compound_info %>% 
   select(MF) %>% 
@@
 Two example sample matrices were used for assessing the molecular formula assignment approach. 
-The first was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with a diverse mixture of `r nrow(spiked_urine_compound_info)` chemical standards from a previously validated assay that included `r spiked_urine_compound_MFs` unique molecular formulas [@beckmann2020].
+The first was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with a diverse mixture of `r nrow(standards_compound_info)` chemical standards from a previously validated assay that included `r standards_compound_MFs` unique molecular formulas [@beckmann2020].
 A full list of the names and InChI chemical identifiers of the spiked chemical standards can be found in Additional file 1.
@@
 Features matching to possible ionisation products of the spiked urine compounds were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
-The calculated *m/z* were then matched to the spiked urine features using a search range of $\pm$ `r spiked_urine_search_ppm` ppm.
+The calculated *m/z* were then matched to the spiked urine features using a search range of $\pm$ `r search_ppm` ppm.
 
@@
 ```{r assignment-success}
-tar_load(spiked_urine_compound_PIPs)
+tar_load(standards_compound_PIPs)
 tar_load(spiked_urine_feature_matches)
-tar_load(spiked_urine_unassigned)
-tar_load(spiked_urine_assigned)
-tar_load(spiked_urine_incorrect_assignments)
+# tar_load(spiked_urine_unassigned)
+# tar_load(spiked_urine_assigned)
+# tar_load(spiked_urine_incorrect_assignments)
 
@@
 
-spiked_urine_MF_PIPs <- spiked_urine_compound_PIPs %>% 
+spiked_urine_MF_PIPs <- standards_compound_PIPs %>% 
   select(MF,Adduct,Isotope) %>% 
@@
 
-percent_unassigned <- {nrow(spiked_urine_unassigned) / 
-    nrow(spiked_urine_feature_matches) * 100} %>% 
-  signif(digits = 3)
+# percent_unassigned <- {nrow(spiked_urine_unassigned) / 
+#     nrow(spiked_urine_feature_matches) * 100} %>% 
+#   signif(digits = 3)
 
-spiked_urine_unassigned_MF_n <- spiked_urine_unassigned %>% 
-  anti_join(
-    spiked_urine_assigned %>% 
-      select(MF) %>% 
-      distinct(), 
-    by = 'MF') %>% 
-  .$MF %>% 
-  unique() %>% 
-  length()
+# spiked_urine_unassigned_MF_n <- spiked_urine_unassigned %>% 
+#   anti_join(
+#     spiked_urine_assigned %>% 
+#       select(MF) %>% 
+#       distinct(), 
+#     by = 'MF') %>% 
+#   .$MF %>% 
+#   unique() %>% 
+#   length()
 
-spiked_urine_correct_assignments_MF_n <- spiked_urine_correct_assignments %>%
-  .$MF %>% 
-  unique() %>% 
-  length()
+# spiked_urine_correct_assignments_MF_n <- spiked_urine_correct_assignments %>%
+#   .$MF %>% 
+#   unique() %>% 
+#   length()
 
-spike_urine_standards_MF_assignment_impossible_n <- spiked_urine_feature_matches_MF_n -
-  nrow(spike_urine_standards_MF_assignment_possible)
+# spike_urine_standards_MF_assignment_impossible_n <- spiked_urine_feature_matches_MF_n -
+#   nrow(spike_urine_standards_MF_assignment_possible)
 
-percent_standards_MF_assigned <- {spiked_urine_correct_assignments_MF_n / 
-  nrow(spike_urine_standards_MF_assignment_possible) * 100} %>% 
-  signif(digits = 3)
+# percent_standards_MF_assigned <- {spiked_urine_correct_assignments_MF_n / 
+#   nrow(spike_urine_standards_MF_assignment_possible) * 100} %>% 
+#   signif(digits = 3)
 ```
@@
 Based on the `r positive_mode_adducts + negative_mode_adducts` adducts and four isotopic rules specified, there were `r nrow(spiked_urine_MF_PIPs)` predicted ionisation product *m/z* for these chemical standards.
-A total of `r nrow(spiked_urine_feature_matches)` of these ionisation products could be directly matched to the observed *m/z* features which included `r spiked_urine_feature_matches_MF_n` of the `r spiked_urine_compound_MFs` unique molecular formulas of the spiked chemical standards.
+A total of `r nrow(spiked_urine_feature_matches)` of these ionisation products could be directly matched to the observed *m/z* features which included `r spiked_urine_feature_matches_MF_n` of the `r standards_compound_MFs` unique molecular formulas of the spiked chemical standards.
 These directly matched *m/z* features are provided in Additional file 5.
-For `r nrow(spiked_urine_unassigned)` (`r percent_unassigned`%) of these matched *m/z* features, no molecular formula was assigned by the approach, which included `r spiked_urine_unassigned_MF_n` of the molecular formulas of the spiked chemical standards.
-However, `r nrow(spiked_urine_assigned)` of the matched features were assigned a molecular formula with `r nrow(spiked_urine_correct_assignments)` of these corresponding to one of the spiked compound ionisation products.
-This included `r spiked_urine_correct_assignments_MF_n` of the spiked chemical standards molecular formulas (Table \@ref(tab:assignment-results)).
+For `r #nrow(spiked_urine_unassigned)` (`r #percent_unassigned`%) of these matched *m/z* features, no molecular formula was assigned by the approach, which included `r #spiked_urine_unassigned_MF_n` of the molecular formulas of the spiked chemical standards.
+However, `r #nrow(spiked_urine_assigned)` of the matched features were assigned a molecular formula with `r #nrow(spiked_urine_correct_assignments)` of these corresponding to one of the spiked compound ionisation products.
+This included `r #spiked_urine_correct_assignments_MF_n` of the spiked chemical standards molecular formulas (Table \@ref(tab:assignment-results)).
 The *m/z* features assigned to molecular formulas corresponding to the spiked chemical standards are provided in Additional file 6.
 
-The remaining `r nrow(spiked_urine_incorrect_assignments)` matched *m/z* features were assigned alternative molecular formula ionisation products.
+The remaining `r #nrow(spiked_urine_incorrect_assignments)` matched *m/z* features were assigned alternative molecular formula ionisation products.
 The *m/z* of the predicted ionisation features that it was not possible to assign molecular formulas were as a result of either not being correlated with other *m/z* features that corresponded to a correct adduct or isotope mass difference or whose subgraph components were not selected, with more plausible solutions instead selected based on the correlation network.
 
-Accounting for the `r spike_urine_standards_MF_assignment_impossible_n` chemical standards that were not assigned due to a lack of relevant correlated mass relationships, the algorithm was able to assign molecular formulas to features corresponding to `r percent_standards_MF_assigned`% of the `r nrow(spike_urine_standards_MF_assignment_possible)` possible chemical standards.
+Accounting for the `r #spike_urine_standards_MF_assignment_impossible_n` chemical standards that were not assigned due to a lack of relevant correlated mass relationships, the algorithm was able to assign molecular formulas to features corresponding to `r #percent_standards_MF_assigned`% of the `r #nrow(spike_urine_standards_MF_assignment_possible)` possible chemical standards.
 This highlights the difficulty in annotating these complex biological mixtures containing hundreds of compounds, which can create many thousands of possible ionisation products.

### 2023-11-06 manuscript/manuscript.Rmd
add the chemical standards evaluation tables to the manuscript

@@
 
-```{r assignment-results}
+```{r assignment-evaluations}
+tar_load(assignment_outcomes_summary)
+
+assignment_outcomes_total <- assignment_outcomes_summary %>% 
+  dplyr::summarise(
+    standards = sum(standards),
+    spiked_urine = sum(spiked_urine)
+  ) %>% 
+  dplyr::mutate(
+    ` ` = 'Total'
+  )
+
+assignment_outcomes_summary %>% 
+  dplyr::bind_rows(
+    assignment_outcomes_total
+  ) %>% 
+  dplyr::relocate(
+    ` `,
+    .before = dplyr::everything()
+  ) %>% 
+  dplyr::relocate(
+    spiked_urine,
+    .after = dplyr::everything()
+  ) %>% 
+  dplyr::mutate(
+    `Assignment outcome` = as.character(
+      `Assignment outcome`
+    ),
+    dplyr::across(
+      dplyr::everything(),
+      ~replace(
+        .x,
+        is.na(.x),
+        ' '
+      )
+    )
+  ) %>%
+  dplyr::rename(
+    `Spiked urine` = spiked_urine,
+    `Standards mix` = standards
+  ) %>% 
+  kableExtra::kbl(
+    format = 'latex',
+    col.names = c('','','Standards mix','Spiked urine'),
+    booktabs = TRUE,
+    linesep = '',
+    align = c(
+      rep('l',2),
+      rep('r',2)
+    ),
+    caption = 'The frequencies of molecular formula (MF) assignment outcomes for theoretical, non-isotopic ionisation products that were matched to \\textit{m/z} features in the chemical standards mix and the same standards mix spiked into human urine.
+These frequencies include only outcomes from the adduct and isotopic assignment iterations and highlight the perfomance of the MF and graphical component selection approaches.'
+  ) %>% 
+  kableExtra::kable_styling(
+    latex_options = "HOLD_position"
+  ) %>% 
+  kableExtra::collapse_rows(
+    1,
+    latex_hline = 'major'
+  )
+```
+
+\linespread{2}
+
+\linespread{1}
+
+```{r compound-evaluations}
+tar_load(standards_compound_outcomes)
+
+standards_compound_outcomes %>%
+  dplyr::mutate(
+    outcome = outcome %>%
+      stringr::str_replace_all(
+        'm/z',
+        '\\\\textit{m/z}'
+      ) %>% 
+      stringr::str_replace_all(
+        '&',
+        '\\\\&'
+      )
+  ) %>%
+  dplyr::relocate(
+    spiked_urine,
+    .after = dplyr::everything()
+  ) %>%
+  dplyr::rename(
+    ` ` = outcome,
+    `Standards mix` = standards,
+    `Spiked urine` = spiked_urine
+  ) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    escape = FALSE,
+    linesep = '',
+    align = c(
+      rep('l',1),
+      rep('r',2)
+    ),
+    caption = 'Assignments results frequencies for the 64 unique molecular formulas of the chemical standards in the standards mix and spiked into human urine.'
+  ) %>%
+  kableExtra::kable_styling(
+    latex_options = "HOLD_position"
+  )
 ```

### 2023-11-06 R/targets/standards_targets.R

@@ -726,7 +726,7 @@ standards_figures_targets <- list(
           dplyr::distinct() %>% 
           dplyr::count(sample) %>% 
           dplyr::rename(
-            `Correlated` = n
+            `Correlations obtained` = n
           ),
         by = 'sample'
       ) %>% 
@@ -789,8 +789,7 @@ standards_figures_targets <- list(
           outcome,
           levels = c(
             'Matched to m/z features',
-            'Correlated',
-            'Relevant relationships',
+            'Correlations obtained',
             'Assigned during A&I assignment step',
             'Assigned during T assignment step',
             'Assigned'

### 2023-11-06 manuscript/manuscript.Rmd
Correct table 2 caption.

@@
     ),
-    caption = 'Assignments results frequencies for the 64 unique molecular formulas of the chemical standards in the standards mix and spiked into human urine.'
+    caption = 'Molecular formula assignment results frequencies for the 64 unique molecular formulas of the chemical standards in the standards mix and spiked into human urine.'
   ) %>%

### 2023-11-06 manuscript/manuscript.Rmd
Include the standards mixture in the methods section

@@
 
-Two example sample matrices were used for assessing the molecular formula assignment approach. 
-The first was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with a diverse mixture of `r nrow(standards_compound_info)` chemical standards from a previously validated assay that included `r standards_compound_MFs` unique molecular formulas [@beckmann2020].
-A full list of the names and InChI chemical identifiers of the spiked chemical standards can be found in Additional file 1.
-The second was healthy leaf tissue samples collected from four ecotypes of the model grass species *Brachypodium distachyon* using the method described in @parker2008.
+Three example sample matrices were used for assessing the molecular formula assignment approach. 
+The first was a diverse mixture of `r nrow(standards_compound_info)` chemical standards from a previously validated assay that included `r standards_compound_MFs` unique molecular formulas [@beckmann2020].
+A full list of the names and InChI chemical identifiers of the chemical standards can be found in Additional file 1.
+The second was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with the sample mixture of chemical standards as used above.
+The third was healthy leaf tissue samples collected from four ecotypes of the model grass species *Brachypodium distachyon* using the method described in @parker2008.
 Further details of the sample collection and extraction are described in the Additional file 2.
@@
 
-For each biological sample matrix, the input data for molecular formula assignment had undergone spectral processing and pre-treatment and consisted of a sample by *m/z* feature intensity matrix containing both positive and negative ionisation mode data.
+For each sample matrix, the input data for molecular formula assignment had undergone spectral processing and pre-treatment and consisted of a sample by *m/z* feature intensity matrix containing both positive and negative ionisation mode data.
 Molecular formula assignment was conducted using the R package *assignments* v`r packageVersion('assignments')` and the approach comprised four major steps which are outlined in Figure \@ref(fig:figure-1-approach).

### 2023-11-06 manuscript/manuscript.Rmd
Update the assignment assessments section of the methods to include the standards mix data

@@
 
-Features matching to possible ionisation products of the spiked urine compounds were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
-The calculated *m/z* were then matched to the spiked urine features using a search range of $\pm$ `r search_ppm` ppm.
+Features matching to possible ionisation products of the chemical standards were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
+The calculated *m/z* were then matched to the *m/z* features in the standards mix and spiked urine data using a search range of $\pm$ `r search_ppm` ppm.
 

### 2023-11-06 manuscript/additional_files.Rmd
add the standards mix to additional files 4-6

@@
 Additional file 5.csv Table S8 -
-The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.
+The *m/z* features directly matched to theoretical ionisation products of the chemical standards in the standards mixed and spiked urine.
 
 Additional file 6.csv Table S9 -
-The *m/z* features assigned to spiked chemical compounds in urine.
+The *m/z* features assigned to chemical standards in the standards mix and spiked urine.

### 2023-11-06 R/targets/manuscript_targets.R

@@ -47,6 +47,8 @@ manuscript_targets <- list(
     table_s7,
     bind_rows(
       list(
+        `NA_Standards mix_FIE-HRMS` = standards_results_molecular_formula_assignment %>% 
+          assignments::assignments(),
         `Human_urine_FIE-HRMS` = spiked_urine_results_molecular_formula_assignment %>% 
           assignments::assignments(),
         `Brachypodium distachyon_leaf_FIE-HRMS` = brachy_FIE_HRMS_results_molecular_formula_assignment %>% 
@@ -62,6 +64,13 @@ manuscript_targets <- list(
           'Matrix',
           'Technique'),
         sep = '_'
+      ) %>% 
+      dplyr::mutate(
+        Organism = Organism %>% 
+          replace(
+            Organism == 'NA',
+            NA
+          )
       ) %>% 
       left_join(
         bind_rows(
@@ -106,13 +115,22 @@ manuscript_targets <- list(
   ## Supplementary table S6 directly matched chemical standard ionisation products
   tarchetypes::tar_file(
     table_s8,
-    spiked_urine_feature_matches %>% 
-      select(MF,
-             Isotope,
-             Adduct,
-             `m/z` = mz,
-             `Ionisation mode` = Mode,
-             `PPM error`) %>% 
+    list(
+      `Standards mix` = standards_feature_matches,
+      `Spiked urine` = spiked_urine_feature_matches
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'Matrix'
+      ) %>% 
+      select(
+        Matrix,
+        MF,
+        Isotope,
+        Adduct,
+        `m/z` = mz,
+        `Ionisation mode` = Mode,
+        `PPM error`
+      ) %>% 
       jfmisc::exportCSV(
         file = 'exports/manuscript/supplementary_materials/Additional file 5.csv'
       )
@@ -121,13 +139,22 @@ manuscript_targets <- list(
   ## Supplementary table S7 assigned chemical standards
   tarchetypes::tar_file(
     table_s9,
-    spiked_urine_correct_assignments %>% 
-      select(MF,
-             Isotope,
-             Adduct,
-             `m/z` = `Measured m/z`,
-             `Ionisation mode` = Mode,
-             `PPM error`) %>% 
+    list(
+      `Standards mix` = standards_correct_assignments,
+      `Spiked urine` = spiked_urine_correct_assignments
+    ) %>%
[…]

### 2023-11-08 R/targets/brachy_targets.R
Add rerun of Brachypodium FIE-HRMS data

@@ -9,18 +9,15 @@ assignment_parameters <- list(
   `RP-UHPLC-HRMS` = assignments::assignmentParameters("RP-LC-HRMS")
 )
 
-assignments::adducts(assignment_parameters$`FIE-HRMS`)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
-assignments::adducts(assignment_parameters$`RP-UHPLC-HRMS`)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
-
 assignment_parameters <- assignment_parameters %>% 
   purrr::map(
     ~{
+      assignments::adducts(.x)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
       assignments::isotopes(.x) <- c('13C','18O')
+      assignments::ppm(.x) <- 4
       .x
     }
   )
-
-assignments::ppm(assignment_parameters$`RP-UHPLC-HRMS`) <- 4
 
 brachy_targets <- purrr::map2(
   techniques,
@@ -38,14 +35,28 @@ brachy_targets <- purrr::map2(
       pre_treatment_parameters <- NULL
     }
     
+    input_targets <- switch(
+      .x,
+      `FIE-HRMS` = hrmtargets::tar_input_piggyback(
+        !!target_name,
+        repo = 'jasenfinch/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS-devel',
+        tag = 'brachypodium_FIE-HRMS',
+        sample_info_file = 'sample_information.csv'  
+      ),
+      `RP-UHPLC-HRMS` = hrmtargets::tar_input_file_path(
+        !!target_name,
+        mzML_files = metaboData::filePaths(
+          .x,
+          'BdistachyonEcotypes',
+          ask = FALSE),
+        sample_info = metaboData::runinfo(
+          .x,
+          'BdistachyonEcotypes',
+          ask = FALSE)) 
+    )
+    
     list( 
-      hrmtargets::tar_input_file_path(!!target_name,
-                                      mzML_files = metaboData::filePaths(.x,
-                                                                         'BdistachyonEcotypes',
-                                                                         ask = FALSE),
-                                      sample_info = metaboData::runinfo(.x,
[…]

### 2023-11-08 R/utils.R

@@ -13,6 +13,6 @@ targets::tar_option_set(
 ## Parallel backend
 metaboMisc::suitableParallelPlan(
   workers = jfmisc::suitableParallelWorkers(
-    RAM_per_worker = 32
+    RAM_per_worker = 12
   )
 )

### 2023-11-08 manuscript/supplementary_tables.Rmd
fix the order of the rows in table S6

--- 2023-08-03 manuscript/supplementary_tables.Rmd
@@
 ```
+
+```{r table-s6}
+tar_read(assignment_results) %>%
+  dplyr::full_join(
+    tar_read(assignment_matches) %>% 
+      dplyr::mutate(
+        matrix = matrix %>% 
+          replace(
+            matrix == 'standards',
+            'standards-mix'
+          )
+      ), 
+    by = c("matrix", "technique")) %>% 
+  mutate(
+    matrix = paste(matrix,technique)
+  ) %>%
+  mutate(
+    n_assignments = glue::glue(
+      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
+    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
+  select(-technique) %>% 
+  dplyr::rename(
+    `\\textit{m/z} features` = n_features,
+    correlations = n_correlations,
+    relationships = n_relationships,
+    `adduct and isotopic step assignments` = n_ai_assignments,
+    `transformation step assignments` = n_t_assignments,
+    `assigned features` = n_assignments,
+    `molecular formulas (MF)` = n_MFs,
+    `KEGG compound matches` = KEGG_compound_matches,
+    `KEGG MF matches` = KEGG_MF_matches,
+    `matched chemical standards` = matched_standards) %>% 
+  tidyr::gather(Variable,Value,-matrix) %>% 
+  mutate(
+    Value = replace(Value,
+                    Value == 'NA',
+                    'N/A')) %>% 
+  tidyr::spread(matrix,Value) %>% 
+  slice(c(1,4,9,2,10,3,8,6,5,7)) %>%
+  set_colnames(
+    stringr::str_split_fixed(colnames(.),
+                             ' ',2)[,2]) %>% 
+  knitr::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    linesep = '',
+    escape = FALSE,
+    align = c(
+      rep('l',1),
+      rep('r',3)
+    ),
+    caption = 'A summary of the molecular formula assignment results for the example sample matrices.
+The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
+  kable_styling() %>%
+  add_header_above(c(
+    " ",
+    "\\\\textit{B. distachyon} leaf" = 2,
+    "Human urine" = 1,
+    "Standards mix" = 1),
+    escape = FALSE)
+```

### 2023-11-08 manuscript/supplementary_tables.Rmd
removed the matched chemical standards row from table S6 as this is now provided in table 2

--- 2023-08-03 manuscript/supplementary_tables.Rmd
@@
 ```
+
+```{r table-s6}
+tar_read(assignment_results) %>%
+  dplyr::full_join(
+    tar_read(assignment_matches) %>% 
+      dplyr::mutate(
+        matrix = matrix %>% 
+          replace(
+            matrix == 'standards',
+            'standards-mix'
+          )
+      ), 
+    by = c("matrix", "technique")) %>% 
+  mutate(
+    matrix = paste(matrix,technique)
+  ) %>%
+  mutate(
+    n_assignments = glue::glue(
+      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
+    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
+  select(-technique) %>% 
+  dplyr::rename(
+    `\\textit{m/z} features` = n_features,
+    correlations = n_correlations,
+    relationships = n_relationships,
+    `adduct and isotopic step assignments` = n_ai_assignments,
+    `transformation step assignments` = n_t_assignments,
+    `assigned \\textit{m/z}` = n_assignments,
+    `molecular formulas (MF)` = n_MFs,
+    `KEGG compound matches` = KEGG_compound_matches,
+    `KEGG MF matches` = KEGG_MF_matches,
+    `matched chemical standards` = matched_standards) %>% 
+  tidyr::gather(Variable,Value,-matrix) %>% 
+  mutate(
+    Value = replace(Value,
+                    Value == 'NA',
+                    'N/A')) %>% 
+  tidyr::spread(matrix,Value) %>% 
+  slice(c(1,4,9,2,10,3,8,6,5)) %>%
+  set_colnames(
+    stringr::str_split_fixed(colnames(.),
+                             ' ',2)[,2]) %>% 
+  knitr::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    linesep = '',
+    escape = FALSE,
+    align = c(
+      rep('l',1),
+      rep('r',3)
+    ),
+    caption = 'A summary of the molecular formula assignment results for the example sample matrices.
+The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
+  kable_styling() %>%
+  add_header_above(c(
+    " ",
+    "\\\\textit{B. distachyon} leaf" = 2,
+    "Human urine" = 1,
+    "Standards mix" = 1),
+    escape = FALSE)
+```

### 2023-11-08 manuscript/supplementary_methods.Rmd
Update the MS analysis section of the supplementary methods with the rerun of the Brachypodium FIE-HRMS data

@@
 
-The `r tar_read(spiked_urine_compound_info) %>% nrow()` chemical standards were dissolved in a solvent mixture of methanol, water and formic acid (70%:30%:0.1%), each 4.5 $\mu$g/ml concentration.
+The `r tar_read(standards_compound_info) %>% nrow()` chemical standards were dissolved in a solvent mixture of methanol, water and formic acid (70%:30%:0.1%), each 4.5 $\mu$g/ml concentration.
 700 $\mu$L of urine and 300 $\mu$L of the standards mixture were combined to produce the spiked urine sample.
@@
 
-FIE-HRMS mass spectra were acquired on an Exactive Orbitrap (ThermoFinnigan, San Jose CA) mass spectrometer coupled to an Accela (ThermoFinnigan, San Jose CA) ultra-performance liquid chromatography (LC) system for the *B.
-distachyon* samples and the Human urine on a Q-Exactive Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system.
+FIE-HRMS mass spectra were acquired on on a Q-Exactive Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system system for the *B.
+distachyon* samples and the Human urine.
 A sample volume of 20 $\mu$L was injected and delivered to the electrospray ionisation (ESI) source via a 'plug' flow solvent of pre-mixed HPLC grade MeOH (Fisher Scientific) and ultra-pure H~2~O (18.2 $\Omega$) at a ratio of 7:3.
@@
 
-For the *B. distachyon* samples, mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes. 
-Positive ions were acquired between 55.000 and 1000.000 *m/z*, and  63.000 and 1000.000 *m/z* for negative ions.
-Mass spectra for the spiked human urine were acquired with two scan events between 55-280 *m/z* and 270-1200 *m/z* for each of the positive and negative acquisition modes.
+The mass spectra were acquired with two scan events between 55-280 *m/z* and 270-1200 *m/z* for each of the positive and negative acquisition modes.
 For all scan events, the scan rate was 1 Hz with a mass resolution of 100,000. 

### 2023-11-08 manuscript/supplementary_tables.Rmd
reorder the columns of table S6

--- 2023-11-08 manuscript/supplementary_tables.Rmd
@@
     `transformation step assignments` = n_t_assignments,
-    `assigned features` = n_assignments,
+    `assigned \\textit{m/z}` = n_assignments,
     `molecular formulas (MF)` = n_MFs,
@@
   tidyr::spread(matrix,Value) %>% 
-  slice(c(1,4,9,2,10,3,8,6,5,7)) %>%
+  slice(c(1,4,9,2,10,3,8,6,5)) %>%
+  dplyr::select(
+    Variable,
+    `standards-mix FIE-HRMS`,
+    `spiked-urine FIE-HRMS`,
+    `brachy FIE-HRMS`,
+    `brachy C18-UHPLC-HRMS`
+  ) %>% 
   set_colnames(
@@
     " ",
-    "\\\\textit{B. distachyon} leaf" = 2,
+    "Standards mix" = 1,
     "Human urine" = 1,
-    "Standards mix" = 1),
+    "\\\\textit{B. distachyon} leaf" = 2),
     escape = FALSE)

### 2023-11-08 manuscript/manuscript.Rmd
rewrite of section 3.2 to include the chemical standards assignment results in the standards mix

@@
 
-```{r assignment-success}
-tar_load(standards_compound_PIPs)
-tar_load(spiked_urine_feature_matches)
-# tar_load(spiked_urine_unassigned)
-# tar_load(spiked_urine_assigned)
-# tar_load(spiked_urine_incorrect_assignments)
+```{r assignment-evaluations}
+tar_load(standards_parameters_molecular_formula_assignment)
+tar_load(standards_assignment_outcomes_non_iso)
+tar_load(spiked_urine_assignment_outcomes_non_iso)
+tar_load(standards_compound_outcomes)
 
-brachy_shared_MFs <- tar_read(brachy_FIE_HRMS_results_molecular_formula_assignment) %>% 
-  assignments::assignments() %>% 
-  select(MF) %>% 
-  inner_join(tar_read(brachy_RP_UHPLC_HRMS_results_molecular_formula_assignment) %>% 
-               assignments::assignments() %>% 
-               select(MF),
-             by = 'MF') %>% 
-  distinct() %>% 
+n_adducts <- standards_parameters_molecular_formula_assignment %>%
+  assignments::adducts() %>% 
+  purrr::map_dbl(length) %>% 
+  sum()
+
+n_features <- list(
+  standards = standards_assignment_outcomes_non_iso,
+  spiked_urine = spiked_urine_assignment_outcomes_non_iso
+) %>% 
+  purrr::map(
+    ~.x %>% 
+      dplyr::pull(
+        Feature
+      ) %>%
+      unique() %>%
+      length()
+  )
+
+standards_n_AI_relevant <- standards_assignment_outcomes_non_iso %>% 
+  dplyr::filter(
+    correlations == TRUE,
+    relationships == TRUE
+  ) %>% 
+  dplyr::pull(MF) %>% 
+  unique() %>% 
+  length()
+
+standards_n_AI_assigned <- standards_assignment_outcomes_non_iso %>% 
+  dplyr::filter(
+    matching_assignment == TRUE
+  ) %>% 
+  dplyr::pull(MF) %>% 
+  unique() %>% 
+  length()
+
+standards_percent_AI_assigned <- {standards_n_AI_assigned / standards_n_AI_relevant * 100} %>% 
+  signif(digits = 3)
+
+standards_n_alternative <- standards_assignment_outcomes_non_iso %>% 
+  dplyr::filter(
+    dplyr::if_all(
+      correlations:assigned,
+      ~.x == TRUE
+    ),
+    matching_adduct == FALSE
+  ) %>% 
   nrow()
 
-brachy_shared_kegg <- intersect(
-  tar_read(brachy_FIE_HRMS_KEGG_matches)$MF %>% 
-    unique(),
-  tar_read(brachy_RP_UHPLC_HRMS_KEGG_matches)$MF %>% 
-    unique()
-) %>% length()
-
-positive_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
- assignments::adducts() %>% 
-  .$p %>% 
-  length()
-
-negative_mode_adducts <- brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
- assignments::adducts() %>% 
-  .$n %>% 
-  length()
-
-spiked_urine_MF_PIPs <- standards_compound_PIPs %>% 
-  select(MF,Adduct,Isotope) %>% 
-  distinct()
-
-spiked_urine_feature_matches_MF_n <- spiked_urine_feature_matches$MF %>% 
-  unique() %>% length()
-
-# percent_unassigned <- {nrow(spiked_urine_unassigned) / 
-#     nrow(spiked_urine_feature_matches) * 100} %>% 
-#   signif(digits = 3)
-
-# spiked_urine_unassigned_MF_n <- spiked_urine_unassigned %>% 
-#   anti_join(
-#     spiked_urine_assigned %>% 
-#       select(MF) %>% 
-#       distinct(), 
-#     by = 'MF') %>% 
-#   .$MF %>% 
-#   unique() %>% 
-#   length()
-
-# spiked_urine_correct_assignments_MF_n <- spiked_urine_correct_assignments %>%
-#   .$MF %>% 
-#   unique() %>% 
-#   length()
-
-# spike_urine_standards_MF_assignment_impossible_n <- spiked_urine_feature_matches_MF_n -
-#   nrow(spike_urine_standards_MF_assignment_possible)
-
-# percent_standards_MF_assigned <- {spiked_urine_correct_assignments_MF_n / 
-#   nrow(spike_urine_standards_MF_assignment_possible) * 100} %>% 
-#   signif(digits = 3)
+compounds_percent_assigned <- standards_compound_outcomes %>% 
+  dplyr::filter(
+    outcome %in% c(
+      "Matched to m/z features",
+      "Assigned"
+    )
+  ) %>% 
+  tidyr::gather(
+    sample,n,
+    -outcome
+  ) %>% 
+  tidyr::spread(
+    outcome,n
+  ) %>% 
+  dplyr::group_by(sample) %>% 
+  dplyr::summarise(
+    percent = {Assigned / `Matched to m/z features` * 100} %>% 
+      signif(digits = 3)
+  ) %>% 
+  dplyr::pull(percent) %>% 
+  purrr::set_names(c(
+    'spiked_urine',
+    'standards'
+  )) %>% 
+  as.list()
 ```
 
-An overview of the molecular formula assignments made in both the urine and *B.
-distachyon* FIE-HRMS matrices is provided in Table \@ref(tab:assignment-results), which shows that similar proportions of *m/z* features were assigned molecular formulas.
-However, there was more than double the relative molecular formula assignment coverage for the C18-UHPLC-HRMS data of *B. distachyon* compared to the FIE-HRMS matrices.
-The chromatographic separation of the C18-UHPLC-HRMS increases the likelihood that a number of correlated ionisation products will be present for a given compound due to a reduction in the effects of ion suppression [@gowda2014].
-There is also a decreased chance of interference in the correlations between related features due to the presence of structural isomers and isobaric compounds.
-A total of `r brachy_shared_MFs` molecular formulas were assigned in both the FIE-HRMS and C18-UHPLC-HRMS data of *B. distachyon*.
+To assess the performance of the molecular formula and graphical component selection approaches, the assignment outcomes were evaluated for the theoretical ionisation products calculated for the chemical standards that matched to *m/z* features in the standards mix and spiked urine.
+The standards mix allows for a direct evaluation of the selection approaches in a relatively simple, synthetic matrix.
+Alternatively, the spiked urine provides an indication of how the selection approaches perform in a more realistic context where there are more *m/z* features present greater competition for assignment.
+To simplify interpretation, these evaluations were based on the assignment outcomes of the non-isotopic ionisation products during the adduct and isotopic assignment iterations.
+This reduces the incidence of overlap between ionisation products, where multiple theoretical ionisation products match to a single *m/z* feature as well as a removes the inclusion of relationships involving transformations that complicates the calculation and interpretation of relevant relationships.
+Based on the `r n_adducts` adduct rules specified, there were `r nrow(standards_assignment_outcomes_non_iso)` and `r nrow(spiked_urine_assignment_outcomes_non_iso)` theoretical ionisation products for these chemical standards that matched to `r n_features$standards` and `r n_features$spiked_urine` *m/z* in the standards mix and spiked urine respectively.
+The theoretical ionisation products that directly matched *m/z* features, including the isotopic theoretical ionisation products, are provided in Additional file 5.
+The frequencies of assignment outcomes for the theoretical ionisation products in these matrices are shown in Table \@ref(tab:ip-outcomes).
 
@@
 
-```{r assignment-evaluations}
+```{r ip-outcomes}
 tar_load(assignment_outcomes_summary)
@@
 
+Of the `r standards_n_AI_relevant` chemical standards that had theoretical ionisation products with relevant relationships in the standards mix, the molecular formula and graphical component selection approaches were able to correctly assign molecular formulas to `r standards_n_AI_assigned` (`r standards_percent_AI_assigned`%) of these.
+For the `r xfun::numbers_to_words(standards_n_alternative)` ionisation products for which an alternative adduct and molecular formula were assigned, two of these were assigned to alternative components with a higher $P_c$ score resulting from a higher component AIS score.
+The other alternative component was selected due to higher $P_c$ score resulting from a higher component degree.
+There was only a single instance where an ionisation product was incorrectly assigned as a result of the molecular formula selection approach.
+
+The incidence of alternative adduct and molecular formula assignment is higher in the spiked urine (Table \@ref(tab:ip-outcomes)).
+However, this is expected as there and are more *m/z* features present and greater numbers of background compounds and therefore greater chance for alternative graphical components to be present in the correlation network.
+It does not necessarily reflect a poor performance of the graphical component selection approach, there is just a higher probability that these *m/z* could potentially be alternative compounds. 
+Alternative assignment resulting from molecular formula selection still remains low.
+
+The assignment results for the chemical standards in the standards mix and spiked urine are provided in Table \@ref(tab:compound-frequencies).
+These show that, where compounds had *m/z* features matching to theoretical ionisation products, the approach was able to assign `r compounds_percent_assigned$standards`% and `r compounds_percent_assigned$spiked_urine`% of these chemical standards in the standards mix and spiked urine respectively.
+The *m/z* for which a chemical standard was assigned are provided in Additional file 6.
+
 \linespread{1}
 
-```{r compound-evaluations}
-tar_load(standards_compound_outcomes)
-
+```{r compound-frequencies}
 standards_compound_outcomes %>%
@@
 
-Although there were proportionately more molecular formulas assigned to *m/z* features in the C18-UHPLC-HRMS compared to the FIE-HRMS *B.
-distachyon* matrix, there was a disproportionate number of these molecular formulas that matched to compounds in the KEGG database specific to *B.
-distachyon* and able to form the relevant ionisation products (Table \@ref(tab:assignment-results)).
-However, only `r brachy_shared_kegg` of the molecular formulas matched to compounds in the KEGG database were found to be common between the two techniques.
-The low number of assigned molecular formulas partly reflects the incompleteness of the KEGG database, especially with respect to the secondary metabolites that are prolific in C18-UHPLC-HRMS [@breitling2013].
-The assignments for all the matrices along with the matched KEGG compound IDs are provided in Additional file 4.
+An overview of the molecular formula assignments made across the example sample matrices is provided in Table S6 (Additional file 3).
+All of the assignments for these example matrices along with any matched KEGG compound IDs are provided in Additional file 4.
+The assignments obtained from the *B.
+distachyon* is provided as an example showing the intended real world application of the molecular formula assignment approach to a sample set that includes multiple underlying experimental classes in the correlation network.
+A similar number of molecular formulas were assigned to the FIE-HRMS matrices for the spiked human urine and *B.
+distachyon* matrices as well as similar numbers of molecular formulas that were matched to KEGG compounds.
 
-To assess the assignments in the example urine matrix with respect to the spiked chemical standards, the predicted ionisation products were first identified by directly matching these to the observed *m/z*.
-Based on the `r positive_mode_adducts + negative_mode_adducts` adducts and four isotopic rules specified, there were `r nrow(spiked_urine_MF_PIPs)` predicted ionisation product *m/z* for these chemical standards.
-A total of `r nrow(spiked_urine_feature_matches)` of these ionisation products could be directly matched to the observed *m/z* features which included `r spiked_urine_feature_matches_MF_n` of the `r standards_compound_MFs` unique molecular formulas of the spiked chemical standards.
-These directly matched *m/z* features are provided in Additional file 5.
-For `r #nrow(spiked_urine_unassigned)` (`r #percent_unassigned`%) of these matched *m/z* features, no molecular formula was assigned by the approach, which included `r #spiked_urine_unassigned_MF_n` of the molecular formulas of the spiked chemical standards.
-However, `r #nrow(spiked_urine_assigned)` of the matched features were assigned a molecular formula with `r #nrow(spiked_urine_correct_assignments)` of these corresponding to one of the spiked compound ionisation products.
-This included `r #spiked_urine_correct_assignments_MF_n` of the spiked chemical standards molecular formulas (Table \@ref(tab:assignment-results)).
-The *m/z* features assigned to molecular formulas corresponding to the spiked chemical standards are provided in Additional file 6.
-
-The remaining `r #nrow(spiked_urine_incorrect_assignments)` matched *m/z* features were assigned alternative molecular formula ionisation products.
-The *m/z* of the predicted ionisation features that it was not possible to assign molecular formulas were as a result of either not being correlated with other *m/z* features that corresponded to a correct adduct or isotope mass difference or whose subgraph components were not selected, with more plausible solutions instead selected based on the correlation network.
-
-Accounting for the `r #spike_urine_standards_MF_assignment_impossible_n` chemical standards that were not assigned due to a lack of relevant correlated mass relationships, the algorithm was able to assign molecular formulas to features corresponding to `r #percent_standards_MF_assigned`% of the `r #nrow(spike_urine_standards_MF_assignment_possible)` possible chemical standards.
-This highlights the difficulty in annotating these complex biological mixtures containing hundreds of compounds, which can create many thousands of possible ionisation products.
-Interference in correlations due to the presence of isobaric compounds and MS instrument noise can weaken correlations between features, particularly in those with low abundance.
+While not the intended primary application of the approach, the C18-UPHLC-HRMS data from *B.
+distachyon* had the highest percentage of *m/z* features assigned along with the greatest number of molecular formulas.
+The chromatographic separation of the C18-UHPLC-HRMS enables the correlation network to be considered within limited retention time windows and increases the likelihood that a number of correlated ionisation products will be present for a given compound.
+Added to this, the LC-MS reduces the effects of ion suppression compare to FIE-HRMS and also decreases the chance of potential interference in the correlations between related features due to the presence of structural isomers and isobaric compounds [@gowda2014].
 
@@
 distachyon* ecotype comparison sample set provided here.
-The assignments obtained from the *B. distachyon* example matrix also show the application of the assignment approach to sample sets that contain multiple underlying experimental classes.
 

### 2023-11-08 manuscript/manuscript.Rmd
Improve the caption and row descriptions of table 2

@@
     outcome %in% c(
-      "Matched to m/z features",
-      "Assigned"
+      "MFs matched to m/z features",
+      "Total MFs assigned"
     )
@@
   dplyr::summarise(
-    percent = {Assigned / `Matched to m/z features` * 100} %>% 
+    percent = {`Total MFs assigned` / `MFs matched to m/z features` * 100} %>% 
       signif(digits = 3)
@@
     ),
-    caption = 'Molecular formula assignment results frequencies for the 64 unique molecular formulas of the chemical standards in the standards mix and spiked into human urine.'
+    caption = 'Molecular formula assignment results for the 64 unique molecular formulas (MFs) of the chemical standards in the standards mix and spiked into human urine.
+The MFs assigned are also shown for the individual adduct and isotopic (A\\&I) iterations and the transformation (T) iterations.'
   ) %>%

### 2023-11-08 R/targets/standards_targets.R

@@ -715,7 +715,7 @@ standards_figures_targets <- list(
       dplyr::distinct() %>% 
       dplyr::count(sample) %>% 
       dplyr::rename(
-        `Matched to m/z features` = n
+        `MFs matched to m/z features` = n
       ) %>% 
       dplyr::left_join(
         assignment_outcomes_non_iso %>% 
@@ -739,7 +739,7 @@ standards_figures_targets <- list(
           dplyr::distinct() %>% 
           dplyr::count(sample) %>% 
           dplyr::rename(
-            `Assigned during A&I assignment step` = n
+            `Assigned during A&I assignment iterations` = n
           ),
         by = 'sample'
       ) %>% 
@@ -761,7 +761,7 @@ standards_figures_targets <- list(
           dplyr::distinct() %>% 
           dplyr::count(sample) %>% 
           dplyr::rename(
-            `Assigned during T assignment step` = n
+            `Assigned during T assignment iterations` = n
           ),
         by = 'sample'
       ) %>% 
@@ -771,7 +771,7 @@ standards_figures_targets <- list(
           dplyr::distinct() %>% 
           dplyr::count(sample) %>% 
           dplyr::rename(
-            `Assigned` = n
+            `Total MFs assigned` = n
           ),
         by = 'sample'
       ) %>% 
@@ -788,11 +788,11 @@ standards_figures_targets <- list(
         outcome = factor(
           outcome,
           levels = c(
-            'Matched to m/z features',
+            'MFs matched to m/z features',
             'Correlations obtained',
-            'Assigned during A&I assignment step',
-            'Assigned during T assignment step',
-            'Assigned'
+            'Assigned during A&I assignment iterations',
+            'Assigned during T assignment iterations',
+            'Total MFs assigned'
           )
         )
       ) %>% 

### 2023-11-08 manuscript/manuscript.Rmd
Abbreviate ionisation products to IPs

@@
 A number of strategies can be employed to annotate peaks in mass spectra.
-These can include simple approaches such as heuristic filtering of molecular formulas generated from *m/z* or direct matching putative ionisation products to metabolite databases [@draper2009; @green2015; @kind2007];
+These can include simple approaches such as heuristic filtering of molecular formulas generated from *m/z* or direct matching putative ionisation products (IPs) to metabolite databases [@draper2009; @green2015; @kind2007];
 but also ensemble approaches that integrate a number of strategies like correlation analysis, ionisation product scoring and biotransformations [@baran2013; @fernandez2014; @kuhl2012; @uppal2017].
@@
 
-Features matching to possible ionisation products of the chemical standards were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
+Features matching to possible IPs of the chemical standards were identified by first applying the MZedDB ionisation 'rules' and calculating their theoretical *m/z* for each of the adducts specified in Section \@ref(relationships) [@draper2009].
 The calculated *m/z* were then matched to the *m/z* features in the standards mix and spiked urine data using a search range of $\pm$ `r search_ppm` ppm.
@@
 
-To assess the performance of the molecular formula and graphical component selection approaches, the assignment outcomes were evaluated for the theoretical ionisation products calculated for the chemical standards that matched to *m/z* features in the standards mix and spiked urine.
+To assess the performance of the molecular formula and graphical component selection approaches, the assignment outcomes were evaluated for the theoretical IPs calculated for the chemical standards that matched to *m/z* features in the standards mix and spiked urine.
 The standards mix allows for a direct evaluation of the selection approaches in a relatively simple, synthetic matrix.
 Alternatively, the spiked urine provides an indication of how the selection approaches perform in a more realistic context where there are more *m/z* features present greater competition for assignment.
-To simplify interpretation, these evaluations were based on the assignment outcomes of the non-isotopic ionisation products during the adduct and isotopic assignment iterations.
-This reduces the incidence of overlap between ionisation products, where multiple theoretical ionisation products match to a single *m/z* feature as well as a removes the inclusion of relationships involving transformations that complicates the calculation and interpretation of relevant relationships.
-Based on the `r n_adducts` adduct rules specified, there were `r nrow(standards_assignment_outcomes_non_iso)` and `r nrow(spiked_urine_assignment_outcomes_non_iso)` theoretical ionisation products for these chemical standards that matched to `r n_features$standards` and `r n_features$spiked_urine` *m/z* in the standards mix and spiked urine respectively.
-The theoretical ionisation products that directly matched *m/z* features, including the isotopic theoretical ionisation products, are provided in Additional file 5.
-The frequencies of assignment outcomes for the theoretical ionisation products in these matrices are shown in Table \@ref(tab:ip-outcomes).
+To simplify interpretation, these evaluations were based on the assignment outcomes of the non-isotopic IPs during the adduct and isotopic assignment iterations.
+This reduces the incidence of overlap between IPs, where multiple theoretical IPs match to a single *m/z* feature as well as a removes the inclusion of relationships involving transformations that complicates the calculation and interpretation of relevant relationships.
+Based on the `r n_adducts` adduct rules specified, there were `r nrow(standards_assignment_outcomes_non_iso)` and `r nrow(spiked_urine_assignment_outcomes_non_iso)` theoretical IPs for these chemical standards that matched to `r n_features$standards` and `r n_features$spiked_urine` *m/z* in the standards mix and spiked urine respectively.
+The theoretical IPs that directly matched *m/z* features, including the isotopic theoretical IPs, are provided in Additional file 5.
+The frequencies of assignment outcomes for the theoretical IPs in these matrices are shown in Table \@ref(tab:ip-outcomes).
 
@@
 
-Of the `r standards_n_AI_relevant` chemical standards that had theoretical ionisation products with relevant relationships in the standards mix, the molecular formula and graphical component selection approaches were able to correctly assign molecular formulas to `r standards_n_AI_assigned` (`r standards_percent_AI_assigned`%) of these.
-For the `r xfun::numbers_to_words(standards_n_alternative)` ionisation products for which an alternative adduct and molecular formula were assigned, two of these were assigned to alternative components with a higher $P_c$ score resulting from a higher component AIS score.
+Of the `r standards_n_AI_relevant` chemical standards that had theoretical IPs with relevant relationships in the standards mix, the molecular formula and graphical component selection approaches were able to correctly assign molecular formulas to `r standards_n_AI_assigned` (`r standards_percent_AI_assigned`%) of these.
+For the `r xfun::numbers_to_words(standards_n_alternative)` IPs for which an alternative adduct and molecular formula were assigned, two of these were assigned to alternative components with a higher $P_c$ score resulting from a higher component AIS score.
 The other alternative component was selected due to higher $P_c$ score resulting from a higher component degree.
@@
 The assignment results for the chemical standards in the standards mix and spiked urine are provided in Table \@ref(tab:compound-frequencies).
-These show that, where compounds had *m/z* features matching to theoretical ionisation products, the approach was able to assign `r compounds_percent_assigned$standards`% and `r compounds_percent_assigned$spiked_urine`% of these chemical standards in the standards mix and spiked urine respectively.
+These show that, where compounds had *m/z* features matching to theoretical IPs, the approach was able to assign `r compounds_percent_assigned$standards`% and `r compounds_percent_assigned$spiked_urine`% of these chemical standards in the standards mix and spiked urine respectively.
 The *m/z* for which a chemical standard was assigned are provided in Additional file 6.
@@
 distachyon* had the highest percentage of *m/z* features assigned along with the greatest number of molecular formulas.
-The chromatographic separation of the C18-UHPLC-HRMS enables the correlation network to be considered within limited retention time windows and increases the likelihood that a number of correlated ionisation products will be present for a given compound.
+The chromatographic separation of the C18-UHPLC-HRMS enables the correlation network to be considered within limited retention time windows and increases the likelihood that a number of correlated IPs will be present for a given compound.
 Added to this, the LC-MS reduces the effects of ion suppression compare to FIE-HRMS and also decreases the chance of potential interference in the correlations between related features due to the presence of structural isomers and isobaric compounds [@gowda2014].
@@
 
-Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted ionisation products could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS metabolome fingerprinting.
+Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted IPs could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS metabolome fingerprinting.
 It was also possible to assign molecular formulas of common primary and secondary metabolites present in the KEGG compound database to *m/z* features from leaf tissue of *B.

### 2023-11-08 manuscript/additional_files.Rmd
re-number additional files 4-6 inline with the restructuring of section 3.2

--- 2023-11-06 manuscript/additional_files.Rmd
@@
 
-Additional file 4.csv - Table S7 -
-The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.
+Additional file 4.csv Table S7 -
+The *m/z* features directly matched to theoretical ionisation products of the chemical standards in the standards mixed and spiked urine.
 
 Additional file 5.csv Table S8 -
-The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.
+The *m/z* features assigned to chemical standards in the standards mix and spiked urine.
 
-Additional file 6.csv Table S9 -
-The *m/z* features assigned to spiked chemical compounds in urine.
+Additional file 6.csv - Table S9 -
+The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.

### 2023-11-08 manuscript/manuscript.Rmd
re-number additional files 4-6 inline with the restructuring of section 3.2

@@
 Based on the `r n_adducts` adduct rules specified, there were `r nrow(standards_assignment_outcomes_non_iso)` and `r nrow(spiked_urine_assignment_outcomes_non_iso)` theoretical IPs for these chemical standards that matched to `r n_features$standards` and `r n_features$spiked_urine` *m/z* in the standards mix and spiked urine respectively.
-The theoretical IPs that directly matched *m/z* features, including the isotopic theoretical IPs, are provided in Additional file 5.
+The theoretical IPs that directly matched *m/z* features, including the isotopic theoretical IPs, are provided in Additional file 4.
 The frequencies of assignment outcomes for the theoretical IPs in these matrices are shown in Table \@ref(tab:ip-outcomes).
@@
 These show that, where compounds had *m/z* features matching to theoretical IPs, the approach was able to assign `r compounds_percent_assigned$standards`% and `r compounds_percent_assigned$spiked_urine`% of these chemical standards in the standards mix and spiked urine respectively.
-The *m/z* for which a chemical standard was assigned are provided in Additional file 6.
+The *m/z* for which a chemical standard was assigned are provided in Additional file 5.
 
@@
 An overview of the molecular formula assignments made across the example sample matrices is provided in Table S6 (Additional file 3).
-All of the assignments for these example matrices along with any matched KEGG compound IDs are provided in Additional file 4.
+All of the assignments for these example matrices along with any matched KEGG compound IDs are provided in Additional file 6.
 The assignments obtained from the *B.

### 2023-11-08 R/targets/manuscript_targets.R

@@ -42,9 +42,57 @@ manuscript_targets <- list(
         file = 'exports/manuscript/supplementary_materials/Additional file 1.csv')
   ),
   
-  ## Supplementary table S5 of all assignments
+  ## Supplementary table S7 directly matched chemical standard ionisation products
   tarchetypes::tar_file(
     table_s7,
+    list(
+      `Standards mix` = standards_feature_matches,
+      `Spiked urine` = spiked_urine_feature_matches
+    ) %>% 
+      dplyr::bind_rows(
+        .id = 'Matrix'
+      ) %>% 
+      select(
+        Matrix,
+        MF,
+        Isotope,
+        Adduct,
+        `m/z` = mz,
+        `Ionisation mode` = Mode,
+        `PPM error`
+      ) %>% 
+      jfmisc::exportCSV(
+        file = 'exports/manuscript/supplementary_materials/Additional file 4.csv'
+      )
+  ),
+  
+  ## Supplementary table S8 assigned chemical standards
+  tarchetypes::tar_file(
+    table_s8,
+    list(
+      `Standards mix` = standards_correct_assignments,
+      `Spiked urine` = spiked_urine_correct_assignments
+    ) %>%
+      dplyr::bind_rows(
+        .id = 'Matrix'
+      ) %>% 
+      select(
+        Matrix,
+        MF,
+        Isotope,
+        Adduct,
+        `m/z` = `Measured m/z`,
+        `Ionisation mode` = Mode,
+        `PPM error`
+      ) %>% 
+      jfmisc::exportCSV(
+        file = 'exports/manuscript/supplementary_materials/Additional file 5.csv'
+      )
+  ),
+  
+  ## Supplementary table S9 of all assignments
+  tarchetypes::tar_file(
+    table_s9,
     bind_rows(
       list(
         `NA_Standards mix_FIE-HRMS` = standards_results_molecular_formula_assignment %>% 
@@ -107,54 +155,6 @@ manuscript_targets <- list(
              `PPM error`,
              `MF Plausibility (%)`,
              `KEGG ID`) %>% 
-      jfmisc::exportCSV(
-        file = 'exports/manuscript/supplementary_materials/Additional file 4.csv'
-      )
-  ),
-  
-  ## Supplementary table S6 directly matched chemical standard ionisation products
-  tarchetypes::tar_file(
-    table_s8,
-    list(
-      `Standards mix` = standards_feature_matches,
[…]

### 2023-11-08 R/targets/manuscript_targets.R
Correct the format of the 37Cl and 41K adducts in additional files 4-6

@@ -60,6 +60,17 @@ manuscript_targets <- list(
         `m/z` = mz,
         `Ionisation mode` = Mode,
         `PPM error`
+      ) %>%
+      dplyr::mutate(
+        Adduct = Adduct %>% 
+          replace(
+            Adduct == '[M+Cl37]1-',
+            '[M+37Cl]1-'
+          ) %>% 
+          replace(
+            Adduct == '[M+K41]1+',
+            '[M+41K]1+'
+          )
       ) %>% 
       jfmisc::exportCSV(
         file = 'exports/manuscript/supplementary_materials/Additional file 4.csv'
@@ -84,6 +95,17 @@ manuscript_targets <- list(
         `m/z` = `Measured m/z`,
         `Ionisation mode` = Mode,
         `PPM error`
+      ) %>% 
+      dplyr::mutate(
+        Adduct = Adduct %>% 
+          replace(
+            Adduct == '[M+Cl37]1-',
+            '[M+37Cl]1-'
+          ) %>% 
+          replace(
+            Adduct == '[M+K41]1+',
+            '[M+41K]1+'
+          )
       ) %>% 
       jfmisc::exportCSV(
         file = 'exports/manuscript/supplementary_materials/Additional file 5.csv'
@@ -155,6 +177,17 @@ manuscript_targets <- list(
              `PPM error`,
              `MF Plausibility (%)`,
              `KEGG ID`) %>% 
+      dplyr::mutate(
+        Adduct = Adduct %>% 
+          replace(
+            Adduct == '[M+Cl37]1-',
+            '[M+37Cl]1-'
+          ) %>% 
+          replace(
+            Adduct == '[M+K41]1+',
+            '[M+41K]1+'
+          )
+      ) %>% 
       jfmisc::exportCSV(
         file = 'exports/manuscript/supplementary_materials/Additional file 6.csv'
       )

### 2023-11-08 manuscript/supplementary_tables.Rmd
Correctly format the adducts in table S2

@@
   ) %>%
-  mutate(`Mass difference` = round(`Mass difference`,5)) %>% 
+  mutate(
+    Adduct = Adduct %>% 
+      stringr::str_replace_all(']1-',']$^{1-}$') %>% 
+      stringr::str_replace_all(']2-',']$^{2-}$') %>% 
+      stringr::str_replace_all('Cl37','$^{37}$Cl') %>% 
+      stringr::str_replace_all(stringr::coll(']1+'),']$^{1+}$') %>% 
+      stringr::str_replace_all(stringr::coll(']2+'),']$^{2+}$') %>% 
+      stringr::str_replace_all('K41','$^{41}$K'),
+    `Mass difference` = round(`Mass difference`,5)
+    ) %>% 
   knitr::kable(caption = 'Information for positive and negative mode adducts used for calculating ionisation products.',
                format = 'latex',
+               escape = FALSE,
                booktabs = TRUE)

### 2023-11-08 manuscript/additional_files.Rmd
Move table S6 back into the main manuscript as table 3

--- 2023-11-06 manuscript/additional_files.Rmd
@@
 Additional file 3.pdf - Supplementary tables -
-Supplementary tables S2, S3, S4, S5, S6 showing adduct, isotopic, transformation, R package benchmarking information and a summary of the molecular formula assignment results for the example sample matrices respectively.
+Supplementary tables S2, S3, S4, S5, showing adduct, isotopic, transformation and R package benchmarking information respectively.
 
-Additional file 4.csv - Table S7 -
+Additional file 4.csv Table S6 -
+The *m/z* features directly matched to theoretical ionisation products of the chemical standards in the standards mixed and spiked urine.
+
+Additional file 5.csv Table S7 -
+The *m/z* features assigned to chemical standards in the standards mix and spiked urine.
+
+Additional file 6.csv - Table S8 -
 The molecular formula assignments for *m/z* across all experimental matrices along with any match KEGG compound IDs.
-
-Additional file 5.csv Table S8 -
-The *m/z* features directly matched to predicted ionisation products of the spiked chemical compounds in urine.
-
-Additional file 6.csv Table S9 -
-The *m/z* features assigned to spiked chemical compounds in urine.

### 2023-11-08 manuscript/manuscript.Rmd
Move table S6 back into the main manuscript as table 3

@@
 
-An overview of the molecular formula assignments made across the example sample matrices is provided in Table S6 (Additional file 3).
+An overview of the molecular formula assignments made across the example sample matrices is provided in Table \@ref(tab:assignment-results).
 All of the assignments for these example matrices along with any matched KEGG compound IDs are provided in Additional file 6.
@@
 Added to this, the LC-MS reduces the effects of ion suppression compare to FIE-HRMS and also decreases the chance of potential interference in the correlations between related features due to the presence of structural isomers and isobaric compounds [@gowda2014].
+
+\linespread{1}
+
+```{r assignment-results}
+tar_read(assignment_results) %>%
+  dplyr::full_join(
+    tar_read(assignment_matches) %>% 
+      dplyr::mutate(
+        matrix = matrix %>% 
+          replace(
+            matrix == 'standards',
+            'standards-mix'
+          )
+      ), 
+    by = c("matrix", "technique")) %>% 
+  mutate(
+    matrix = paste(matrix,technique)
+  ) %>%
+  mutate(
+    n_assignments = glue::glue(
+      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
+    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
+  select(-technique) %>% 
+  dplyr::rename(
+    `\\textit{m/z} features` = n_features,
+    correlations = n_correlations,
+    relationships = n_relationships,
+    `adduct and isotopic step assignments` = n_ai_assignments,
+    `transformation step assignments` = n_t_assignments,
+    `assigned \\textit{m/z}` = n_assignments,
+    `molecular formulas (MF)` = n_MFs,
+    `KEGG compound matches` = KEGG_compound_matches,
+    `KEGG MF matches` = KEGG_MF_matches,
+    `matched chemical standards` = matched_standards) %>% 
+  tidyr::gather(Variable,Value,-matrix) %>% 
+  mutate(
+    Value = replace(Value,
+                    Value == 'NA',
+                    'N/A')) %>% 
+  tidyr::spread(matrix,Value) %>% 
+  slice(c(1,4,9,2,10,3,8,6,5)) %>%
+  dplyr::select(
+    Variable,
+    `standards-mix FIE-HRMS`,
+    `spiked-urine FIE-HRMS`,
+    `brachy FIE-HRMS`,
+    `brachy C18-UHPLC-HRMS`
+  ) %>% 
+  set_colnames(
+    stringr::str_split_fixed(colnames(.),
+                             ' ',2)[,2]) %>% 
+  knitr::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    linesep = '',
+    escape = FALSE,
+    align = c(
+      rep('l',1),
+      rep('r',4)
+    ),
+    caption = 'A summary of the molecular formula assignment results for the example sample matrices.') %>%
+  kable_styling(
+    font_size = 10,
+    latex_options = "HOLD_position"
+  ) %>%
+  add_header_above(c(
+    " ",
+    "Standards mix" = 1,
+    "Human urine" = 1,
+    "\\\\textit{B. distachyon} leaf" = 2),
+    escape = FALSE)
+```
+
+\linespread{2}
 

### 2023-11-08 manuscript/supplementary_tables.Rmd
Move table S6 back into the main manuscript as table 3

@@
 ```
-
-```{r table-s6}
-tar_read(assignment_results) %>%
-  dplyr::full_join(
-    tar_read(assignment_matches) %>% 
-      dplyr::mutate(
-        matrix = matrix %>% 
-          replace(
-            matrix == 'standards',
-            'standards-mix'
-          )
-      ), 
-    by = c("matrix", "technique")) %>% 
-  mutate(
-    matrix = paste(matrix,technique)
-  ) %>%
-  mutate(
-    n_assignments = glue::glue(
-      '{n_assignments} ({signif(n_assignments/n_features * 100,digits = 3)}\\%)'),
-    across(where(is.numeric),~prettyNum(.x,big.mark = ','))) %>%
-  select(-technique) %>% 
-  dplyr::rename(
-    `\\textit{m/z} features` = n_features,
-    correlations = n_correlations,
-    relationships = n_relationships,
-    `adduct and isotopic step assignments` = n_ai_assignments,
-    `transformation step assignments` = n_t_assignments,
-    `assigned \\textit{m/z}` = n_assignments,
-    `molecular formulas (MF)` = n_MFs,
-    `KEGG compound matches` = KEGG_compound_matches,
-    `KEGG MF matches` = KEGG_MF_matches,
-    `matched chemical standards` = matched_standards) %>% 
-  tidyr::gather(Variable,Value,-matrix) %>% 
-  mutate(
-    Value = replace(Value,
-                    Value == 'NA',
-                    'N/A')) %>% 
-  tidyr::spread(matrix,Value) %>% 
-  slice(c(1,4,9,2,10,3,8,6,5)) %>%
-  dplyr::select(
-    Variable,
-    `standards-mix FIE-HRMS`,
-    `spiked-urine FIE-HRMS`,
-    `brachy FIE-HRMS`,
-    `brachy C18-UHPLC-HRMS`
-  ) %>% 
-  set_colnames(
-    stringr::str_split_fixed(colnames(.),
-                             ' ',2)[,2]) %>% 
-  knitr::kable(
-    format = 'latex',
-    booktabs = TRUE,
-    linesep = '',
-    escape = FALSE,
-    align = c(
-      rep('l',1),
-      rep('r',3)
-    ),
-    caption = 'A summary of the molecular formula assignment results for the example sample matrices.
-The number of matched chemical standards is given as the number of unique molecular formulas.') %>%
-  kable_styling() %>%
-  add_header_above(c(
-    " ",
-    "Standards mix" = 1,
-    "Human urine" = 1,
-    "\\\\textit{B. distachyon} leaf" = 2),
-    escape = FALSE)
-```

### 2023-11-08 R/targets/manuscript_targets.R

@@ -42,9 +42,9 @@ manuscript_targets <- list(
         file = 'exports/manuscript/supplementary_materials/Additional file 1.csv')
   ),
   
-  ## Supplementary table S7 directly matched chemical standard ionisation products
+  ## Supplementary table S6 directly matched chemical standard ionisation products
   tarchetypes::tar_file(
-    table_s7,
+    table_s6,
     list(
       `Standards mix` = standards_feature_matches,
       `Spiked urine` = spiked_urine_feature_matches
@@ -77,9 +77,9 @@ manuscript_targets <- list(
       )
   ),
   
-  ## Supplementary table S8 assigned chemical standards
+  ## Supplementary table S7 assigned chemical standards
   tarchetypes::tar_file(
-    table_s8,
+    table_s7,
     list(
       `Standards mix` = standards_correct_assignments,
       `Spiked urine` = spiked_urine_correct_assignments
@@ -112,9 +112,9 @@ manuscript_targets <- list(
       )
   ),
   
-  ## Supplementary table S9 of all assignments
+  ## Supplementary table S8 of all assignments
   tarchetypes::tar_file(
-    table_s9,
+    table_s8,
     bind_rows(
       list(
         `NA_Standards mix_FIE-HRMS` = standards_results_molecular_formula_assignment %>% 
@@ -208,9 +208,9 @@ manuscript_targets <- list(
           c(supplementary_methods[1],
             supplementary_tables[1],
             table_s1,
+            table_s6,
             table_s7,
-            table_s8,
-            table_s9),
+            table_s8),
           flags = '-r9Xj')
       
       zipfile

### 2023-11-09 R/targets/MF_selection_targets.R
Set the ppm threshold to 4 for the KEGG compound MF ranking

@@ -1,5 +1,5 @@
 
-KEGG_alternative_MFs_ppm <- 6
+KEGG_alternative_MFs_ppm <- 4
 
 MF_selection_targets <- list(
   

### 2023-11-09 R/targets/standards_targets.R

@@ -1,6 +1,4 @@
 
-search_ppm <- 4
-
 assignment_parameters <- {
   p <- assignments::assignmentParameters("FIE-HRMS") 
   assignments::adducts(p)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")

### 2023-11-09 manuscript/manuscript.Rmd
Correct the figure 2 caption

@@
 ```{r figure-2-kegg-mf-selection,fig.height=4,fig.cap='Molecular formula plausibility ($P_{MF}$) rankings of KEGG compound molecular formulas.
-The $P_{MF}$ rankings were based on the monoisotopic masses of 6745 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 6 ppm.'}
+The $P_{MF}$ rankings were based on the monoisotopic masses of 6745 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 4 ppm.'}
 tar_read(KEGG_compound_MF_ranks_mass_plot) +

### 2023-11-09 manuscript/manuscript.Rmd
Update the assignment parameters in the methods section

@@
   str_as_written_list()
+
 positive_mode_adducts <-   brachy_FIE_HRMS_parameters_molecular_formula_assignment %>% 
+  assignments::adducts() %>% 
+  .$p %>% 
+  stringr::str_replace_all(stringr::coll(']1+'),']^1+^') %>% 
+  stringr::str_replace_all(stringr::coll(']2+'),']^2+^') %>% 
+  stringr::str_replace_all('K41','^41^K') %>% 
+  str_as_written_list()
+
+standards_pos_adducts <-   standards_parameters_molecular_formula_assignment %>% 
   assignments::adducts() %>% 
@@
 All adduct mass, charge and molecule information can be found in Additional file 3 (Table S2).
-The isotopes included single and double ^13^C and ^18^O for the *B.distachyon* matrices and additionally ^34^S for the spiked urine.
+The isotopes included ^13^C and ^18^O for the *B.distachyon* matrices and additionally ^34^S for the standards mix and spiked urine.
 The isotopic mass information used can be found in Additional file 3 (Table S3).
@@
 The calculated frequencies were rounded to the nearest integer value.
-Molecular formula generation was conducted for each $M$ value within a mass deviation range of $\pm$ 6 ppm.
+Molecular formula generation was conducted for each $M$ value within a mass deviation range of $\pm$ `r search_ppm` ppm.
 This generation was exhaustive, using the brute force approach of the HR2 generator, without the application of the golden rules proposed by @kind2007.
@@
 To calculate $AIS$, the possible adducts within each ionisation mode were ranked from most probable to least probable and assigned a score ($a$) with the most probable being $n-1$, $n$ being the number of adducts within that ionisation mode and least probable adduct a score of 0.
-The adduct rank scores used for the assignment were based on the adduct order specified in Section \@ref(relationships) for all matrices.
+The adduct rank scores used for the assignment were based on the adduct order specified in Section \@ref(relationships) for the *B.
+distachyon* matrices.
+For the standards mix and spiked urine, an alternative order of `r standards_pos_adducts` was used for the positive mode adducts.
 
 Similarly, for the possible isotopes, these were also ranked from most (non-isotopic) to least probable and assigned a score ($i$) as for the adducts.
-The isotopic rank scores used for the assignment of the *B.distachyon* matrices were as follows: Non-isotopic = 3, ^13^C = 2, ^18^O = 1, ^13^C2 = 0.
-For the spiked urine matrices, the isotopic rank scores were: Non-isotopic = 4, ^13^C = 3, ^18^O = 2, ^34^S = 1, ^13^C2 = 0.
+The isotopic rank scores used for the assignment of the *B.distachyon* matrices were as follows: Non-isotopic = 2, ^13^C = 1, ^18^O = 0.
+For the standards mix and spiked urine matrices, the isotopic rank scores were: Non-isotopic = 3, ^13^C = 2, ^18^O = 1, ^34^S = 0.
 

### 2023-11-09 R/targets/standards_targets.R

@@ -1,10 +1,12 @@
 
+search_ppm <- 4 
+
 assignment_parameters <- {
   p <- assignments::assignmentParameters("FIE-HRMS") 
   assignments::adducts(p)$n <- c("[M-H]1-","[M+Cl]1-","[M+Cl37]1-","[M+K-2H]1-","[M-2H]2-","[2M-H]1-")
   assignments::adducts(p)$p <- c("[M+Na]1+","[M+H]1+","[M+K]1+","[M+K41]1+","[M+2H]2+","[2M+H]1+")
   assignments::isotopes(p) <- c("13C","18O","34S")
-  assignments::ppm(p) <- 4
+  assignments::ppm(p) <- search_ppm
   p
 }
 

### 2023-11-09 manuscript/manuscript.Rmd
update the methods section to include the bug fixes in assignments version 1.0.1

@@
 The $AIS$ scores within each component were totalled and divided by half the total number of adduct and isotope combinations to give a component $AIS$ score ($AIS_c$).
+If an *m/z* feature was represented by more than one adduct and isotope combination in a component, the combination with the highest AIS score was retained.
 
@@
 $$P_c = d \times AIS_c \times w$$
-Where $d$ is the component degree and $w$ is the average component edge weight. 
+Where $d$ is the component degree and $w$ is the average absolute component edge weight. 
 The higher $P_c$, the more plausible the component subgraph.

### 2023-11-09 manuscript/manuscript.Rmd
Update the abstract to include the results from the standards mix

@@
 ```{r assignment-summary}
-# tar_load(assignment_results)
-# tar_load(assignment_matches)
-# tar_load(spike_urine_standards_MF_assignment_possible)
-# tar_load(spiked_urine_correct_assignments)
-# 
-# total_assigned <- round(sum(assignment_results$n_assignments) / 
-#                           sum(assignment_results$n_features) * 100,1)
-# total_mfs <- sum(assignment_results$n_MFs)
-# total_kegg <- sum(assignment_matches$KEGG_compound_matches)
-# 
-# spiked_urine_possible <- nrow(spike_urine_standards_MF_assignment_possible)
-# spiked_urine_correct_percent <- spiked_urine_correct_assignments$MF %>% 
-#   unique() %>% 
-#   length() %>% 
-#   {. / spiked_urine_possible * 100} %>% 
-#   round(digits = 1)
+tar_load(standards_compound_outcomes)
+tar_load(assignment_results)
+tar_load(assignment_matches)
+
+compounds_possible <- standards_compound_outcomes %>% 
+  dplyr::filter(
+    outcome == 'MFs matched to m/z features'
+  ) %>% 
+  dplyr::pull(
+    standards
+  )
+
+compounds_percent_assigned <- standards_compound_outcomes %>% 
+  dplyr::filter(
+    outcome %in% c(
+      "MFs matched to m/z features",
+      "Total MFs assigned"
+    )
+  ) %>% 
+  tidyr::gather(
+    sample,n,
+    -outcome
+  ) %>% 
+  tidyr::spread(
+    outcome,n
+  ) %>% 
+  dplyr::group_by(sample) %>% 
+  dplyr::summarise(
+    percent = {`Total MFs assigned` / `MFs matched to m/z features` * 100} %>% 
+      signif(digits = 3)
+  ) %>% 
+  dplyr::pull(percent) %>% 
+  purrr::set_names(c(
+    'spiked_urine',
+    'standards'
+  )) %>% 
+  as.list()
+
+average_percent_assigned <- assignment_results %>% 
+  dplyr::mutate(
+    percent_assigned = n_assignments / n_features * 100
+  ) %>% 
+  dplyr::pull(percent_assigned) %>% 
+  mean() %>% 
+  signif(digits = 3)
+
+average_kegg_compounds <- assignment_matches %>% 
+  dplyr::pull(KEGG_compound_matches) %>% 
+  mean(na.rm = TRUE) %>% 
+  round()
 ```
@@
 The R package *assignments* was developed as a user-friendly implementation of the approach and is freely available from https://github.com/aberHRML/assignments.
-The approach was tested in human urine samples spiked with known chemical standards and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
-Across the three test matrices, `r #total_assigned`% of the m/z features were assigned a molecular formula giving a total of `r #total_mfs` molecular formulas and matching to `r #total_kegg` compounds in the KEGG compound database.
-Of the `r #spiked_urine_possible` spiked chemical standards that it was possible to assign in the human urine samples, the approach was able to assign m/z features to `r #spiked_urine_correct_percent`% of these.
+The approach was tested in a mix of known chemical standards, human urine samples spiked the chemical standards mix and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
+Of the `r compounds_possible` chemical standards that had theoretical ionisation products matching *m/z* features in the standards mix, the approach assigned `r compounds_percent_assigned$standards`% of these.
+Across the example matrices, an average of `r average_percent_assigned`% of the *m/z* features were assigned a molecular formula matching an average of `r average_kegg_compounds` compounds in the KEGG compound database.
 
@@
 tar_load(spiked_urine_assignment_outcomes_non_iso)
-tar_load(standards_compound_outcomes)
 
@@
   nrow()
-
-compounds_percent_assigned <- standards_compound_outcomes %>% 
-  dplyr::filter(
-    outcome %in% c(
-      "MFs matched to m/z features",
-      "Total MFs assigned"
-    )
-  ) %>% 
-  tidyr::gather(
-    sample,n,
-    -outcome
-  ) %>% 
-  tidyr::spread(
-    outcome,n
-  ) %>% 
-  dplyr::group_by(sample) %>% 
-  dplyr::summarise(
-    percent = {`Total MFs assigned` / `MFs matched to m/z features` * 100} %>% 
-      signif(digits = 3)
-  ) %>% 
-  dplyr::pull(percent) %>% 
-  purrr::set_names(c(
-    'spiked_urine',
-    'standards'
-  )) %>% 
-  as.list()
 ```
@@
 ```{r assignment-results}
-tar_read(assignment_results) %>%
+assignment_results %>%
   dplyr::full_join(
-    tar_read(assignment_matches) %>% 
+    assignment_matches %>% 
       dplyr::mutate(

### 2023-11-09 manuscript/manuscript.Rmd
fix wording in abstract

@@
 The R package *assignments* was developed as a user-friendly implementation of the approach and is freely available from https://github.com/aberHRML/assignments.
-The approach was tested in a mix of known chemical standards, human urine samples spiked the chemical standards mix and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
+The approach was tested in a mix of known chemical standards, human urine samples spiked with the chemical standards mix and a data set comparing leaf tissue samples of different ecotypes of the model grass species *Brachypodium distachyon*.
 Of the `r compounds_possible` chemical standards that had theoretical ionisation products matching *m/z* features in the standards mix, the approach assigned `r compounds_percent_assigned$standards`% of these.

### 2023-11-09 manuscript/manuscript.Rmd
Add the term tandem MS
Replace 'LC-MS techniques' with 'LC-MS derived data'

@@
 Tandem MS can also be employed for LC-MS approaches and provides a convenient method for structural validation [@neumann2010].
-However, this is unsuitable for metabolome fingerprinting-based techniques due to the presence of structural isomers and isobaric compounds.
-Many of the available software packages implementing these strategies are suitable for LC-MS techniques only.
+However, tandem MS is unsuitable for metabolome fingerprinting-based techniques due to the presence of structural isomers and isobaric compounds.
+Many of the available software packages implementing the strategies outlined above are suitable for LC-MS derived data only.
 

### 2023-11-09 manuscript/manuscript.Rmd
replace 'interpret' with 'deduce'

@@
 
-For investigators, manual consideration of putative annotations is time-consuming and often difficult to interpret due to the complex interactions between ions and molecules and their mathematical possibilities forming *m/z* features.
+For investigators, manual consideration of putative annotations is time-consuming and often difficult to deduce due to the complex interactions between ions and molecules and their mathematical possibilities forming *m/z* features.
 It is important that software package implementations can automate this process but also provide tractable results and simple input parameters for easy and wide applicability.

### 2023-11-09 manuscript/manuscript.Rmd
replace 'tractable' with 'traceable'

@@
 For investigators, manual consideration of putative annotations is time-consuming and often difficult to deduce due to the complex interactions between ions and molecules and their mathematical possibilities forming *m/z* features.
-It is important that software package implementations can automate this process but also provide tractable results and simple input parameters for easy and wide applicability.
+It is important that software package implementations can automate this process but also provide traceable results and simple input parameters for easy and wide applicability.
 

### 2023-11-09 manuscript/manuscript.Rmd
clarify sentence about use of correlations

@@
 
-The use of correlations to associate relationships between *m/z* features provides an essential layer of data-driven, internal validation of the mathematically derived solutions.
+The use of correlations to identify relationships between *m/z* features provides an essential layer of data-driven, internal validation of theoretically derived associations.
 For instance, if two features that are posited to have the same composition (eg.

### 2023-11-09 manuscript/manuscript.Rmd
clarify replicated classes

@@
 Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS fingerprinting and LC-MS profiling-based metabolomics data.
-This is aimed at biological experiments that include numerous replicates of sample classes for the assignment of low accurate mass (< 1000 atomic mass units (amu)) metabolite elemental compositions of biological origin.
+This is aimed at biological experiments that includes replicated samples (> 4) of experimental classes for the assignment of low accurate mass (< 1000 atomic mass units (amu)) metabolite elemental compositions of biological origin.
 It is an ensemble approach that combines a number of annotation strategies including correlation analysis, molecular formula selection, adduct and isotopic scoring and graphical component selection.

### 2023-11-09 manuscript/manuscript.Rmd
correct the use of the phrase 'low accurate mass'

--- 2023-11-09 manuscript/manuscript.Rmd
@@
 Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS fingerprinting and LC-MS profiling-based metabolomics data.
-This is aimed at biological experiments that include numerous replicates of sample classes for the assignment of low accurate mass (< 1000 atomic mass units (amu)) metabolite elemental compositions of biological origin.
+This is aimed at biological experiments that include experimental classes with greater than four replicate samples and for the assignment of accurate masses up to 1000 atomic mass units (amu).
 It is an ensemble approach that combines a number of annotation strategies including correlation analysis, molecular formula selection, adduct and isotopic scoring and graphical component selection.

### 2023-11-09 manuscript/manuscript.Rmd
add 'biological origin' to the abstract
replace 'pre-ionisation' with 'precursor'

@@
 We present a data-driven ensemble approach for the ‘first pass’ automated assignment of molecular formulas to mass spectra acquired from high-resolution electrospray ionisation mass spectrometry (ESI-MS).
-This is primarily intended for metabolomics experiment data using non-targeted metabolome fingerprinting and LC-MS profiling.
+This is primarily intended for metabolomics experiment data of biological origin, using non-targeted metabolome fingerprinting and LC-MS profiling.
 The approach incorporates the use of *m/z* correlation networks, molecular formula generation and selection as well as graphically based component plausibility selection to assign molecular formulas to *m/z* features.
@@
 Possible adduct, isotopic and transformation mass relationships were calculated for each pair of correlated *m/z*.
-The pre-ionisation masses ($M$) were first calculated for all adduct, isotopic and transformation combinations for each of the *m/z*. 
+The precursor masses ($M$) were first calculated for all adduct, isotopic and transformation combinations for each of the *m/z*. 
 An $M$ for a given ionisation product *m/z*, adduct, isotope and transformation was calculated using the following equation:

### 2023-11-09 manuscript/manuscript.Rmd
fix sentence grammar for Figure 3 reference

@@
 
-As shown in Figure \@ref(fig:figure-3-feature-solutions), there can be multiple graphical molecular formula component solutions can exist for a given *m/z* feature.
+As shown in Figure \@ref(fig:figure-3-feature-solutions), there can be multiple graphical component solutions that can exist for a given *m/z* feature.
 The component plausibility score was the primary criterion for which components were selected. 

### 2023-11-09 manuscript/manuscript.Rmd
replace 'Bonferroni corrected' with 'Bonferroni correction applied'

@@
 This is especially important in non-targeted metabolomics data sets that can contain thousands of *m/z* features and p-value adjustment is required to account for multiple testing. 
-For example, to detect a significant lower limit of Pearson's correlation coefficient of 0.7 for a matrix containing `r sample_size_n_features` *m/z* features, an approximate minimum sample size of `r sample_size` observations would be required (p < 0.05, Bonferroni corrected) [@cohen2013].
+For example, to detect a significant lower limit of Pearson's correlation coefficient of 0.7 for a matrix containing `r sample_size_n_features` *m/z* features, an approximate minimum sample size of `r sample_size` observations would be required (p < 0.05, Bonferroni correction applied) [@cohen2013].
 Metabolomics fingerprinting and profiling experiments typically require high replication of 10-15 observations per class in order to discriminate between the sample classes [@enot2008]. 

### 2023-11-09 manuscript/manuscript.Rmd
log10 scale the MF plausibility rank axes in Figure 2

@@
 ```{r figure-2-kegg-mf-selection,fig.height=4,fig.cap='Molecular formula plausibility ($P_{MF}$) rankings of KEGG compound molecular formulas.
-The $P_{MF}$ rankings were based on the monoisotopic masses of 6745 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 4 ppm.'}
+The $P_{MF}$ rankings were based on the monoisotopic masses of 6744 valid KEGG compound molecular formulas for which molecular formulas were generated using a mass error of $\\pm$ 4 ppm.'}
 tar_read(KEGG_compound_MF_ranks_mass_plot) +

### 2023-11-09 R/targets/MF_selection_targets.R

@@ -52,6 +52,7 @@ MF_selection_targets <- list(
       ggplot2::ggplot(ggplot2::aes(x = Rank,
                                    y = `Measured M`)) +
       ggplot2::geom_point(alpha = 0.3,size = 1) +
+      ggplot2::scale_x_log10() +
       jfmisc::theme_neat() +
       ggplot2::labs(x = '*P<sub>MF</sub>* rank',
                     y = 'Monoisotopic mass') +
@@ -65,6 +66,7 @@ MF_selection_targets <- list(
       ggplot2::ggplot(ggplot2::aes(x = Rank,
                                    y = N)) +
       ggplot2::geom_point(alpha = 0.3,size = 1) +
+      ggplot2::scale_x_log10() +
       jfmisc::theme_neat() +
       ggplot2::labs(x = '*P<sub>MF</sub>* rank',
                     y = 'Number of generated molecular formulas') +

### 2023-11-10 manuscript/manuscript.Rmd
improve the wording of the sentence about ESI-MS
Clarify sentence about m/z correlations

@@
 Metabolite identification remains one of the main bottlenecks both in time and resources in the analysis of non-targeted mass spectrometry based metabolomics approaches.
-Electrospray ionisation mass spectrometry (ESI-MS) is preferred as this is a soft ionisation technique that enables easier compound identification [@domingo2018].
+Electrospray ionisation mass spectrometry (ESI-MS) is a soft ionisation technique that simplifies compound identification [@domingo2018].
 Irrespectively, it is possible for a single metabolite to be represented within ESI-MS spectra by multiple ionisation product adducts and isotopes [@draper2013].
@@
 The use of correlations to identify relationships between *m/z* features provides an essential layer of data-driven, internal validation of theoretically derived associations.
-For instance, if two features that are posited to have the same composition (eg.
-a C^13^ relationship) but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about associations (eg.
+For instance, if two features are posited to have the same molecular formula based on a mass difference equivalent to ^13^C but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about other associations (eg.
 metabolic pathway changes) of those features with other putatively assigned features within the matrix.

### 2023-11-10 R/targets/brachy_targets.R
increase the RT limit to 2 seconds

@@ -19,6 +19,8 @@ assignment_parameters <- assignment_parameters %>%
     }
   )
 
+assignment_parameters$`RP-UHPLC-HRMS`@RT_diff_limit <- 2/60
+
 brachy_targets <- purrr::map2(
   techniques,
   assignment_parameters,

### 2023-11-11 manuscript/manuscript.Rmd
Add retention time difference limit for LC-MS to adduct and isotopic assignment iteration section
replace unambiguous with singular

@@
 Molecular formula assignment was conducted using the R package *assignments* v`r packageVersion('assignments')` and the approach comprised four major steps which are outlined in Figure \@ref(fig:figure-1-approach).
-Where assignment was possible, *m/z* features were designated with unambiguous molecular formulas, isotopes and ionisation adduct assignments. 
+Where assignment was possible, *m/z* features were designated with singular molecular formula, isotope and ionisation adduct assignments. 
 The following sections outline the methodology behind each major step of the approach.
@@
 This first assignment step used only positively correlated feature relationships that corresponded to adduct and isotopic relationships without transformation. 
+For the C18-UHPLC-HRMS data, these correlated feature relationships were limited to *m/z* features with a retention time difference of up to 2 seconds.
 This assignment step consisted of molecular formula generation, molecular formula selection (Section \@ref(MF-selection)) and then multiple iterations of network graph component assembly and network graph component selection (Section \@ref(component-selection)).

### 2023-11-11 manuscript/manuscript.Rmd
Replace the use of the term metabolome fingerprinting with direct infusion where necessary

@@
 Compound identification remains one of the primary bottlenecks in the analysis of non-targeted mass spectrometry metabolomics data.
-Some annotation approaches are available; however, these are primarily for liquid chromatography-mass spectrometry (LC-MS) profiling and are unsuitable for non-targeted metabolome fingerprinting techniques.
+Some annotation approaches are available; however, these are primarily for liquid chromatography-mass spectrometry (LC-MS) profiling and are unsuitable for non-targeted direct infusion techniques.
 
@@
 We present a data-driven ensemble approach for the ‘first pass’ automated assignment of molecular formulas to mass spectra acquired from high-resolution electrospray ionisation mass spectrometry (ESI-MS).
-This is primarily intended for metabolomics experiment data of biological origin, using non-targeted metabolome fingerprinting and LC-MS profiling.
+This is primarily intended for metabolomics experiment data of biological origin, using non-targeted direct infusion and LC-MS profiling.
 The approach incorporates the use of *m/z* correlation networks, molecular formula generation and selection as well as graphically based component plausibility selection to assign molecular formulas to *m/z* features.
@@
 Ultra-high resolution MS instrumentation utilising Orbitrap mass analyser technology is now commonplace in many laboratories, which can yield many thousands of *m/z* features in a single analytical run [@hu2005].
-This has provided an unprecedented opportunity for metabolite annotation and has been applied to approaches such as 'first pass' flow infusion electrospray high-resolution MS (FIE-HRMS) fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
+This has provided an unprecedented opportunity for metabolite annotation and has been applied to direct infusion approaches like flow infusion electrospray high-resolution MS (FIE-HRMS), commonly applied for 'first pass' metabolome fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
 However, mass accuracy alone does not meet the requirements for metabolite annotation [@kind2006].
@@
 Tandem MS can also be employed for LC-MS approaches and provides a convenient method for structural validation [@neumann2010].
-However, tandem MS is unsuitable for metabolome fingerprinting-based techniques due to the presence of structural isomers and isobaric compounds.
+However, tandem MS is unsuitable for direct infusion based techniques due to the presence of structural isomers and isobaric compounds.
 Many of the available software packages implementing the strategies outlined above are suitable for LC-MS derived data only.
@@
 
-Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS fingerprinting and LC-MS profiling-based metabolomics data.
+Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted, direct infusion and LC, ESI-MS based metabolomics data. 
 This is aimed at biological experiments that include experimental classes with greater than four replicate samples and for the assignment of accurate masses up to 1000 atomic mass units (amu).
@@
 
-The sample matrices were each analysed using FIE-HRMS fingerprinting and C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling.
+The sample matrices were each analysed using FIE-HRMS and C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling.
 These data are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949.
-The FIE-HRMS fingerprinting data were spectrally processed by performing spectral binning using the R package *binneR* v`r packageVersion('binneR')` [@finch2022].
+The FIE-HRMS data were spectrally processed by performing spectral binning using the R package *binneR* v`r packageVersion('binneR')` [@finch2022].
 Spectral processing of the C18-UHPLC-HRMS profiling data was conducted using the R package *xcms* v`r packageVersion('xcms')` [@smith2006].
@@
 For example, to detect a significant lower limit of Pearson's correlation coefficient of 0.7 for a matrix containing `r sample_size_n_features` *m/z* features, an approximate minimum sample size of `r sample_size` observations would be required (p < 0.05, Bonferroni correction applied) [@cohen2013].
-Metabolomics fingerprinting and profiling experiments typically require high replication of 10-15 observations per class in order to discriminate between the sample classes [@enot2008]. 
+Metabolome profiling experiments typically require high replication of 10-15 observations per class in order to discriminate between the sample classes [@enot2008]. 
 Therefore, a sample set comparing just four classes brings this to within a practical range for applying this approach in many cases, such as the example *B.
@@
 
-The approaches of the *MFAssignR* R package and *Formularity* software provide alternative means of assigning molecular formulas to both fingerprinting and LC-MS data, with *Formularity* utilising a compound identification database to aid molecular formula selection [@schum2020;@tolic2017].
+The approaches of the *MFAssignR* R package and *Formularity* software provide alternative means of assigning molecular formulas to both direct infusion and LC-MS data, with *Formularity* utilising a compound identification database to aid molecular formula selection [@schum2020;@tolic2017].
 While these approaches provide the advantage of being able to assign molecular formulas using a single replicate with substantially low computational demands, they do not incorporate the use of correlations between *m/z* features.
@@
 
-Although examples used here include FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using non-targeted ultra-high resolution mass accuracy ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
+Although the examples used here include FIE-HRMS and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using non-targeted ultra-high resolution mass accuracy ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
 Similarly, data from other chromatographic techniques for LC-HRMS profiling could be used such as hydrophilic interaction chromatography (HILIC), where suitable adducts and adduct rankings could instead be specified [@spagou2010].
@@
 
-Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted IPs could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS metabolome fingerprinting.
+Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted IPs could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS.
 It was also possible to assign molecular formulas of common primary and secondary metabolites present in the KEGG compound database to *m/z* features from leaf tissue of *B.
-distachyon* using both FIE-HRMS fingerprinting and C18-UHPLC-HRMS profiling techniques.
+distachyon* using both FIE-HRMS and C18-UHPLC-HRMS profiling techniques.
 

### 2023-11-11 manuscript/manuscript.Rmd
Update the conclusion to include the standards mixture

@@
 
-Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted IPs could be recovered from a subset of chemical standards spiked into an example human urine sample analysed using FIE-HRMS.
+Using correlation networks of *m/z* feature abundances, molecular formula selection and graphical component selection, assigned molecular formulas for predicted IPs could be recovered from a mixture of chemical standards as well as an example human urine sample spiked with this mixture, analysed using FIE-HRMS.
 It was also possible to assign molecular formulas of common primary and secondary metabolites present in the KEGG compound database to *m/z* features from leaf tissue of *B.

### 2023-11-11 manuscript/manuscript.Rmd
correct sentence about mass accuracy for metabolite annotation
Clarify what transformations can be used

@@
 This has provided an unprecedented opportunity for metabolite annotation and has been applied to direct infusion approaches like flow infusion electrospray high-resolution MS (FIE-HRMS), commonly applied for 'first pass' metabolome fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
-However, mass accuracy alone does not meet the requirements for metabolite annotation [@kind2006].
+However, high mass accuracy alone is insufficient for unambiguous metabolite annotation [@kind2006].
 
@@
 The additional input of knowledge for transformations specific to the biological sample can increase assignment success.
-However, the approach could be applied without the use of transformations where these are not relevant.
+These transformations could encompass any biological or non-biological alteration that may occur to the precursor molecules and result in correlated *m/z* that could be used for molecular formula assignment.
+In-source fragmentation would be one such example.
+The approach can alternatively be applied without the use of transformations where these are not relevant.
 The assigned molecular formulas could be putatively matched to compound databases such as KEGG using ionisation rules such as those of the MZedDB database and then apply functional enrichment analyses using a tool such as the R package *FELLA* [@draper2009;@picart2018].

### 2023-11-11 manuscript/manuscript.Rmd
clarify the types of parameters that can be customised

@@
 This data should be spectrally processed and appropriately pre-treated prior to assignment.
-All assignment parameters are fully customisable, which includes the specification of custom adducts, isotopes and transformations.
+All assignment parameters are fully customisable, which includes error thresholds, correlation analysis parameters and the specification of custom adducts, isotopes and transformations.
 It is possible for all the assignment criteria and results to be recovered for each assignment step, enabling the manual curation of assignments by the user.

### 2023-11-11 manuscript/manuscript.Rmd
Add the impact of pre-treatment steps

@@
 Spectral processing of the C18-UHPLC-HRMS profiling data was conducted using the R package *xcms* v`r packageVersion('xcms')` [@smith2006].
-The data were then pre-treated by performing batch correction, feature occupancy filtering, imputation, relative standard deviation filtering based on QC samples and total ion count normalisation using the R package *metabolyseR* v`r packageVersion('metabolyseR')` available from GitHub (https://github.com/jasenfinch/metabolyseR).
+The data were then pre-treated by performing batch correction, feature occupancy filtering, imputation, relative standard deviation (RSD) filtering based on quality control (QC) samples and total ion count (TIC) normalisation using the R package *metabolyseR* v`r packageVersion('metabolyseR')` available from GitHub (https://github.com/jasenfinch/metabolyseR).
 Further details of the ESI-MS analyses and data preparation are available in Additional file 2.
@@
 ```{r adducts}
+tar_load(standards_parameters_molecular_formula_assignment)
 tar_load(brachy_FIE_HRMS_parameters_molecular_formula_assignment)
@@
 The input data consists of a sample by accurate *m/z* feature intensity matrix with at least four samples.
-This data should be spectrally processed and appropriately pre-treated prior to assignment.
+It is recommended that these data are pre-treated prior to assignment.
+Pre-treatment steps such as occupancy or QC based RSD filtering can ensure the robustness of the *m/z* that are considered. 
+Other steps such as batch correction, imputation and TIC normalisation can improve the strength of *m/z* correlations by removing variability introduced during MS analysis.
+
 All assignment parameters are fully customisable, which includes error thresholds, correlation analysis parameters and the specification of custom adducts, isotopes and transformations.
@@
 ```{r assignment-evaluations}
-tar_load(standards_parameters_molecular_formula_assignment)
 tar_load(standards_assignment_outcomes_non_iso)

### 2023-11-11 manuscript/manuscript.Rmd
fix typo
include data preparation information

@@
 
-The input data consists of a sample by accurate *m/z* feature intensity matrix with at least four samples.
-It is recommended that these data are pre-treated prior to assignment.
-Pre-treatment steps such as occupancy or QC based RSD filtering can ensure the robustness of the *m/z* that are considered. 
+The input data simply consists of a sample by accurate *m/z* feature intensity matrix with at least four samples.
+It is recommended that these data are pre-treated prior to assignment and data preparation can be performed by a software tool of the user's preference.
+Pre-treatment steps such as occupancy or QC-based RSD filtering can ensure the robustness of the *m/z* that are considered. 
 Other steps such as batch correction, imputation and TIC normalisation can improve the strength of *m/z* correlations by removing variability introduced during MS analysis.

### 2023-11-11 manuscript/manuscript.Rmd
Correct component removal statement
add reason for isotopic component removal

@@
 For each selected molecular formula, a component subgraph was constructed from the correlation network graph using all matching features, with features allowed across multiple components.
-All components with a size of less than one and containing only isotopic features or adducts were removed.
+All components with a size of less than two were removed.
+Isotopic *m/z* present without their non-isotopic counterparts are unlikely for molecular formulas below 1000 amu.
+Therefore, components containing only isotopic features or isotopic adducts were excluded.
 

### 2023-11-12 manuscript/manuscript.Rmd
improve the row names in Table 3

@@
     relationships = n_relationships,
-    `adduct and isotopic step assignments` = n_ai_assignments,
-    `transformation step assignments` = n_t_assignments,
+    `A\\&I iterations assignments` = n_ai_assignments,
+    `T iterations assignments` = n_t_assignments,
     `assigned \\textit{m/z}` = n_assignments,
@@
     ),
-    caption = 'A summary of the molecular formula assignment results for the example sample matrices.') %>%
+    caption = 'A summary of the molecular formula assignment results for the example sample matrices.
+The \\textit{m/z} assigned for the adduct and isotopic (A\\&I) iterations and the transformation (T) iterations steps are included.') %>%
   kable_styling(

### 2023-11-12 manuscript/manuscript.Rmd
Reword sentence in Section 3.2
correct sentence structure

@@
 Ultra-high resolution MS instrumentation utilising Orbitrap mass analyser technology is now commonplace in many laboratories, which can yield many thousands of *m/z* features in a single analytical run [@hu2005].
-This has provided an unprecedented opportunity for metabolite annotation and has been applied to direct infusion approaches like flow infusion electrospray high-resolution MS (FIE-HRMS), commonly applied for 'first pass' metabolome fingerprinting and liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
+This has provided an unprecedented opportunity for metabolite annotation and has been applied to direct infusion approaches like flow infusion electrospray high-resolution MS (FIE-HRMS), commonly applied for 'first pass' metabolome fingerprinting, as well as liquid chromatography (LC) HRMS profiling for non-targeted metabolomic analyses.
 However, high mass accuracy alone is insufficient for unambiguous metabolite annotation [@kind2006].
@@
 
-While not the intended primary application of the approach, the C18-UPHLC-HRMS data from *B.
-distachyon* had the highest percentage of *m/z* features assigned along with the greatest number of molecular formulas.
+The C18-UPHLC-HRMS data from *B. distachyon* had the highest percentage of *m/z* features assigned along with the greatest number of molecular formulas. 
 The chromatographic separation of the C18-UHPLC-HRMS enables the correlation network to be considered within limited retention time windows and increases the likelihood that a number of correlated IPs will be present for a given compound.

### 2023-11-12 manuscript/manuscript.Rmd
Fix sentence structure

@@
 The use of correlations to identify relationships between *m/z* features provides an essential layer of data-driven, internal validation of theoretically derived associations.
-For instance, if two features are posited to have the same molecular formula based on a mass difference equivalent to ^13^C but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about other associations (eg.
-metabolic pathway changes) of those features with other putatively assigned features within the matrix.
+For instance, if two features are posited to have the same molecular formula based on a mass difference equivalent to ^13^C but are not significantly correlated within the matrix, it would then be illogical to infer biological assumptions about associations (eg.
+metabolic pathway changes) with other putatively assigned features.
 

### 2023-11-12 manuscript/manuscript.Rmd
fix sentence structure

@@
 
-Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted, direct infusion and LC, ESI-MS based metabolomics data. 
+Here, we outline the development of a data-driven, automated approach for 'first pass' putative molecular formula assignment in non-targeted ESI-MS based metabolomics data. 
 This is aimed at biological experiments that include experimental classes with greater than four replicate samples and for the assignment of accurate masses up to 1000 atomic mass units (amu).

### 2023-11-12 manuscript/manuscript.Rmd
Various minor corrections

@@
 A full list of the names and InChI chemical identifiers of the chemical standards can be found in Additional file 1.
-The second was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with the sample mixture of chemical standards as used above.
+The second was human urine purchased from BioIVT (Catalog# HUMANURINEPNN, Lot# 0393918) and spiked with the sample mixture of chemical standards used above.
 The third was healthy leaf tissue samples collected from four ecotypes of the model grass species *Brachypodium distachyon* using the method described in @parker2008.
@@
 
-The sample matrices were each analysed using FIE-HRMS and C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling.
+The sample matrices were all analysed using FIE-HRMS as well as C18 reverse phase ultra-high performance liquid chromatography high-resolution mass spectrometry (C18-UHPLC-HRMS) profiling of the *B.
+distachyon* leaf tissue.
 These data are available via the EMBL-EBI Metabolights database with the study identifier MTBLS6949.
@@
 The incidence of alternative adduct and molecular formula assignment is higher in the spiked urine (Table \@ref(tab:ip-outcomes)).
-However, this is expected as there and are more *m/z* features present and greater numbers of background compounds and therefore greater chance for alternative graphical components to be present in the correlation network.
+However, this is expected as there and are more *m/z* features present with greater numbers of background compounds and therefore a greater chance for alternative graphical components to be present in the correlation network.
 It does not necessarily reflect a poor performance of the graphical component selection approach, there is just a higher probability that these *m/z* could potentially be alternative compounds. 
@@
 The assignments obtained from the *B.
-distachyon* is provided as an example showing the intended real world application of the molecular formula assignment approach to a sample set that includes multiple underlying experimental classes in the correlation network.
+distachyon* matrices are provided as an example showing the intended real world application of the molecular formula assignment approach to a sample set that includes multiple underlying experimental classes in the correlation network.
 A similar number of molecular formulas were assigned to the FIE-HRMS matrices for the spiked human urine and *B.
@@
 The chromatographic separation of the C18-UHPLC-HRMS enables the correlation network to be considered within limited retention time windows and increases the likelihood that a number of correlated IPs will be present for a given compound.
-Added to this, the LC-MS reduces the effects of ion suppression compare to FIE-HRMS and also decreases the chance of potential interference in the correlations between related features due to the presence of structural isomers and isobaric compounds [@gowda2014].
+Added to this, the LC-MS reduces the effects of ion suppression compared to FIE-HRMS and also decreases the chance of potential interference in the correlations between related features due to the presence of structural isomers and isobaric compounds [@gowda2014].
 
@@
 
-Although the examples used here include FIE-HRMS and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using non-targeted ultra-high resolution mass accuracy ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
+Although the examples used here include FIE-HRMS and C18-UHPLC-HRMS profiling techniques, other non-targeted metabolomics techniques using ultra-high resolution ESI-MS data such as nanospray direct infusion would also be suitable [@southam2017].
 Similarly, data from other chromatographic techniques for LC-HRMS profiling could be used such as hydrophilic interaction chromatography (HILIC), where suitable adducts and adduct rankings could instead be specified [@spagou2010].

### 2023-11-14 R/targets/brachy_targets.R
point piggyback input targets to parent repo

@@ -41,7 +41,6 @@ brachy_targets <- purrr::map2(
       .x,
       `FIE-HRMS` = hrmtargets::tar_input_piggyback(
         !!target_name,
-        repo = 'jasenfinch/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS-devel',
         tag = 'brachypodium_FIE-HRMS',
         sample_info_file = 'sample_information.csv'  
       ),

### 2023-11-14 R/targets/standards_targets.R

@@ -66,7 +66,6 @@ standards_processing_targets <- matrix_info %>%
         hrmtargets::tar_input_piggyback(
           !!rlang::sym(.x$sample),
           tag = !!.x$tag,
-          repo = 'jasenfinch/Data_driven_matrix-wide_molecular_formula_assignment_for_high_resolution_ESI-MS-devel',
           sample_info_file = .x$sample_info),
         
         ## Perform spectral binning

### 2023-11-14 manuscript/supplementary_tables.Rmd
Remove 13C2 isotope from Table S3, Additional file 3

@@
 mzAnnotation::isotope_rules() %>% 
-  dplyr::filter(Isotope %in% c('13C','13C2','18O','34S')) %>% 
+  dplyr::filter(Isotope %in% c('13C','18O','34S')) %>% 
   dplyr::select(Isotope,
@@
   dplyr::mutate(Isotope = c('\\textsuperscript{13}C',
-                            '\\textsuperscript{13}C2',
                             '\\textsuperscript{18}O',

### 2023-11-14 manuscript/supplementary_methods.Rmd
fix instrument name in supplementary methods

@@
 
-FIE-HRMS mass spectra were acquired on on a Q-Exactive Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system system for the *B.
+FIE-HRMS mass spectra were acquired on on a Thermo Q-Exactive Plus Orbitrap (ThermoFinnigan, San Jose CA) coupled to a Dionex (ThermoFinnigan, San Jose CA) ultra-performance LC system system for the *B.
 distachyon* samples and the Human urine.
@@
 
-C18-UHPLC-HRMS analyses were performed on an Orbitrap Fusion Tribrid mass spectrometer (Thermo Scientific) that was coupled to a Ultimate 3000 liquid chromatography tower (Dionex, Thermo Scientific). 
+C18-UHPLC-HRMS analyses were performed on an Thermo Orbitrap Fusion Tribrid mass spectrometer (Thermo Scientific) that was coupled to a Ultimate 3000 liquid chromatography tower (Dionex, Thermo Scientific).
 Mass spectra were acquired using a single scan filter for each of the positive and negative ionisation modes between a range of 100-2000 m/z.
